\section{Summary}
Silicon photonics engineer with 20 years in CMOS-compatible photonic integration, spanning datacenter optics, consumer sensing, and defense applications. 53 granted US patents in optical phased arrays and LiDAR, and publications in Nature and Nature Communications (5,000+ citations). Principal Investigator on \$3.9M+ in DOE and DoD SBIRs at Analog Photonics, and co-author on the founding proposal for AIM Photonics, the \$600M+ Manufacturing USA institute. Currently building 800G/1.6T intrarack optical links at AWS; completed Andrew Ng's Deep Learning Specialization.% \vspace{-0.3em}
\section{Experience}
\newenvironment{jobitems}{%
  \par\vspace{0.1em}%
  \begin{itemize}[nosep,leftmargin=1.15em,itemsep=0.2em]%
}{%
  \end{itemize}%
}
% \vspace{-0.5em}
% --- AWS ---
\begin{resumetabular}
2026 \,\, &
\textbf{Amazon Web Services (AWS)}, Optical Network Development Engineer, Cupertino, CA
\begin{jobitems}
  \item Design and validate 800G intrarack optical links (100G/lane, OSFP) deployed in AWS datacenter networks
  \item Implement in-band 1PPS delivery of GPS-sourced timing from TOR to NIC, holding jitter under 10ns over existing high-speed optical links
  \item Build an asynchronous side channel on the same optical link for telemetry and firmware updates, targeting 1-4 Mb/s alongside 1PPS delivery
  \item Define architecture and vendor requirements for next-generation 1.6T links (200G/lane)
\end{jobitems}
\end{resumetabular}

\vspace{0.3em}

% --- Apple ---
\begin{resumetabular}
  20--26 &
  \textbf{Apple Camera Technology Group}, Silicon Photonics Lead, Cupertino, CA
  \begin{jobitems}
    \item Coordinated cross-discipline team (VCSEL, TIA, driver, optics, applications) on R\&D for SMI-based OCT consumer sensing; set system requirements (SNR, noise budget) and made component-level design decisions
    \item Led R\&D on coherent laser Doppler imaging for wearable health sensing
    \item Built beam-scanning rig for 3D OCT capture and set up high-density SMI VCSEL array with coverglass backreflection compensation
    \item Designed FMCW chirp pre- and post-processing to linearize all lasers across the SMI VCSEL array
    \item Led MATLAB platform for end-to-end coherent sensing system simulation
  \end{jobitems}
  \end{resumetabular}

\vspace{0.3em}

% --- Analog Photonics ---
\begin{resumetabular}
14--20 &
\textbf{Analog Photonics}, Director of Advanced Projects and Photonics Integration
\begin{jobitems}
  \item Co-authored winning proposal for AIM Photonics, the \$600M+ Manufacturing USA institute for integrated photonics (\$110M AFRL, \$250M New York State, \$245M+ public/private match), 2015
  \item Primary author of the first industry PDK at AIM Photonics, the foundry design kit used by external silicon-photonics customers; sustained \$1M+ annual AP funding through the institute
  \item Principal Investigator on DOE SBIR program for photonic memory controller module (\$2.8M across Phases I, II, IIB): 30 Gb/s MZI-based athermal optical links and microring WDM toward 800GE on-board optics; multi-company consortium with Freedom Photonics and PLCC
  \item Co-developed 400G-FR4 silicon photonics transceiver chipsets for CPO, OBO, and pluggable form factors (94 Gb/s/mm density, $<$10 pJ/bit, -5.4 dBm OMA at KP4 pre-FEC BER); demonstrated at OFC 2019
  \item Invented two LiDAR architectures: WDM LiDAR combining multi-wavelength sources with optical phased array, and switched-array LiDAR using curved-focal-plane emitters with log2(N) control signals; demonstrated chip-scale in silicon
  \item Principal Investigator on five additional SBIRs (\$1M+ combined) across Army, Navy, and DoD: hybrid MOPA laser at 1550 nm ($>$1 W output, $>$1 GHz/\textmu s chirp) with OPA-based coherent combining, UAV/tactical free-space optical comms, 3D optical sensors for Future Vertical Lift, and silicon photonics optogenetic transceivers
\end{jobitems}
\end{resumetabular}

\vspace{0.3em}

% --- MIT ---
\begin{resumetabular}
11--14 &
\textbf{MIT Photonics Microsystems Group}, Post-doctoral Associate
\begin{jobitems}
  \item Designed, fabricated, and demonstrated DFB and DBR erbium-doped lasers integrated into a CMOS-compatible silicon photonics fab, enabling on-chip light sources without III-V bonding
  \item Developed germanium-on-silicon photodetectors from process up at 1550 nm: $\sim$1 A/W responsivity, 50 GHz bandwidth, nA-scale dark current
  \item Contributed to large-scale 2D nanophotonic optical phased array demonstration (Nature, 2013)
  \item Contributed to ultralow-power athermal silicon modulator demonstration (Nature Communications, 2014)
  \item Designed silicon photonics device library: PN-junction MZI modulators, microdisk and microring filters and modulators, edge couplers, and directional grating couplers
\end{jobitems}
\end{resumetabular}\section{Education}
\vspace{-0.5em}
\begin{resumetabular}
03--11 &
\textbf{Georgia Institute of Technology (Georgia Tech), Atlanta, GA} \\
& MS and PhD, Electrical Engineering (Photonics) | Minor: Physics \\
& Thesis: ``High Quality Integrated Silicon Nitride Nanophotonic Structures for Visible Light Applications.'' \\[0.5em]
99--03 &
\textbf{Aryamehr (Sharif) University of Technology, Tehran, Iran} \\
& BSc, Electrical Engineering | Top 10 in EE, Emphasis: Fields \& Waves in Communications \\
\end{resumetabular}

\section{Continuing Education}
\vspace{-0.5em}
\begin{resumetabular}
23--25 &
\textbf{DeepLearning.AI (Coursera)} \\
& Deep Learning Specialization (Andrew Ng) \\
& Machine Learning Specialization (Andrew Ng) \\
\end{resumetabular}\section{Selected Patents}
\vspace{0.2cm}

\noindent\textbf{Analog Photonics (2019--2024)}
\begin{enumerate}
    \item ``Multi-static coherent LiDAR,'' \href{https://patents.google.com/patent/US12050265B2}{\underline{US Patent 12,050,265 B2}}, 2024.
    \item ``Optical phased array light steering,'' \href{https://patents.google.com/patent/US12085833B2}{\underline{US Patent 12,085,833 B2}}, 2024.
    \item ``Large scale steerable coherent optical switched arrays,'' \href{https://patents.google.com/patent/US20240329324A1}{\underline{Patent Application US20240329324A1}}, 2024.
    \item ``Aberration correction of optical phased arrays,'' \href{https://patents.google.com/patent/US12517351B2}{\underline{US Patent 12,517,351 B2}}, 2026.
    \item ``Photonic die alignment,'' \href{https://patents.google.com/patent/US11693196B2}{\underline{US Patent 11,693,196 B2}}, 2023.
    \item ``Photonics fabrication process performance improvement,'' \href{https://patents.google.com/patent/US11762146B2}{\underline{US Patent 11,762,146 B2}}, 2023.
    \item ``Tunable optical structures,'' \href{https://patents.google.com/patent/US11269236B2}{\underline{US Patent 11,269,236 B2}}, 2022.
    \item ``Optical switching using spatially distributed phase shifters,'' \href{https://patents.google.com/patent/US11422431B2}{\underline{US Patent 11,422,431 B2}}, 2022.
    \item ``Speckle reduction in photonic phased arrays,'' \href{https://patents.google.com/patent/US11353769B2}{\underline{US Patent 11,353,769 B2}}, 2022.
    \item ``Two dimensional optical phased arrays using edge-coupled integrated circuits,'' \\
    \href{https://patents.google.com/patent/WO2022051453A1}{\underline{Patent Application WO2022051453A1}}, 2022.
    \item ``Array-based free-space optical communication links,'' \href{https://patents.google.com/patent/US11196486B2}{\underline{US Patent 11,196,486 B2}}, 2021.
    \item ``Wavelength division multiplexed lidar,'' \href{https://patents.google.com/patent/US11061140B2}{\underline{US Patent 11,061,140 B2}}, 2021.
    \item ``Optical phase shifter device,'' \href{https://patents.google.com/patent/US10809591B2}{\underline{US Patent 10,809,591 B2}}, 2020.
    \item ``Large scale steerable coherent optical switched arrays,'' \href{https://patents.google.com/patent/US10338321B2}{\underline{US Patent 10,338,321 B2}}, 2019.
\end{enumerate}

\noindent\textbf{Apple (2023--2024)}
\begin{enumerate}
    \item Depth sensing using multiple coherent emitters with interleaved VCSEL/detector arrays (\href{https://patents.google.com/patent/WO2024177716A1}{\underline{WO2024177716A1}}, 2024)
    \item Coherent optical waveguide for wearable biometric sensing (\href{https://patents.google.com/patent/US20240057896A1}{\underline{US20240057896A1}}, 2024)
    \item ``LiDAR array with vertically-coupled transceivers,'' \href{https://patents.google.com/patent/US20230366986A1}{\underline{Patent Application US20230366986A1}}, 2023.
\end{enumerate}

\noindent\textbf{MIT (2019)}
\begin{enumerate}
    \item ``Photonic devices and methods of using and making photonic devices,'' \href{https://patents.google.com/patent/US10461489B2}{\underline{US Patent 10,461,489 B2}}, 2019.
    \item ``Apparatus, systems, and methods for waveguide-coupled resonant photon detection,'' \\
    \href{https://patents.google.com/patent/US20170219776A1}{\underline{Patent Application US20170219776A1}}, 2017.
\end{enumerate}
\section{Selected Publications}
\vspace{0.2cm}
\begingroup
\hyphenpenalty=10000
\exhyphenpenalty=10000
\emergencystretch=1.5em

\noindent\textbf{Nature and High-Impact Journals}
\begin{enumerate}
    \item J. Sun, \ldots, E. S. Hosseini, ``Large-scale nanophotonic phased array,'' \href{https://doi.org/10.1038/nature12107}{\underline{\textit{Nature}}}, 2013. (1,670 citations)
    \item E. Timurdogan, \ldots, E. Shah Hosseini, ``An ultralow power athermal silicon modulator,'' \href{https://doi.org/10.1038/ncomms5107}{\underline{\textit{Nature Communications}}}, 2014. (569 citations)
\end{enumerate}

\noindent\textbf{Silicon Photonics and Integrated Optics}
\begin{enumerate}
    \item E. Shah Hosseini, \ldots, ``High quality planar silicon nitride microdisk resonators for integrated photonics in the visible wavelength range,'' \href{https://doi.org/10.1364/OE.17.014596}{\underline{\textit{Optics Express}}}, 2009. (219 citations)
    \item J. Sun, \ldots, E. S. Hosseini, ``Large-scale silicon photonic circuits for optical phased arrays,'' \href{https://doi.org/10.1109/JSTQE.2013.2283793}{\underline{\textit{IEEE JSTQE}}}, 2013. (210 citations)
    \item E. Shah Hosseini, \ldots, ``Systematic design and fabrication of high-Q single-mode pulley-coupled planar silicon nitride microdisk resonators at visible wavelengths,'' \href{https://doi.org/10.1364/OE.18.012725}{\underline{\textit{Optics Express}}}, 2010. (190 citations)
    \item A. H. Atabaki, E. S. Hosseini, \ldots, ``Optimization of metallic microheaters for high-speed reconfigurable silicon photonics,'' \href{https://doi.org/10.1364/OE.18.019329}{\underline{\textit{Optics Express}}}, 2010. (158 citations)
    \item J. Sun, E. S. Hosseini, \ldots, ``Two-dimensional apodized silicon photonic phased arrays,'' \href{https://doi.org/10.1364/OL.39.003476}{\underline{\textit{Optics Letters}}}, 2014. (128 citations)
    \item J. D. B. Bradley, E. S. Hosseini, \ldots, ``Monolithic erbium- and ytterbium-doped microring lasers on silicon chips,''\newline
    \href{https://doi.org/10.1364/OE.22.012140}{\underline{\textit{Optics Express}}}, 2014. (117 citations)
    \item Purnawirman, \ldots, E. S. Hosseini, ``C- and L-band erbium-doped waveguide lasers with wafer-scale silicon nitride cavities,'' \href{https://doi.org/10.1364/OL.38.003760}{\underline{\textit{Optics Letters}}}, 2013. (99 citations)
    \item E. S. Hosseini, \ldots, ``CMOS-compatible 75 mW erbium-doped distributed feedback laser,'' \href{https://doi.org/10.1364/OL.39.026392}{\underline{\textit{Optics Letters}}}, 2014. (88 citations)
    \item Purnawirman, \ldots, E. S. Hosseini, ``Ultra-narrow-linewidth Al$_2$O$_3$:Er$^{3+}$ lasers with a wavelength-insensitive waveguide design on a wafer-scale silicon nitride platform,''
    \href{https://doi.org/10.1364/OE.25.013705}{\underline{\textit{Optics Express}}}, 2017. (56 citations)
    \item Z. Su, \ldots, E. S. Hosseini, ``Four-port integrated polarizing beam splitter,'' \href{https://doi.org/10.1364/OL.39.005451}{\underline{\textit{Optics Letters}}}, 2014. (56 citations)
\end{enumerate}

\noindent\textbf{Foundry and Systems (2018)}
\begin{enumerate}
    \item E. Timurdogan, \ldots, E. S. Hosseini, ``AIM process design kit (AIMPDKv2.0): Silicon photonics passive and active component libraries on a 300mm wafer,''\newline
    \href{https://doi.org/10.1364/OFC.2018.M3F.1}{\underline{\textit{OFC}}}, 2018. (54 citations)
    \item C. V. Poulton, \ldots, E. Hosseini, ``High-performance integrated optical phased arrays for chip-scale beam steering and LiDAR,'' \href{https://doi.org/10.1364/CLEO_AT.2018.ATu3R.2}{\underline{\textit{CLEO}}}, 2018. (50 citations)
\end{enumerate}

\endgroup
