\chapter{One-night optical systems interview review}
\label{app:interview-review}

Use this appendix as a standalone review sheet. Its purpose is not to cover the
most optics. Its purpose is to make your engineering process visible under
time pressure.

\section{The answer spine}

Move every answer through the same sequence. The diagram is a memory aid, not
the answer. In the interview you should be able to speak one clear paragraph
for each arrow, naming what you know, what is still uncertain, and what
measurement you would take next. Do not jump from a symptom to a component.
The systems loop in \Cref{sec:systems-engineering-loop}, the debugging pyramid
in \Cref{sec:debug-pyramid}, and the failure-analysis method in
\Cref{ch:failure-modes} are the full versions of this spine.
\[
\begin{split}
\text{requirements} &\longrightarrow \text{architecture}
\longrightarrow \text{measurements}\\
&\longrightarrow \text{observations}
\longrightarrow \text{hypotheses}\\
&\longrightarrow \text{isolation}
\longrightarrow \text{root cause}\\
&\longrightarrow \text{corrective action}
\longrightarrow \text{recurrence control}.
\end{split}
\]

\paragraph{Requirements.}
Start by stating what the system must do before you name a part. Reach, lane
rate, aggregate capacity, power envelope, lifetime, cost, and manufacturing
volume are the constraints that decide everything downstream. A 500~m DR
link\termnote{DR} at 100~G/lane with a tight power budget is a different
problem from a 2~km FR\termnote{FR} link or a dense WDM\termnote{WDM}
co-packaged engine. Say the requirement out loud even when the interviewer
hands you a failed module, because the requirement tells you which margins
matter and which architectures were even eligible. If you cannot name the
requirement, you cannot tell whether a measurement is relevant.

\paragraph{Architecture.}
Translate the requirement into an architecture path: fiber plant, wavelength,
source, modulator, receiver, and digital stack. A VCSEL\termnote{VCSEL} path
commits you to 850~nm multimode fiber and direct modulation. A
DFB\termnote{DFB} or EML\termnote{EML} path at 1310~nm commits you to
single-mode fiber and leaves a choice among direct modulation, electro-
absorption, a silicon MZM\termnote{MZM}, or a ring\termnote{MRM}. Each path
then sets detector material, thermal control, test coverage, and service
policy. Architecture is where you show that component choices are not
preferences; they are consequences of the requirement
(\Cref{tab:source-mod-matrix}).

\paragraph{Measurements.}
Name the instruments you would use and the uncertainty each one removes.
Always state the reference plane, pattern, temperature, and pass criterion
with the instrument. A number without a plane is not a measurement.

A power meter answers whether the optical power ledger moved, and at which
reference plane. An LIV\termnote{LIV} setup answers whether the laser itself
changed threshold or slope. An OSA\termnote{OSA} answers wavelength and
SMSR\termnote{SMSR}.

A DCA\termnote{DCA} answers OMA\termnote{OMA}, ER\termnote{ER},
RLM\termnote{RLM}, TDECQ\termnote{TDECQ}, and eye shape. A
BERT\termnote{BERT} and FEC\termnote{FEC} counters answer the
BER\termnote{BER} waterfall, floor, and burst pattern.

A VNA\termnote{VNA} answers electrical or electro-optic bandwidth. An
ORL\termnote{ORL} meter and inspection scope answer whether the plant is
reflecting or dirty.

\paragraph{Observations.}
Report what the instruments actually showed, not what you hoped they would
show. Separate facts from interpretation. ``Received power held at
$-4$~dBm while pre-FEC BER rose from $10^{-8}$ to $10^{-5}$ at $70^\circ$C''
is an observation. ``The laser is dying'' is not. Scope the observation:
one lane or all lanes, one unit or a lot, sudden or gradual, recoverable on
cool-down or permanent. Preserve telemetry, screenshots, and the failing
state before you reseat, clean, reboot, or rewrite a calibration table.
Observations that disappear under debugging are observations you never had.

\paragraph{Hypotheses.}
Turn the observations into a short list of competing causes, ranked by how
well they explain the full pattern. Use the power-versus-signal-quality fork
as the first split. If average power moved, hypothesize source output,
coupling, connector loss, multiplexer loss, or monitor calibration. If power
held but BER worsened, hypothesize eye closure from ER, OMA, RLM, or
TDECQ\termnote{TDECQ} degradation; RIN\termnote{RIN} under reflection; ISI
or jitter; wavelength walking off a filter; receiver sensitivity loss; or a
wrong calibration table. Keep the list short enough that the next measurement
can kill more than one item. A hypothesis you cannot falsify with a bench
step does not belong on the list yet.

\paragraph{Isolation.}
Choose the next measurement for its ability to separate the remaining
hypotheses, not because it is the instrument you like. A golden swap of
transmitter versus receiver splits Tx from Rx. An optical-versus-electrical
lane swap in a multi-lane module splits the optical path from the driver or
TIA\termnote{TIA}. A bias sweep at the failing temperature asks whether the
optimum moved away from the stored table. An LIV\termnote{LIV} compared with
ship data asks whether the device aged or only the setpoint drifted. A
BER-versus-power waterfall asks whether the curve shifted or floored.
Isolation is the discipline of picking the cut that removes the most
uncertainty per hour of bench time.

\paragraph{Root cause.}
State the mechanism that survived isolation, with the evidence that killed
the alternatives. ``EAM\termnote{EAM} bias table segment wrong above
$60^\circ$C'' is a root cause. ``Bad eye at high temperature'' is still a
symptom. Name the physical or process mechanism when you can: facet wear,
monitor-PD corruption, FAU\termnote{FAU} misalignment, MPI\termnote{MPI}
from a dirty connector pair, TEC\termnote{TEC} railed under load, supplier
lot with high threshold. If the evidence only reaches ``calibration drift''
and not a deeper mechanism, say so. Overclaiming the root cause is worse than
stopping one layer early with honesty.

\paragraph{Corrective action.}
Fix the mechanism you named, under the condition that failed. Retune the
table, replace the lot, change the thermal design, clean and inspect the
plant, filter the supply, or rewrite the ATP\termnote{ATP} limit. Verify the
fix by reproducing the original failing condition: same temperature, same
host, same fiber, same pattern. A fix that only works on a quiet room-
temperature bench is not a fix. Containment comes first when the fleet is
already exposed: stop ship, quarantine lots, or derate while the permanent
fix is built.

\paragraph{Recurrence control.}
Close with the control that stops the same escape next month. That control
is usually a new or tightened test, an SPC\termnote{SPC} chart, a telemetry
alarm, a process-control limit, a supplier screen, or a firmware guard that
refuses to boot with a railed actuator. Name the owner and the measurable
closure criterion. Root cause without recurrence control is a story, not an
engineering result. In the interview, ending on recurrence control is what
separates a debug narrative from a product-engineering answer.

\section{Ten concepts to know cold}

\subsection{Start at the system}

For a new link, walk downward from capacity to the service model. Capacity
sets lane rate. Lane rate and reach together set the fiber plant and the
source class. Power and cost then decide whether you can afford a
TEC\termnote{TEC}, a retimer, or a dense WDM\termnote{WDM} engine. Only after
those constraints are named do you choose the modulator and the receiver,
then the DSP\termnote{dspserdes} and FEC\termnote{FEC} stack, then the
validation and field-service model.
\[
\begin{split}
\text{capacity} &\longrightarrow \text{lane rate}
\longrightarrow \text{reach}
\longrightarrow \text{power}\\
&\longrightarrow \text{fiber plant}
\longrightarrow \text{source}
\longrightarrow \text{modulator}\\
&\longrightarrow \text{receiver}
\longrightarrow \text{digital signal processing and FEC}\\
&\longrightarrow \text{validation and service model}.
\end{split}
\]

A good answer explains why each choice constrains the next one. A
VCSEL\termnote{VCSEL} path points toward 850~nm, multimode
fiber\termnote{MMF}, silicon detection, and direct modulation with a short
reach. A DFB\termnote{DFB} path at 1310~nm points toward single-mode
fiber\termnote{SMF}, germanium detection, and a choice among a directly
modulated laser\termnote{DML}, an EML\termnote{EML}, or an external
silicon-photonic modulator. Neither path is ``better.'' Each closes a
different reach, power, cost, and manufacturing problem. The decision matrix
is in \Cref{tab:source-mod-matrix}.

\subsection{Scope before root cause}

Before you open a instrument, ask how wide the failure is. One lane, one
module, one lot, one vendor, one rack, or the fleet? A new failure or a
long-running trend? Sudden, gradual, intermittent, or temperature-dependent?
Did firmware, supplier, process, fixture, workload, or environment change
just before the symptom appeared? Scope often removes more hypotheses than
the first bench measurement, because a fleet-wide gradual drift cannot be a
single dirty connector, and a single-lane sudden failure after a rework
cannot be a population aging problem.

Preserve the failing state and its telemetry before you reseat, clean,
reboot, or change calibration. Capture CMIS\termnote{CMIS} monitors, pre-
FEC\termnote{FEC} BER\termnote{BER}, bias currents, temperatures,
LOS\termnote{LOS} and LOL\termnote{LOL} flags, and firmware versions. An
intermittent that disappears under debugging is still a real failure; you
just destroyed the evidence
(\Cref{sec:fleet-triage,tab:failure-analysis-checklist}).

\subsection{Use the power-versus-signal-quality fork}

First ask whether received optical power changed. That one question splits
the debug tree.

If power changed, the power ledger moved. Inspect source output and enable
state, coupling and FAU\termnote{FAU} alignment, connector cleanliness,
ORL\termnote{ORL}, fiber and multiplexer loss, and monitor-photodiode
calibration. An APC\termnote{APC} loop that trusts a drifting monitor will
hold a wrong launch power while reporting that everything is fine. Confirm
with an external power meter at a named reference plane.

If power held but BER worsened, the loss is in signal quality or in the
receiver. Inspect OMA, ER, RLM, and TDECQ\termnote{TDECQ} on a DCA;
RIN\termnote{RIN} under controlled ORL; jitter and ISI; wavelength alignment
to filters or rings; receiver sensitivity and equalization; and calibration
tables that set modulator bias. Also ask the receiver-side equivalent: did
the required optical power increase even though the power meter reading did
not? A sensitivity shift can look exactly like transmitter degradation until
you do a golden swap (\Cref{sec:debug-fork,sec:validation-fork}).

\subsection{Track four margin ledgers}

Links rarely fail from one dramatic excursion. They fail when several small
shifts spend different ledgers at once.

\begin{description}
\item[Power] Launch power, insertion loss, coupling, connector loss,
      multiplexer loss, and receiver sensitivity. Track average power and
      OMA\termnote{OMA} separately: average power can look healthy while OMA
      collapses.
\item[Noise] RIN\termnote{RIN}, receiver thermal noise, shot noise,
      crosstalk, supply noise through weak PSRR, and MPI from reflective
      interfaces. Noise that scales with the signal makes a BER floor that
      more launch power cannot fix.
\item[Timing] Bandwidth, dispersion, jitter, ISI, and equalization reserve.
      A SerDes\termnote{SerDes} that is already using most of its FFE and DFE
      taps has little timing margin left even if the eye still opens on a DCA.
\item[Spectral] Wavelength, SMSR\termnote{SMSR}, filter or ring passband,
      thermal drift, and lock range. A ring whose heater is near its DAC rail
      is spending spectral margin even while BER still passes.
\end{description}

Keep the four ledgers as a mental checklist when you narrate a failure. Name
which ledger moved, by how much, and what spent it. The complete treatment is
in \Cref{sec:laser-margin-erosion}.

\subsection{Use the validation ladder}

Validation is staged uncertainty reduction. For every stage, name the
question the stage answers and the uncertainty it removes. A test that
answers no question is cost, not confidence.

\begin{description}
\item[Bring-up] Does the hardware power, initialize, emit, lock, and pass
      data on a known-good host? Removes integration unknowns only.
\item[Characterization] How does performance vary with temperature,
      voltage, lane, wavelength, pattern, and unit? Maps the population, not
      a hero sample. Record LIV, spectrum, RIN, OMA, ER, RLM, and
      TDECQ\termnote{TDECQ} as distributions.
\item[Margin testing] How far is the design from each failure boundary?
      Push power, temperature, ORL\termnote{ORL}, and supplies to the cliff
      and measure the distance in dB, degrees, and volts.
\item[Interoperability] Does it work with target hosts, production fiber,
      real modules, and real firmware? Removes combination failures that
      golden benches miss.
\item[Environmental testing] What changes under temperature cycling,
      humidity, vibration, shock, and connector mating?
\item[Reliability qualification] Which parameters drift with stress and
      time? HTOL\termnote{HTOL}, with a named mechanism and justified
      activation energy, projects wear-out as a FIT\termnote{FIT} number.
\item[Production readiness] Can the ATP\termnote{ATP} catch bad units and
      process drift at useful cost and cycle time, with station correlation
      and guard bands sized to repeatability?
\item[Fleet monitoring] Which telemetry catches margin erosion before
      service failure, and does the RMA\termnote{RMA} Pareto match the life
      model?
\end{description}

\Cref{tab:ladder} maps these stages to activities and instruments. The
worked interview answer in \Cref{sec:interview-worked-validation} walks the
same ladder as spoken prose.

\subsection{Know what each instrument answers}

Do not recite instrument names. Say what uncertainty each one removes.

\begin{description}
\item[Power meter] Did optical power change, and at which reference plane?
      External meters catch monitor-PD lies.
\item[LIV setup] Did laser\termnote{LIV} threshold, slope efficiency,
      voltage, kink behavior, or thermal rollover change relative to ship
      data?
\item[Optical spectrum analyzer or wavemeter] Did wavelength,
      SMSR\termnote{SMSR}, mode structure, or spectral width change on an
      OSA? Essential for filter and ring alignment.
\item[RIN setup] Is transmitter intensity noise\termnote{RIN} limiting BER,
      and does it depend on ORL or on the product bias board rather than the
      laser itself?
\item[Digital communication analyzer] Are OMA, ER, RLM,
      TDECQ\termnote{TDECQ}, jitter, and eye shape acceptable on a DCA? This
      is the transmitter quality ledger in one instrument.
\item[Bit-error-ratio tester and FEC counters] What is the BER waterfall,
      floor, burst pattern, lane correlation, and dwell-time behavior from a
      BERT\termnote{BERT}? FEC histograms separate Gaussian noise from
      clustered MPI or intermittents. The next subsection walks each shape
      in study detail.
\item[Vector network analyzer] Is electrical or electro-optic bandwidth,
      impedance, or reflection causing eye closure? Use a
      VNA\termnote{VNA}.
\item[ORL meter and inspection scope] Is the optical plant adding loss or
      reflection\termnote{ORL} that the laser will convert into RIN or MPI?
\item[Thermal chamber] Is behavior reversible with temperature, and which
      actuator (TEC\termnote{TEC}, heater, bias DAC) loses headroom at the
      failing corner?
\item[Bias sweep] Where is the optimum operating point for the
      EAM\termnote{EAM}, MZM, or ring, and did its curve move away from the
      stored calibration table?
\end{description}

Always name the reference plane, setup calibration, test pattern, bandwidth,
temperature, sample identity, and pass criterion with the result
(\Cref{tab:measurement-mapping,tab:laser-meas}).

\subsection{Read a BER waterfall: shift, floor, and burst pattern}
\label{sec:interview-waterfall}

These three words show up in almost every debug answer. Know what each one
looks like on the bench, what it rules in, and what it rules out. The
operational procedures are in \Cref{sec:fm-ber-increase,sec:fm-ber-floor}.

\paragraph{What a BER waterfall is.}
A BER waterfall is a plot of bit error ratio versus received optical power.
You build it by sweeping a calibrated VOA\termnote{VOA} in the path, holding
pattern, temperature, and host fixed, and counting errors long enough at each
power that the BER\termnote{BER} is statistically meaningful. Plot received
power on the horizontal axis (name the reference plane: usually TP3 at the
receiver connector) and pre-FEC BER on a log vertical axis. A healthy link
falls steeply as power rises: more photons, better signal-to-noise ratio,
fewer errors. That falling curve is the waterfall. A single BER point at the
operating power is not a waterfall. Without the sweep you cannot tell whether
you are short on power or limited by noise that scales with the signal.

\paragraph{What a shifted waterfall means.}
A parallel shift means the whole curve moved left or right while keeping a
similar slope. The link still improves when you add power; it just needs a
different power to hit the same BER. A rightward shift (worse sensitivity)
means you now need more received power for the same pre-FEC BER. Common causes
are lost launch or coupling (power ledger), a quieter or noisier receiver
(TIA\termnote{TIA} noise, responsivity), eye closure that lowers effective
OMA\termnote{OMA} at constant average power (ER\termnote{ER}, RLM\termnote{RLM},
TDECQ\termnote{TDECQ}), wavelength walking onto a filter edge, or equalizer
misadaptation. A leftward shift means the link got healthier. The diagnostic
move after you see a shift is a golden swap: known-good transmitter, then
known-good receiver, to decide which side owns the sensitivity change. Raising
launch power can still help a shifted link, because the waterfall has not
stopped responding to power.

\paragraph{What a BER floor means.}
A floor is a horizontal asymptote: as you raise received power, BER improves
for a while and then stops. More power no longer helps. That happens when the
dominant noise scales with the signal itself. Relative intensity
noise\termnote{RIN} is the textbook case: intensity fluctuations grow in
proportion to optical power, so the signal-to-noise ratio saturates and
\[
Q_{\max}=1/\sqrt{\mathrm{RIN}\cdot\mathrm{BW}}.
\]
Multipath interference\termnote{MPI}, bias-rail noise through weak
PSRR\termnote{PSRR}, and aggressor crosstalk that tracks the aggressor power
make the same shape. The interview trap is to keep raising launch power or
cleaning for loss when the curve has already floored. The fix is to remove the
multiplicative noise source (ORL\termnote{ORL} and isolator hygiene, supply
filtering, connector reflections, crosstalk isolation), not to buy more
photons. Confirm the floor with a full sweep before you name the mechanism;
then bisect with a quiet laser source versus the product bias board, an ORL
reflector sweep, and the FEC error timing below.

\paragraph{What a burst pattern means.}
Average BER alone hides how errors arrive in time. A burst pattern means
errors cluster: many errored symbols in a short window, then quiet intervals,
rather than a steady sprinkle of random bit flips. FEC\termnote{FEC}
histograms and pre-FEC error counters with timestamps make this visible.
Gaussian thermal noise and well-behaved RIN tend to spread errors. MPI from a
pair of reflective interfaces, connector intermittents, ESD events, supply
glitches, and unlocked CDR\termnote{CDR} intervals tend to cluster them. Lane-
correlated bursts across a module point at a shared supply, clock, or thermal
event. A single-lane burst pattern points at that lane's optical path or
connector. In the fleet, bursty pre-FEC counters with stable average power are
often intermittents: preserve the counters and CMIS\termnote{CMIS} event log
before you reseat anything, or the evidence disappears
(\Cref{sec:fm-intermittent}).

\paragraph{How to say the three together.}
``I sweep BER versus received power. A parallel shift is a sensitivity or
power-ledger problem; more power can still help. A floor is multiplicative
noise; more power cannot help. A bursty FEC histogram means clustered
impairments such as MPI or intermittents, not plain Gaussian noise.'' Offer
to draw the shifted curve next to the floored curve on the whiteboard.

\subsection{Separate thermal response from aging}

Thermal effects are usually reversible when you return to the starting
temperature: wavelength shift, ring detuning, EAM\termnote{EAM} or
MZM\termnote{MZM} bias movement, TEC\termnote{TEC} loading, and
receiver-noise increase. Aging is cumulative and does not reverse on
cool-down: threshold rise, slope loss, contact degradation, defect growth,
and permanent absorption or spectral change.

The practical test is simple. Return to the starting temperature and compare
with pre-stress data at the same junction temperature. Full recovery supports
a thermal or control hypothesis, often a calibration-table segment or an
actuator that ran out of range. A permanently moved LIV\termnote{LIV}
baseline supports aging, damage, or assembly change. Mixing the two owners
wastes weeks: a reliability team cannot fix a wrong table, and a firmware
team cannot fix a worn facet (\Cref{sec:laser-thermal-vs-aging}).

\subsection{Accelerated life tests need a mechanism}

HTOL\termnote{HTOL} predicts field life only if elevated temperature and
bias accelerate the same physical mechanism the fleet will see. The stress
must not introduce a different mechanism, such as solder creep at a
temperature the product never reaches, and it must not omit a dominant field
stress, such as thermal cycling or connector wear.

In the interview, state the assumed mechanism, the activation energy and why
it applies to this process, the sample size and confidence bound, the stress
and use temperatures, the measured drift parameter (threshold, slope,
wavelength), and the field-data check. Treat the projected FIT\termnote{FIT}
result as a hypothesis that field returns must confirm or falsify. A FIT
number without those pieces is a spreadsheet, not a prediction. The life
acceleration model is Arrhenius\termnote{Arrhenius}
(\Cref{sec:laser-aging}).

\subsection{Calibration is part of the product}

The operating points a product actually stores are as much the product as the
die. Know them by name: laser bias or APC\termnote{APC} target;
EAM\termnote{EAM} bias versus temperature; MZM\termnote{MZM} quadrature;
ring heater or wavelength-lock point; monitor-photodiode\termnote{PD}
calibration; and receiver and power-monitor offsets. Tables are usually
segmented by temperature, and segment boundaries are a real failure mode.

Recalibrate when evidence demands it, not by habit: loop residual stops
converging, an actuator approaches its rail, telemetry disagrees with an
external reference, temperature leaves the table range, or repair, rework,
or firmware changes invalidate coefficients. ATP\termnote{ATP} must verify
calibration at the temperature corners the fleet will see, not only at
station ambient (\Cref{sec:laser-calibration}).

\subsection{Corrective action must prevent recurrence}

Root cause is not the end of the answer. Close with immediate containment;
the design, process, calibration, firmware, or supplier correction;
verification under the original failing condition; an acceptance test,
process control, alarm, or telemetry control that catches recurrence; and an
owner with a measurable closure criterion. An interview answer that stops at
``we replaced the laser'' sounds like repair. An answer that ends on the new
ATP\termnote{ATP} corner or the new telemetry alarm sounds like product
engineering.

\section{Two stories to prepare}

Each story is a spoken version of the answer spine. Prepare one component or
bench story and one system, production, or fleet story. Use only work you
personally performed, and separate your contribution from the team's. Walk
the same eight beats every time:

\begin{enumerate}
\item Requirement and system context: what the link or product had to do.
\item Observed symptom and scope: what failed, how wide, sudden or gradual.
\item Competing hypotheses: the short list after the power fork.
\item Measurements chosen and why: which instrument, which uncertainty.
\item Evidence that eliminated each hypothesis.
\item Root cause: the mechanism that survived, with evidence.
\item Corrective action: the fix, verified under the failing condition.
\item Recurrence control and measured result: the test, alarm, or process
      change, and the metric that proved it worked.
\end{enumerate}

\section{Questions to rehearse aloud, with model answers}

Rehearse these until the structure is automatic. Each answer starts with scope
or requirements, names measurements and the hypotheses they separate, and ends
with a control. Do not recite; adapt the skeleton to the follow-up questions.

\subsection{How would you validate a new optical transmitter from bring-up
through production?}
\label{sec:interview-worked-validation}

This is the question most likely to open the interview, so it gets the full
treatment. The frame to state before any detail: validation is staged
uncertainty reduction. Each stage answers a question the previous stage could
not, and a test that answers no question is cost, not confidence. Walk the
stages in order, and at every one name the reference plane where the
measurement is taken and the criterion that decides pass
(\Cref{tab:ladder}).

\paragraph{Stage 1: bring-up.}
For bring-up, get one unit alive on a known-good bench and remove the
integration unknowns. Power the device through its specified sequence and
confirm rails, inrush, and power state transitions. Initialize the firmware and
read identity and status over the management
interface\termnote{CMIS}, because a part that cannot report its own state
cannot be debugged later. Enable the laser and confirm first light on a
calibrated power meter at the specified output reference plane, not at
whatever connector is convenient. Then close a real loop: lock the receiver,
run a PRBS\termnote{PRBS} pattern from a BERT\termnote{BERT} or host SerDes,
and count errors. The uncertainty removed is basic: the design functions, the
chips talk to each other, and the test setup itself is sane. The trap is that
everything here is golden: pristine connectors, one temperature, one host, one
unit. Passing bring-up says nothing about the population, the corners, or the
lifetime. Its only job is to make the next stage's failures meaningful.

\paragraph{Stage 2: characterization.}
For characterization, map what the design actually does across its operating
range, on enough units to see variation. Sweep temperature, supply voltage,
bias, and test pattern, and at each point record the transmitter's vital
signs: the LIV curve\termnote{LIV} for threshold, slope efficiency, and
thermal rollover; the optical spectrum on an OSA\termnote{OSA} for center
wavelength and SMSR\termnote{SMSR}; RIN\termnote{RIN} measured at the stated
ORL\termnote{ORL}, because a quiet-bench RIN number is flattery; and the eye
on a DCA\termnote{DCA} for ER\termnote{ER}, OMA\termnote{OMA}, and
TDECQ\termnote{TDECQ}. The key discipline is to record distributions, not a
golden sample: ten units across two lots tell you what the process gives you,
one hero unit tells you what marketing wants. The uncertainty removed is where
the population sits relative to the spec at every corner. The output is not a
pass stamp; it is the dataset that will later set production test limits and
telemetry alarm thresholds.

\paragraph{Stage 3: margin testing.}
For margin, stop asking whether the part passes and ask how far it sits from
the cliff. Push each parameter to its failure boundary: sweep received power
down with a calibrated VOA\termnote{VOA} until the pre-FEC BER\termnote{preFEC}
waterfall crosses the FEC threshold; take case temperature to the limit and
past it; pull supply rails to their corners; degrade ORL with a controlled
reflector and watch the error floor. Measure the distance to failure in dB,
degrees, and volts, then compare it against the variation found in
characterization. The uncertainty removed is the one that kills fleets: a
design can pass every nominal test while sitting half a decibel from its
limit, and normal lot spread plus aging will spend that half decibel in the
first year. Margin testing also reveals which of the four margins, power,
noise, timing, or spectral, runs out first, which tells you what telemetry
must watch (\Cref{sec:laser-margin-erosion}).

\paragraph{Stage 4: interoperability.}
For interoperability, take away everything golden. Replace the BERT with the
target hosts, plural, because SerDes equalizers from different vendors adapt
differently to the same transmitter. Replace the lab jumper with production
fiber plant: real connector losses, real MPO\termnote{MPO} polarity, realistic
ORL from the connectors the datacenter will actually use. Run real link
bring-up sequences, firmware interactions, and traffic patterns rather than
one endless PRBS. The uncertainty removed lives in combinations rather than
components: a transmitter and a receiver that each meet spec can still fail
together, because the standard's transmitter eye and stressed-receiver
tests\termnote{SECQ} only guarantee interoperability if both sides were tested
against the specified reference, and reference receivers are idealizations.
Interop failures found here cost a lab week; found in production they cost a
fleet rollback.

\paragraph{Stage 5: environmental stress and reliability qualification.}
For qualification, ask which mechanisms move with stress and time. Each stress
is chosen for the failure mechanism it accelerates, not run as a ritual:
temperature cycling exercises solder joints, die attach, and fiber attach;
damp heat exercises corrosion and delamination; vibration and shock exercise
the mechanical path; HTOL\termnote{HTOL} accelerates the wear-out of the
active region so that weeks of stress project years of field life through an
Arrhenius model\termnote{Arrhenius}. State the projection's fine print
without being asked: the activation energy must be justified for this
mechanism on this process, the sample size sets the confidence interval, and
the projection holds only if the stress accelerates the same mechanism the
field will see. The uncertainty removed is the shape of lifetime: infant
mortality rate for the burn-in decision, and wear-out rate as a
FIT\termnote{FIT} number the fleet math can consume.

\paragraph{Stage 6: production readiness.}
For production, the question changes from is the design good to can the
factory tell a good unit from a bad one, at rate, at acceptable cost. Write
the ATP\termnote{ATP} so that every limit traces to the link budget or to the
characterization data; a limit nobody can trace will be waived under ship
pressure. Correlate stations by running the same golden units on every
station and site, so a limit means the same thing everywhere. Size guard
bands from measurement repeatability, because a test that cannot repeat
tighter than its limit ships escapes and scraps good units in equal measure.
Verify calibration at the temperature corners the fleet will see, not only at
station ambient. Archive raw data, instrument identity, and
calibration-table versions so any shipped result can be replayed. Then keep
SPC\termnote{SPC} charts on threshold current, power, and wavelength, because
process drift between qualification lots is caught here or not at all.

\paragraph{Stage 7: fleet telemetry and the feedback loop.}
Deployment is the last validation stage, not the end of validation. Fleet
telemetry watches the margins that stage 3 identified: per-lane transmit
power, bias current, pre-FEC BER, temperature, and actuator drive such as
TEC\termnote{TEC} current, with alarms on trends and disagreements rather
than only on hard thresholds. Field returns close the loop: the
RMA\termnote{RMA} Pareto is compared against the qualification projection,
and divergence means the life model was wrong, not that the fleet is unlucky.
A high NFF\termnote{NFF} rate is itself a finding, usually pointing at
intermittents or triage gaps. The uncertainty removed is the honest one: what
every earlier stage missed.

\paragraph{How to say it aloud.}
In the interview, give the frame in one sentence, then walk the seven stages
in order, one sentence each, naming for each stage one instrument and the
uncertainty it removes. Then offer to go deep on whichever stage the
interviewer cares about. That structure shows judgment before detail, and it
turns a memorized list into a conversation.

\subsection{BER worsens at high temperature but average power is stable. What do
you do?}
\label{sec:interview-worked-hot-ber}

This is a classic fork question. The frame to state first: average power held
means the power ledger is intact, so the loss lives in signal quality or in a
receiver that got noisier. Scope before instruments, measure at the failing
temperature, and end with the control that stops recurrence
(\Cref{sec:fm-temperature}).

\paragraph{Step 1: scope the failure.}
Ask how wide the problem is before touching a knob. One unit or many? One lane
or all lanes? One host or every host? One lot or the population? Then ask
whether the BER recovers when the unit is cooled. Recovery makes it a thermal
operating-point problem, not aging. Aging does not reverse on a cool-down. A
single-unit, single-lane, recoverable failure points at a local bias table,
heater, or alignment. A multi-unit, multi-lane, recoverable failure points at a
system thermal design or a shared calibration method. Scope cuts the candidate
list before the first instrument is opened.

\paragraph{Step 2: apply the power-versus-signal-quality fork.}
Average power held by the monitor or an external meter means the APC
loop\termnote{APC} is still hitting its setpoint. The candidates that remain
are the ones that close the eye without moving average power: extinction
ratio\termnote{ER} collapse, modulator bias parked on the wrong part of its
curve, wavelength walking off a filter or ring passband, a TEC\termnote{TEC}
or heater running out of range, and receiver noise rising with temperature.
Do not chase laser threshold or fiber loss first. Those would have moved the
power meter.

\paragraph{Step 3: measure at the failing temperature.}
Take the unit to the temperature where it fails and stay there. On a
DCA\termnote{DCA}, read the eye for ER, OMA\termnote{OMA}, and
TDECQ\termnote{TDECQ}. Sweep the EAM\termnote{EAM} or MZM\termnote{MZM} bias
and ask whether the optimum moved away from the stored table value. Read
wavelength on an OSA\termnote{OSA} and compare it with the filter or ring
passband. Read TEC current and heater DAC codes for actuator headroom: an
actuator near its rail is consumed margin even when performance still
passes. If a bias sweep restores a clean eye, the laser is healthy and the
calibration table is wrong. If the eye stays closed across the full bias
range, look at wavelength lock, thermal crosstalk, or the receiver.

\paragraph{Step 4: fix the root cause and close the escape.}
A collapsed eye that a bias sweep restores is usually a temperature-segment
boundary in the calibration table, or a table that was taken at station
ambient and never verified at the loaded high-temperature corner. Fix the
table or the thermal design, then add that loaded corner to the ATP that
missed it. If the actuator is railed, the fix is thermal design or a wider
tuning range, not a tighter BER limit. Recurrence control is the new test
condition, not a hope that the next lot will behave.

\paragraph{How to say it aloud.}
Scope, fork, measure at the failing temperature, name the hypothesis the
measurement eliminates, then name the control. Offer to walk the bias-sweep
or the calibration-table story if the interviewer wants depth.

\subsection{How do you distinguish laser aging from calibration drift?}
\label{sec:interview-worked-aging-vs-cal}

This question tests whether you separate device physics from control-loop
bookkeeping. Aging changes the device. Calibration drift changes the
operating point applied to a healthy device. The frame is an external
reference plus a recovery test
(\Cref{sec:laser-calibration,sec:laser-thermal-vs-aging}).

\paragraph{Step 1: state the two hypotheses cleanly.}
Laser aging raises threshold and lowers slope efficiency at a fixed junction
temperature. The APC\termnote{APC} loop then raises bias to hold power, and
eventually the eye or the RIN\termnote{RIN} budget fails even though average
power still looks fine. Calibration drift leaves the LIV\termnote{LIV} curve
unchanged but parks the product at the wrong bias, wrong EAM\termnote{EAM}
voltage, wrong MZM\termnote{MZM} quadrature, or wrong ring heater code. The
symptoms can look identical in telemetry. The separation requires an external
reference the product does not control.

\paragraph{Step 2: remeasure the device against ship data.}
Bring the unit to a controlled temperature and rerun LIV. Compare threshold
and slope with the shipping record at the same junction temperature. A
threshold rise or slope drop is device aging. If LIV matches ship data, the
device is still healthy and the problem is in the operating point or the
monitor path. Next compare the monitor photodiode\termnote{PD} current
against an external power meter at the same launch. A monitor that disagrees
with the external meter will drive the APC loop to the wrong place silently.
Then sweep modulator bias and ask whether the optimum moved away from the
stored table. A healthy LIV with a moved optimum is calibration drift.

\paragraph{Step 3: use time behavior as a second separator.}
Aging accumulates monotonically across stress intervals. Plot threshold or
bias against cumulative stress hours and look for a steady climb. Calibration
error appears as a step after a thermal excursion, a repair, a rework, a
firmware flash, or a table rewrite, and it is fully corrected by
recalibration. A unit that returns to ship performance after a table reload
was never aged. A unit whose LIV still fails after a fresh calibration is
aged, and recalibration only masks the wear until the next corner.

\paragraph{Step 4: choose the corrective action by the mechanism.}
Aging routes to life modeling, derating, burn-in, or field replacement.
Calibration drift routes to table version control, temperature-segment
verification, monitor-PD integrity checks, and an ATP\termnote{ATP} corner
that exercises the table under load. Mixing the two owners wastes weeks: a
reliability team cannot fix a wrong table, and a firmware team cannot fix a
worn facet.

\paragraph{How to say it aloud.}
``Aging changes the device; calibration drift changes the setpoint. I
separate them with an external LIV and a recovery test, then use time
behavior to confirm.'' Offer the monitor-PD corruption story as the silent
failure mode.

\subsection{How would you qualify a second laser or photonic-integrated-circuit
supplier?}
\label{sec:interview-worked-second-source}

This question tests supplier judgment, not vendor names. The frame: the first
supplier's failure distribution does not transfer. Qualify against the
requirements slice, not against the incumbent's datasheet
(\Cref{tab:laser-prd,sec:supplier-exec}).

\paragraph{Step 1: freeze the requirements, not the part number.}
Write what the new part must close: link budget, RIN\termnote{RIN} at the
stated ORL\termnote{ORL}, bias window, wavelength class, thermal class, and
lifetime FIT\termnote{FIT} target. Those are the acceptance criteria. The
incumbent's datasheet is evidence that one process can meet them, not a
template the second source must copy. A second source that matches the
datasheet but fails the link-budget corners is not qualified.

\paragraph{Step 2: characterize distributions, not samples.}
Measure threshold, slope, wavelength, SMSR\termnote{SMSR}, and RIN across
wafers, lots, and temperature. Compare spreads against the incumbent, not
only means. A hero sample from a new supplier proves nothing about the
process. Ask for wafer maps and lot genealogy so edge-of-wafer outliers are
visible before they enter your module line. Record the same reference planes
and fixtures you use on the incumbent, or the comparison is fiction.

\paragraph{Step 3: run the full qualification on this process.}
HTOL\termnote{HTOL} with the supplier's own activation-energy justification
for the named mechanism, temperature cycling, damp heat, and burn-in. Wear-
out mechanisms and infant-mortality rates are process-specific: epi, facet
coat, attach, and hermeticity all differ. Borrowing the incumbent's $E_a$ is
the most common quiet mistake. State the projection's fine print: sample
size, confidence bounds, and evidence that the stress accelerates the field
mechanism without inventing a new one.

\paragraph{Step 4: correlate ATPs and plan the ramp.}
Run the same units on the supplier's line and yours so an ATP\termnote{ATP}
limit means the same thing at both sites. Size guard bands from the combined
repeatability. Ramp with lot traceability, a FAIR\termnote{FAIR}, and an
agreed reaction plan for excursions. Keep fleet RMA\termnote{RMA} codes
split by supplier so field data can falsify the qualification. A second
source that cannot be traced in the fleet is not a second source; it is a
blind risk.

\paragraph{How to say it aloud.}
``Requirements first, distributions second, process-specific qual third, ATP
correlation and split field codes fourth.'' Offer to go deep on HTOL validity
or on what you would put in the reaction plan.

\subsection{A single lane is weak in a multi-lane module. How do you isolate
optical, electrical, thermal, and assembly causes?}
\label{sec:interview-worked-lane}

Multi-lane modules give you a free control group: the sibling lanes. The
frame is pattern recognition across lanes before any single-lane deep dive
(\Cref{sec:fm-lane-imbalance}).

\paragraph{Step 1: confirm one outlier, not a gradient.}
Measure every lane for OMA\termnote{OMA}, ER\termnote{ER},
TDECQ\termnote{TDECQ}, and wavelength at the same temperature and pattern. A
single-lane cliff is a local defect. A smooth edge-to-center gradient is
thermal or FAU\termnote{FAU} alignment. A checkerboard pattern often points
at electrical routing or a shared supply. Write the pattern down before
changing anything. The siblings tell you what ``normal'' is for this unit.

\paragraph{Step 2: bisect optical versus electrical.}
Swap the electrical lane assignment, or drive the suspect optical path with
a known-good driver channel. If the weakness follows the optical path, the
candidates are the laser or EML\termnote{EML} element, the modulator, the
attach, or the fiber. If it follows the electrical channel, the candidates
are the driver, the TIA\termnote{TIA}, the package routing, or the host
SerDes\termnote{SerDes} lane. This one swap often cuts the tree in half.

\paragraph{Step 3: separate assembly from device and thermal.}
Per-lane coupled power with a normal LIV\termnote{LIV} on the weak lane
points at fiber-array alignment or a contaminated ferrule: the device is
fine, the light is not leaving. A thermal map or per-lane temperature
monitors separate an edge-lane thermal gradient from a device defect. If the
imbalance grows over HTOL\termnote{HTOL} or burn-in time, one array element
is aging faster, which is a lot or screening question, not a design question.

\paragraph{Step 4: pick the corrective action from the pattern.}
Across lanes, units, and lots, the pattern chooses the fix: a rework
instruction for alignment, a coupling-spec change for the FAU, a thermal-
design change for edge lanes, a driver or TIA lot screen, or a supplier
corrective action on one array element. Do not rewrite the whole module
spec for a single-lane assembly escape.

\paragraph{How to say it aloud.}
``Siblings first, then optical-versus-electrical swap, then assembly versus
device versus thermal, then the pattern picks the control.'' Offer the
FAU-alignment story as the most common production escape.

\subsection{Received power is unchanged but required receiver power increased.
What hypotheses remain?}
\label{sec:interview-worked-sensitivity}

This is the other face of the power-versus-signal-quality fork. Average
received power held means the loss is not in the link-budget power ledger.
The frame: eye quality or the receiver itself
(\Cref{sec:fm-ber-increase}).

\paragraph{Step 1: restate what the observation rules out.}
Unchanged received power rules out average launch drop, connector loss, and
fiber attenuation as the primary cause. What remains is signal-quality
erosion on the transmitter or channel, or a receiver whose sensitivity got
worse. Say that out loud so the interviewer hears the fork.

\paragraph{Step 2: list the remaining hypotheses by location.}
Transmitter: eye closure at constant average power from modulator bias
drift, falling ER, rising RIN under reflection, or increased jitter and
chirp. Channel: a new reflection raising MPI\termnote{MPI}, or wavelength
moving relative to a filter edge, both of which degrade quality without
moving the power meter. Receiver: TIA noise, photodiode responsivity, or
equalizer misadaptation. Keep the list short and location-tagged.

\paragraph{Step 3: isolate with golden swaps and a waterfall.}
Swap a known-good transmitter into the link, then a known-good receiver.
Whichever side restores the old sensitivity owns the fault. Sweep BER versus
received power with a VOA\termnote{VOA}. A parallel shift of the waterfall
means the sensitivity moved. A floor that will not go down with more power
points at multiplicative noise: RIN, MPI, or crosstalk. Finish with a
stressed-eye check\termnote{SECQ} of the receiver against a known
transmitter if the golden swap indicts the receive path.

\paragraph{Step 4: close with the mechanism, not the symptom.}
A shifted waterfall with a clean golden transmitter is a receiver problem:
noise, responsivity, or adaptation. A floor is a noise problem; chase ORL,
supply PSRR\termnote{PSRR}, and aggressor lanes before replacing a laser. The
recurrence control depends on the mechanism: a calibration table, a
connector hygiene rule, a supply filter, or a receiver lot screen.

\paragraph{How to say it aloud.}
``Power ledger intact, so quality or receiver. Golden swap, then waterfall
shape, then mechanism.'' Offer to draw the shifted-versus-floored waterfall
on the whiteboard.

\subsection{Why can a link show a BER floor that more launch power does not
fix?}
\label{sec:interview-worked-ber-floor}

This question has a short physics answer and a longer debug answer. Give both
(\Cref{sec:fm-ber-floor,sec:rin}).

\paragraph{Step 1: state the physics in one sentence.}
A BER floor means the dominant noise scales with the signal. Raising launch
power raises that noise in proportion, so the signal-to-noise ratio
saturates. For relative intensity noise the ceiling is
$Q_{\max}=1/\sqrt{\mathrm{RIN}\cdot\mathrm{BW}}$. More power cannot buy you
past that ceiling. Additive noise (thermal, dark current) does not make a
floor; multiplicative noise does.

\paragraph{Step 2: name the usual multiplicative sources.}
Laser RIN, often feedback-driven through poor ORL\termnote{ORL}. Multipath
interference from a pair of reflective interfaces. Bias-rail noise converting
to intensity noise through the laser when PSRR is weak. Crosstalk that
scales with the aggressor. Each of these raises the error rate in a way that
looks like ``the link just will not clean up,'' and each is fixed by
removing the noise source, not by raising OMA.

\paragraph{Step 3: diagnose by waterfall shape and bisect.}
Sweep received power and plot the BER waterfall. Confirm it floors rather
than shifts. Then bisect: measure RIN with a quiet lab source versus the
product bias board to separate intrinsic laser RIN from board-induced
noise. Sweep ORL with a controlled reflector and watch the floor rise and
fall. Read the FEC\termnote{FEC} error histogram: MPI and burst noise cluster
errors in time, while Gaussian noise spreads them. That histogram often
picks the mechanism before a scope does.

\paragraph{Step 4: fix the mechanism, then prevent recurrence.}
Isolator or connector hygiene for feedback. Supply filtering and PSRR
improvement for electrical noise. Aggressor silencing or better isolation for
crosstalk. Replace the laser only if intrinsic RIN is truly out of spec after
the board and ORL have been cleared. Add the ORL or supply stress to the
qualification and to the ATP if that path was never tested.

\paragraph{How to say it aloud.}
``Floor means multiplicative noise; more power cannot help. I confirm with a
waterfall, then bisect RIN, ORL, and the FEC histogram.'' Offer the $Q_{\max}$
line if they want the equation.

\subsection{What makes an HTOL projection credible?}
\label{sec:interview-worked-htol}

A FIT\termnote{FIT} number without a mechanism is a spreadsheet. The frame
is four conditions, said in order (\Cref{sec:laser-aging}).
HTOL\termnote{HTOL} is only as credible as those four.

\paragraph{Condition 1: a named failure mechanism.}
The projection applies to a specific drift parameter: threshold current,
slope efficiency, wavelength, or a defined hard failure such as
COD\termnote{COD}. ``The laser'' is not a mechanism. Name the parameter, the
end-of-life criterion, and the distribution you are projecting.

\paragraph{Condition 2: a justified activation energy.}
$E_a$ must be measured or justified for that mechanism on that process, with
sample size and confidence bounds. Copying an $E_a$ from another product,
another vendor, or a textbook table is the most common way a projection lies.
State the Arrhenius\termnote{Arrhenius} form and the temperatures used, and
be ready to say what happens if $E_a$ is wrong by 0.1~eV.

\paragraph{Condition 3: the stress matches the field mechanism.}
The stress must accelerate the mechanism the fleet will see, without adding a
new one. A stress temperature that triggers solder creep the field never sees
produces a pessimistic fiction. A bench HTOL that omits thermal cycling,
connector wear, or bias-rail transients produces an optimistic fiction. Say
which field mechanisms the HTOL does and does not cover.

\paragraph{Condition 4: a field feedback loop.}
Compare the RMA\termnote{RMA} Pareto against the projection. Divergence is
evidence the model is wrong, not that the fleet is unlucky. Update $E_a$, the
mechanism list, or the use condition when the field disagrees. A projection
that never meets field data is theater.

\paragraph{How to say it aloud.}
``Named mechanism, justified $E_a$, stress matches field, field feedback.
Without those four, FIT is a spreadsheet.'' Offer one example of a stress that
invents a false mechanism.

\subsection{Which data must an automated test save so a failure can be
replayed?}
\label{sec:interview-worked-test-data}

This question is about whether you have ever had to defend a ship decision
months later. The frame: enough data that a unit failing today can be re-
analyzed without the station or the person.

\paragraph{What identity and setup must be saved.}
Sample identity down to serial, lot, and date code. Instrument identity and
calibration state, including the path-loss table version. Fixture, cable, and
reference-plane description, because a power number without a plane is not a
measurement. Software, firmware, and calibration-table versions on the device
under test. Test conditions: temperature, supplies, pattern, dwell, and
timestamps that let results correlate with environmental logs.

\paragraph{What measurement content must be saved.}
Raw measured data, not only pass/fail. Eyes, LIV\termnote{LIV} points,
spectra, and BER\termnote{BER} sweeps that can be re-plotted.
Station-to-station golden-unit correlation results, so measurement drift can
be separated from product drift. Without the raw data, a later engineer can
only argue about the limit, not about the unit.

\paragraph{The replay test of sufficiency.}
Given a field return, can you reproduce the exact shipping measurement and
decide whether the unit changed or the measurement did? If the answer is no,
the archive is incomplete. That question is the acceptance criterion for the
data system, not a nice-to-have.

\paragraph{How to say it aloud.}
``Identity, instrument state, reference plane, raw data, and versions. The
test is whether I can replay ship data on a return.'' Offer one war story
about a missing path-loss table if you have one.

\subsection{When would you choose an EML, silicon MZM, or ring modulator?}
\label{sec:interview-worked-modulator}

Choose by constraint, not preference. State the requirement slice first,
then name what each path buys and costs
(\Cref{tab:source-mod-matrix,tab:tx-modulator}).

\paragraph{When an EML wins.}
An EML\termnote{EML} wins for single-wavelength DR\termnote{DR} or
FR\termnote{FR} at 100--200G/lane when cost, maturity, and one-chip
integration dominate. You get low chirp relative to a DML\termnote{DML}, a
mature supply chain, and a bias-versus-temperature calibration that the
industry already knows how to qualify. You pay for an InP\termnote{InP}
process, EAM\termnote{EAM} aging as a second wear mechanism, and a
calibration table that must be verified at temperature corners.

\paragraph{When a silicon MZM wins.}
A silicon MZM\termnote{MZM} wins when the modulator must sit on the
PIC\termnote{PIC} beside other functions and the link wants a broad, flat
passband with no wavelength lock. You get CMOS\termnote{CMOS}-compatible
integration and a wide optical bandwidth. You pay for die area, drive swing,
RF co-design of the traveling-wave electrode, and a bias-control loop that
holds quadrature. Validation must cover driver matching and bias-rail
behavior across temperature.

\paragraph{When a ring wins.}
A ring\termnote{MRM} wins when many wavelengths must fit on one die, in
dense WDM\termnote{WDM} and co-packaged engines. It is small and WDM-native.
You pay for a wavelength-lock loop, heater power, thermal-crosstalk
validation\termnote{thermalCrosstalk}, and a resonance-alignment failure mode
the others do not have. The validation burden is the decision: if you cannot
afford lock-range and crosstalk testing, you cannot afford the ring.

\paragraph{How to say it aloud.}
``Constraint first. EML for mature single-$\lambda$ DR/FR. Silicon MZM for
broadband PIC integration. Ring for dense WDM with the lock and heater cost
accepted.'' Offer the validation-burden line as the closer.

\subsection{How do you decide whether a field issue is performance,
reliability, or manufacturability?}
\label{sec:interview-worked-triage-class}

Wrong bucket, wrong owner, wasted weeks. Classify on the ticket before
failure analysis starts (\Cref{sec:fleet-triage}).

\paragraph{Bucket 1: performance.}
Ask: did the unit ever meet spec under the field condition? If the design or
operating point never closed the budget at that corner, it is performance.
The fix is retune, derate, a thermal redesign, or a honest spec change.
Shipping more of the same design will not help.

\paragraph{Bucket 2: reliability.}
Ask: did it meet spec at ship and then degrade with time or stress? That is
reliability. The fix is a life model, a screen, derating, burn-in, or field
replacement. Telemetry trends (bias rising at constant power, actuator
heading toward rail) usually show this before the hard failure.

\paragraph{Bucket 3: manufacturability.}
Ask: does the failure cluster by lot, date code, site, or process step
rather than by time? That is manufacturability. The fix is containment,
supplier corrective action (8D\termnote{eightD}), and an ATP\termnote{ATP}
or SPC\termnote{SPC} change. A lot-correlated escape is not a physics
mystery; it is a process-control gap.

\paragraph{How telemetry answers all three before a pull.}
Install age, lot correlation, and trend shape (sudden, gradual, or corner-
dependent) usually pick the bucket before a unit is removed. Classify first,
then pull hardware with a plan that matches the bucket. Performance issues
need a design owner. Reliability issues need a life-model owner.
Manufacturability issues need a supplier and process owner.

\paragraph{How to say it aloud.}
``Ever meet spec? Then degrade with time? Or cluster by lot? Those three
questions pick the owner.'' Offer a one-line example of each bucket from
your experience if you have one.

\subsection{What would you put in fleet telemetry, and why?}
\label{sec:interview-worked-telemetry}

Telemetry exists to catch margin erosion early and to triage without pulling
hardware. Log what discriminates hypotheses, not every register on the chip
(\Cref{sec:cmis,sec:fleet-triage}).

\paragraph{Per-lane observables.}
Transmit and receive optical power, laser bias current, and pre-FEC
BER\termnote{BER} with FEC\termnote{FEC} error histograms. Bias rising at
constant power is aging. Power dropping at constant bias is the optical path.
Clustered errors mean bursts (MPI\termnote{MPI}, connector, ESD event)
rather than Gaussian noise. Per-lane data is what makes the sibling-lane
method possible in the field.

\paragraph{Module observables.}
Temperature, supply rails, TEC\termnote{TEC} or heater drive, and lock-loop
error. Actuators near their rails are consumed margin even while performance
still passes. A railed TEC today is a BER fail next summer. Alarm on actuator
headroom, not only on hard optical thresholds. Read state over
CMIS\termnote{CMIS}.

\paragraph{Events and context.}
LOS\termnote{LOS} and LOL\termnote{LOL} history with timestamps, resets, and
firmware versions, preserved through reboots so intermittents leave evidence.
Lot and date code, install age, and rack position, so one query separates
unit, lot, and environment. Without context, every incident looks unique.

\paragraph{Alarm philosophy.}
Alarm on trends and disagreements, such as monitor versus expected power or
bias versus a peer lane, not only on hard thresholds. Hard thresholds catch
dead units. Trends catch dying ones. The point of fleet telemetry is the
dying ones.

\paragraph{How to say it aloud.}
``Per-lane power, bias, and pre-FEC BER; module temperature and actuator
drive; events with context; alarms on trends.'' Justify each item with the
hypothesis it separates, then stop.
\section{Must-know abbreviations}

A fuller glossary follows this appendix. For one-night preparation, know these
without notes. Entries are alphabetical by the leading abbreviation.

\begin{description}
\item[8D / CAPA] Eight-discipline problem solving and corrective and preventive
      action. Structured containment, root cause, correction, and recurrence
      control for supplier or production failures.
\item[APC] Automatic power control. Feedback from a monitor photodiode that
      holds average launch power; a drifting monitor corrupts it silently.
\item[ATP] Acceptance test plan. Production test limits, methods, reference
      planes, and reaction rules.
\item[BER] Bit error ratio. Pre-FEC BER is the validation metric; post-FEC is
      the service metric. Always say which.
\item[BERT] Bit-error-ratio tester. Generates PRBS and counts errors for
      waterfalls, floors, and dwell tests.
\item[CDR] Clock-data recovery. Extracts bit clock from the data stream; LOL
      means unlock even if light is present.
\item[CMIS] Common Management Interface Specification. Module state, controls,
      alarms, and telemetry.
\item[COD] Catastrophic optical damage. Sudden, irreversible laser-facet failure.
\item[COM] Channel operating margin. Statistical electrical-link margin after
      loss, noise, crosstalk, and equalization.
\item[DCA] Digital communication analyzer. Sampling oscilloscope for eyes,
      OMA, ER, RLM, and TDECQ.
\item[DFB] Distributed-feedback laser. Single-mode 1310~nm class source; also
      the gain section in an EML.
\item[DML] Directly modulated laser. Simple and efficient, but chirp limits reach.
\item[DPPM] Defective parts per million. Incoming or outgoing quality rate.
\item[DR / FR] Datacenter reach ($\sim$500~m) and far reach (2~km). IEEE
      single-mode classes at 1310~nm.
\item[DVT / PVT] Design validation test and production validation test.
\item[EAM] Electro-absorption modulator. Voltage-controlled absorption; pairs
      with a DFB in an EML.
\item[ELSFP] External Laser Small Form-Factor Pluggable. Replaceable continuous-
      wave source for co-packaged optics.
\item[EML] Electro-absorption modulated laser. DFB plus EAM on one InP chip;
      dominant 100--200G/lane pluggable transmitter.
\item[ER] Extinction ratio. $P_1/P_0$ in dB. Higher ER widens OMA at fixed
      average power; trades against chirp and swing.
\item[FAIR] First-article inspection report. Re-qualify after tooling, site,
      epi, or firmware change before open volume.
\item[FAU] Fiber array unit. Precision multi-fiber attachment to a photonic
      integrated circuit.
\item[FEC / KP4] Forward error correction; KP4 is Reed--Solomon RS(544,514).
      Pre-FEC BER and error histograms show margin before FEC failure.
\item[FIT] Failures in time. Failures per $10^9$ device-hours.
\item[GR-468] Telcordia optoelectronic qualification framework (HTOL,
      environmental, mechanical).
\item[HTOL] High-temperature operating life. Accelerated stress used with an
      Arrhenius model to project field wear-out; credible only with a named
      mechanism.
\item[LIV] Light--current--voltage curve. Threshold, slope efficiency, kink-free
      range, and thermal rollover versus bias.
\item[LOS / LOL] Loss of signal and loss of lock. LOS points first toward power;
      LOL can occur with adequate power but poor timing or signal quality.
\item[MPI] Multipath interference. Coherent beating from reflective paths;
      often floors BER and clusters FEC errors.
\item[MPO] Multi-fiber push-on connector. Parallel-optics plant (8--32 fibers).
\item[MRM] Microring modulator. Compact, WDM-native silicon modulator; needs
      wavelength lock and heater budget.
\item[MSA] Multi-source agreement. An industry specification for interoperable
      products.
\item[MZM] Mach--Zehnder modulator. Broadband interferometric modulator (Si or
      TFLN); needs quadrature bias control.
\item[NFF / RMA] No fault found and return merchandise authorization. High NFF
      rates often indicate weak triage or intermittent faults.
\item[OMA] Optical modulation amplitude. Outer level swing $P_1-P_0$.
\item[ORL] Optical return loss. Reflected power toward the laser; low ORL raises
      RIN and can seed burst errors.
\item[OSA] Optical spectrum analyzer. Wavelength, SMSR, and side-mode structure.
\item[OSFP / QSFP-DD] High-density pluggable module form factors with different
      mechanical and thermal limits.
\item[PIC / SOI] Photonic integrated circuit on silicon-on-insulator. The common
      silicon-photonics chip and substrate.
\item[PRBS] Pseudo-random binary sequence. Repeatable pattern for eye and BER
      measurements.
\item[PSRR] Power-supply rejection ratio. Weak PSRR can turn electrical rail
      noise into optical intensity noise.
\item[RIN] Relative intensity noise. Laser amplitude noise (dB/Hz); often
      measured under a stated ORL.
\item[RLM] Relative level mismatch. PAM4 level-spacing quality.
\item[SECQ] Stressed eye closure quaternary. Receiver-side margin measured with a
      calibrated stressed optical signal.
\item[SerDes] Serializer/deserializer. Host high-speed I/O; equalization reserve
      is a timing-margin ledger.
\item[SMSR] Side-mode suppression ratio. Power difference between the lasing mode
      and the strongest side mode.
\item[SPC] Statistical process control. Tracks process distributions and trends.
\item[TDECQ] Transmitter and dispersion eye closure quaternary. Headline PAM4
      transmitter-quality metric after a reference receiver and bounded FFE.
\item[TEC] Thermoelectric cooler. A TEC near its current limit has little thermal
      control margin left.
\item[TIA] Transimpedance amplifier. Converts photodiode current to voltage.
\item[VCSEL] Vertical-cavity surface-emitting laser. 850~nm class for multimode
      short-reach links.
\item[VNA] Vector network analyzer. Electrical or electro-optic $S$-parameters
      versus frequency.
\item[VOA] Variable optical attenuator. Calibrated loss for sensitivity and
      BER-waterfall sweeps.
\item[WDM] Wavelength division multiplexing. Multiple wavelengths on one fiber.
\end{description}

\section{One-night study order}

\begin{enumerate}
\item Memorize the answer spine, scope questions, power fork, and four margins.
\item Build and rehearse two real stories.
\item Review the instruments and say what uncertainty each measurement removes.
\item Review calibration, aging versus thermal response, and HTOL validity.
\item Review the abbreviations above.
\item Rehearse the questions without notes.
\item Stop adding new optics. Sleep.
\end{enumerate}

\keyidea{Show how you reduce uncertainty. Start with requirements and scope,
measure at a named reference plane, eliminate hypotheses, verify the root cause
under the failing condition, and add a control that prevents recurrence.}
