\chapter{Failure analysis handbook}
\label{ch:failure-modes}

This appendix is a symptom-first field guide. Start with what the bench, production
line, or fleet reports. List the plausible failure domains, choose the smallest
measurement that can split them, and preserve the failing state until the evidence
has been captured. Each case follows the same path: observe, hypothesize, measure,
confirm, correct, and add a control that prevents recurrence.

\begin{description}
\item[BER increase] Check received OMA, eye quality, RIN, ORL, receiver noise,
      jitter, and crosstalk (\Cref{sec:fm-ber-floor}).
\item[One lane degraded] Split source, modulator or ring, filter or MUX, fiber
      attach, driver, and receiver (\Cref{sec:fm-lane-imbalance}).
\item[Yield drop] Clear the tester, then split lot, site, process step, assembly,
      firmware, and calibration (\Cref{sec:fm-yield-drop}).
\item[Temperature sensitivity] Track power, wavelength, lock error, TEC or heater
      headroom, receiver margin, and package stress (\Cref{sec:fm-temperature}).
\item[Intermittent bursts] Preserve counters before reseating. Check connector
      contamination, ORL, supply noise, and fiber attach
      (\Cref{sec:fm-contamination}).
\end{description}

The cases below provide the detailed measurements and corrective actions. Use the
failure and debug callouts in Chapters 3--9 as shorter versions of the same method.

\section{BER floor}
\label{sec:fm-ber-floor}

\paragraph{Symptom.}
Pre-FEC BER improves as you increase transmit power, then stops improving and
flattens at a constant floor regardless of how much more power you add.

\paragraph{Physics.}
A BER floor means a noise source that grows with signal power dominates the link.
The classic cause is \emph{relative intensity noise} (RIN): because
$\sigma_\mathrm{RIN} \propto I$, once RIN dominates the noise budget, signal and
noise grow together and $Q$ saturates at
$Q_\mathrm{max} = 1/\sqrt{\mathrm{RIN}_\mathrm{lin} \cdot \mathrm{BW}}$
(\Cref{sec:rin}). Other causes: multipath interference (MPI) from dirty
connectors or high back-reflection, or broadband noise on the laser bias rail
that converts to equivalent RIN (\Cref{sec:laser-drivers}).

\paragraph{Debug steps.}
\begin{enumerate}
\item Confirm the floor exists: sweep received power (or Tx OMA) and plot BER vs.\
      power. A healthy link has a steep waterfall; a floor appears as a horizontal
      asymptote.
\item \textbf{Bisect optical vs.\ electrical RIN.} Measure RIN with a quiet SMU
      powering the laser (intrinsic RIN). Then repeat with the product bias board
      connected. If the floor moves, the electrical path is injecting noise
      (\Cref{sec:laser-drivers}).
\item \textbf{Sweep ORL.} Add a controlled reflector. If BER floor worsens with
      lower ORL, the laser is feedback-sensitive. Check isolator, connector, and
      fiber-attach cleanliness.
\item \textbf{Check for MPI.} Look at the ESA (electrical spectrum analyzer) for
      discrete spurs at frequencies corresponding to round-trip delays of reflective
      interfaces. MPI creates bursty errors; the FEC histogram shows clustered
      symbol errors rather than random ones.
\item \textbf{Confirm RIN spec.} Compare measured $\mathrm{RIN}_x\mathrm{OMA}$ at
      the stated ORL against the PMD or ATP limit (e.g.\ $-136$~dB/Hz at 17.1~dB
      ORL for DR-class links). If the part exceeds spec, reject.
\end{enumerate}

\paragraph{Resolution.}
If intrinsic RIN is the limiter, the laser is marginal or aged; replace or derate.
If electrical RIN, fix the bias supply (better PSRR, star ground, local
decoupling). If ORL-driven, clean or replace connectors and verify isolator
function. If MPI from multiple reflections, reduce the number of mated interfaces
or improve their ORL.

\section{Low extinction ratio}
\label{sec:fm-low-er}

\paragraph{Symptom.}
Transmitter OMA looks low on the DCA even though average power is in range. TDECQ
may or may not fail depending on how the reference equalizer compensates.

\paragraph{Physics.}
Extinction ratio $\mathrm{ER} = P_1/P_0$ sets how far apart the one and zero
optical levels are for a given average power. Low ER means the zero level is too
high (the modulator does not fully extinguish) or the one level is too low. In an
EML, ER is set by the EAM reverse bias: insufficient bias leaves residual
transmission in the off-state. In a DML, ER depends on modulation depth relative to
threshold. The OMA penalty for finite ER is
$\mathrm{PP} = (\mathrm{ER}+1)/(\mathrm{ER}-1)$: at 10~dB ER the penalty is
$\sim$0.87~dB; at 6~dB ER it rises to $\sim$2.2~dB (\Cref{sec:sensitivity}).

\paragraph{Debug steps.}
\begin{enumerate}
\item Measure ER on the DCA (outer OMA / average power, or directly from the
      histogram levels). Compare against the PMD limit.
\item \textbf{EML:} Sweep EAM bias. ER should peak at the optimal bias point; if
      the curve has shifted (aged EAM), the operating point needs recalibration.
      Check EAM bias DAC code vs.\ datasheet.
\item \textbf{DML:} Check bias current vs.\ LIV. If bias is close to threshold,
      modulation depth is limited. Increase bias (but watch thermal rollover and
      RIN).
\item \textbf{MZM:} Check quadrature bias. If the MZM has drifted off quadrature
      ($V_\pi/2$ point), extinction degrades. Log the bias-control loop error
      signal; a saturated loop indicates drift beyond correction range.
\item \textbf{Ring:} Check resonance alignment. If the ring is detuned from the
      laser wavelength, extinction drops. Monitor the thermal tuner current and
      wavelength-lock error.
\end{enumerate}

\paragraph{Resolution.}
Recalibrate the modulator operating point. For EML aging, update the EAM bias
setpoint in firmware or flag the module for replacement if the absorption curve has
shifted beyond the correctable range. For MZM drift, verify the bias controller and
its monitor PD. For rings, retune or check for neighbor thermal crosstalk
(\Cref{sec:thermal-xtalk}).

\section{Lane imbalance}
\label{sec:fm-lane-imbalance}

\paragraph{Symptom.}
In a multi-lane module (DR4, FR4, DR8), one or more lanes show significantly
different OMA, TDECQ, or pre-FEC BER compared with siblings in the same module.
The weak lane may be marginal or failing while others are healthy.

\paragraph{Physics.}
Multi-lane modules share a substrate, laser array (or CW-WDM source), and thermal
environment. Lane-to-lane variation comes from: (1)~die-level non-uniformity in the
laser or modulator array (threshold, slope, $V_\pi$, coupling), (2)~packaging
variation in fiber-array alignment (one channel of the FAU slightly misaligned),
(3)~driver or TIA channel mismatch on the electronic IC, or (4)~thermal gradient
across the die (edge lanes hotter or cooler than center lanes).

\paragraph{Debug steps.}
\begin{enumerate}
\item Measure all lanes: OMA, ER, TDECQ, RIN, wavelength. Identify the outlier.
\item \textbf{Is it optical or electrical?} Swap the electrical lane assignment
      (if the host supports lane remapping) or test the suspect optical lane with a
      known-good driver channel. If the problem follows the optical path, it is
      laser/modulator/fiber-attach. If it follows the electrical channel, it is
      driver/TIA.
\item \textbf{Check coupling.} Measure per-lane fiber-coupled power with an
      integrating sphere or power meter array. A single weak lane with normal LIV
      suggests FAU misalignment.
\item \textbf{Thermal map.} Use an IR camera or CMIS per-lane monitors (if
      available) to check for hot spots. Edge lanes near the package wall or near a
      TEC boundary may run hotter.
\item \textbf{Laser array aging.} If the imbalance grows over time (HTOL or field
      life), one laser in the array is aging faster (threshold rise, slope drop).
      Compare LIV curves at $t_0$ and now.
\end{enumerate}

\paragraph{Resolution.}
FAU misalignment is a manufacturing escape; tighten incoming inspection or
first-article coupling specs. Thermal gradient: redesign TEC zoning or derate the
hot lane. Laser aging: flag the lot and check burn-in screening effectiveness.
Driver mismatch: work with the IC supplier on channel-to-channel gain flatness.

\section{Wavelength drift}
\label{sec:fm-wavelength-drift}

\paragraph{Symptom.}
In a WDM system, one or more channels walk off the ITU grid or the ring/filter
passband. BER degrades as the channel moves off the receiver filter or MUX/DEMUX
passband. In ring-modulator CPO, this manifests as sudden unlock of the wavelength
control loop.

\paragraph{Physics.}
Laser wavelength drifts with temperature ($d\lambda/dT \approx 0.1$~nm/$^\circ$C
for InP DFB) and bias current ($d\lambda/dI \approx 0.01$~nm/mA). If the TEC or
wavelength-locker servo cannot track, the channel walks off its assigned slot.
In microring systems, the ring resonance drifts at $\sim$80~pm/$^\circ$C
($\sim$10~GHz/$^\circ$C in Si), and neighbor heaters create thermal crosstalk that
pushes adjacent channels (\Cref{sec:thermal-xtalk}).

\paragraph{Debug steps.}
\begin{enumerate}
\item Measure wavelength on an OSA or wavemeter. Compare to the target grid.
\item \textbf{Laser-side:} Check TEC current. If saturated (at max or min drive),
      the thermal load exceeds TEC capacity. Check case temperature and airflow.
\item \textbf{Locker-side:} Read the wavelength-locker error signal. A healthy loop
      holds near zero; drift means the servo is losing lock. Check locker etalon
      alignment and PD balance.
\item \textbf{Ring CPO:} Check the ring thermal tuner DAC code. If it has railed
      (max heater power), the ring cannot reach the target wavelength. Check for
      neighbor heating (all adjacent lanes at full traffic and max case $T$).
\item \textbf{Aging:} If drift is progressive over weeks/months, suspect laser mode
      hop or gradual TEC degradation (reduced $\Delta T$ capacity).
\end{enumerate}

\paragraph{Resolution.}
TEC saturation: reduce case temperature (improve airflow or liquid cooling) or
derate the laser operating current. Ring unlock: increase heater headroom in the
design, reduce thermal crosstalk with layout changes, or shift the CW-WDM source
grid to re-center the ring tuning range. Aging: schedule preventive replacement
(ELSFP hot-swap, \Cref{sec:elsfp}).

\section{Eye closure (high TDECQ)}
\label{sec:fm-eye-closure}

\paragraph{Symptom.}
TDECQ exceeds the PMD limit even though average power and ER look acceptable. The
DCA eye appears compressed or distorted after the reference equalizer is applied.

\paragraph{Physics.}
TDECQ measures how much noise the transmitter can tolerate before BER exceeds the
FEC threshold, relative to an ideal transmitter (\Cref{sec:tdecq}). High TDECQ
means the equalized eye is poor. Common causes: (1)~insufficient EO bandwidth
(modulator or driver roll-off), (2)~poor level linearity (RLM $<$ 0.95;
driver or modulator compression), (3)~pattern-dependent effects (ISI from
bandwidth limit, reflections, or impedance mismatch), (4)~chromatic dispersion on
FR-class fiber eating into the margin.

\paragraph{Debug steps.}
\begin{enumerate}
\item Inspect the raw (unequalized) eye on the DCA. Is it bandwidth-limited
      (rounded transitions), compressed (uneven levels), or noisy (RIN/jitter)?
\item \textbf{Bandwidth:} Measure EO $S_{21}$ of the modulator (if accessible) or
      the combined Tx path. Compare to the Nyquist frequency. If 3-dB BW is below
      Nyquist, bandwidth is the limiter.
\item \textbf{RLM:} Compute relative level mismatch from the DCA histogram. If
      $\mathrm{RLM} < 0.95$, the PAM4 levels are unevenly spaced. Check driver
      linearity (swept DAC code) and modulator transfer function (bias sweep).
\item \textbf{Reflections:} Check $S_{11}$ of the RF path (driver to modulator)
      and the optical return loss. Reflections cause post-cursor ISI the FFE
      cannot fully cancel.
\item \textbf{Dispersion:} Measure TDECQ with and without the test fiber (TECQ vs.\
      TDECQ). If TDECQ is significantly worse than TECQ, dispersion is the marginal
      contributor. Check Tx wavelength and chirp against the fiber length and
      dispersion coefficient.
\end{enumerate}

\paragraph{Resolution.}
Bandwidth: upgrade driver or modulator (higher-BW die or peaking network). RLM:
tune driver pre-emphasis (DAC levels for each PAM4 symbol). Reflections: fix
wirebond, impedance discontinuity, or connector. Dispersion: tighten wavelength
tolerance or shorten the fiber (re-route).

\section{Thermal runaway}
\label{sec:fm-thermal-runaway}

\paragraph{Symptom.}
Module temperature rises continuously until shutdown (CMIS over-temperature alarm)
or until the laser degrades catastrophically. May start subtly: BER creeps up as
case $T$ rises during traffic ramp or with neighbor-lane loading.

\paragraph{Physics.}
In a faceplate pluggable, the module dissipates $\sim$8--20~W into the cage
heatsink. If airflow is blocked, the cage is overloaded (too many high-power
modules), or the TEC inside the module is fighting a losing battle against
junction temperature, the thermal loop runs away. In CPO, the optical engine sits
on the switch substrate next to a high-power ASIC ($>$500~W); even small failures
in the cooling path concentrate heat on the laser or ring heaters.

\paragraph{Debug steps.}
\begin{enumerate}
\item Read CMIS module temperature and compare to the module's rated case $T$
      range. If case $T$ exceeds the max, the system cooling is inadequate.
\item \textbf{Check TEC current.} A TEC at max drive current is saturated; it
      cannot pump more heat. The junction temperature is higher than the case $T$
      suggests.
\item \textbf{Measure LIV at temperature.} If threshold rises and slope drops
      steeply with $T$, the laser is near thermal rollover (\Cref{sec:laser-aging}).
      The operating point may be marginal.
\item \textbf{Neighbor loading.} Bring all lanes and neighbor modules to full
      traffic simultaneously. If the problem only appears under full-cage load, the
      thermal design margin is insufficient.
\item \textbf{Airflow audit.} Inspect the faceplate for blocked vents, incorrect
      fan speed, or missing blanking panels (bypass airflow). For liquid-cooled
      systems (CPO, XPO), check coolant flow rate and inlet temperature.
\end{enumerate}

\paragraph{Resolution.}
System-level: improve airflow, lower ambient, or reduce module count per cage.
Module-level: derate the laser (lower bias current reduces self-heating) or switch
to a lower-power module style (LPO instead of retimed, \Cref{sec:conditioning}).
CPO: ensure the cold-plate thermal interface material (TIM) is intact and the
liquid loop meets flow-rate spec. Long-term: specify a tighter thermal class in the
laser requirements (\Cref{sec:laser-reqs}).

\section{Connector contamination}
\label{sec:fm-contamination}

\paragraph{Symptom.}
Intermittent link failures, burst errors, or elevated BER that clears after
reseating or cleaning a connector. Rx power may fluctuate or show sudden drops.
Often affects one direction of a duplex link.

\paragraph{Physics.}
A particle of dust on an MT, LC, or MPO ferrule endface scatters and absorbs
light, raising insertion loss and back-reflection (lowering ORL). Particles as
small as 1~$\mu$m on a single-mode core (9~$\mu$m diameter) can cause $>$1~dB
loss. Elevated ORL feeds back into the laser and raises RIN, causing burst errors
even when average power looks acceptable. In high-power CW-WDM and ELSFP systems,
trapped particles can burn onto the fiber endface and cause permanent damage.

\paragraph{Debug steps.}
\begin{enumerate}
\item \textbf{Inspect first.} Use a fiber-endface inspection scope (200--400$\times$)
      on every mated connector before any other debug. Look for particles in the
      core zone, scratches, pits, and residue.
\item \textbf{Clean and re-inspect.} Use a dry-click cleaner or lint-free wipe with
      IPA. Re-inspect. If the endface still fails IEC 61300-3-35 zone criteria,
      replace the jumper.
\item \textbf{Measure IL and ORL.} After cleaning, measure insertion loss and ORL
      across the mated pair. Compare to the link-budget allocation
      (\Cref{sec:link-budget}).
\item \textbf{Correlate with BER.} If BER clears after cleaning, the contamination
      was the root cause. Log the connector location and date code for trend
      analysis.
\item \textbf{Repeat offenders.} If the same location re-contaminates quickly, check
      for airborne particulates (raised-floor debris, construction dust) or
      improper cap/cover discipline during service events.
\end{enumerate}

\paragraph{Resolution.}
Immediate: clean and verify. Preventive: install dust caps on every unused port,
enforce ``inspect before connect'' policy in the service runbook, use sealed
cassettes or connectorized trunk cables that minimize open-ferrule exposure. For
high-power paths (ELSFP, CW-WDM), any connector that shows burn damage must be
replaced, not re-cleaned. Track contamination-related RMAs as a distinct failure
code (not ``laser failure'') so FIT accounting stays honest
(\Cref{sec:fleet-triage}).

\section{Yield drop}
\label{sec:fm-yield-drop}

\paragraph{Symptom.}
First-pass yield falls below its stable baseline. The loss may cluster on one ATP
row, lane, tester, shift, supplier lot, assembly site, or firmware revision.

\paragraph{Likely causes.}
Process drift, incoming material variation, fiber-array alignment, die-attach or
solder change, tester or fixture drift, stale calibration, a software limit change,
or a guardband that no longer matches measurement spread.

\paragraph{Debug steps.}
\begin{enumerate}
\item Contain suspect work in process and freeze tester software, limits, fixtures,
      and calibration records.
\item Build a Pareto by ATP row, lot, date code, site, tester, and shift. Do not
      average away a one-lane or one-station signature.
\item Run a golden unit across stations and the same failed unit on the reference
      bench. This clears the measurement system before supplier action begins.
\item Compare upstream wafer, die, assembly, and incoming data at the first point
      where good and bad populations separate.
\item Select failure analysis that distinguishes the remaining mechanisms, then
      verify the correction on a controlled lot and watch the field code.
\end{enumerate}

\paragraph{Corrective action.}
Restore the changed process or measurement input, add a statistical control at the
first observable point, revise ATP only with correlation data, and require a new
first-article check before releasing volume (\Cref{sec:supplier-exec}).

\section{Temperature sensitivity}
\label{sec:fm-temperature}

\paragraph{Symptom.}
BER, TDECQ, optical power, or lock stability degrades during a case-temperature
ramp or under loaded-neighbor conditions. The unit may recover when cooled.

\paragraph{Likely causes.}
Laser threshold and slope drift, wavelength movement, ring-resonance drift, TEC or
heater saturation, EAM or MZM bias error, receiver-noise rise, package stress, or a
thermal gradient that affects one lane.

\paragraph{Debug steps.}
\begin{enumerate}
\item Record case temperature, external optical power, wavelength, bias, TEC or
      heater current, lock error, OMA, ER, TDECQ, and pre-FEC BER on one time axis.
\item Repeat with neighbors off and on. A shared shift points toward thermal or
      supply coupling; a lone lane points toward its local path.
\item Hold the source fixed and move the wavelength-selective element, then reverse
      the test. This splits laser drift from ring or filter drift.
\item Rerun LIV and receiver sensitivity hot and cold. Separate lost launch power
      from lost receiver margin.
\item Inspect package and cooling interfaces if optical and electrical blocks pass
      alone but fail in the assembled system.
\end{enumerate}

\paragraph{Corrective action.}
Restore thermal headroom, correct calibration and control limits, reduce coupling,
or derate the operating point. Add the loaded-neighbor temperature ramp to the ATP
or design-validation plan that missed it.

\keyidea{A useful failure analysis starts with a symptom and ends with a new
control. Preserve the failing state, split shared from local behavior, clear the
measurement system, and choose one measurement that can disprove the leading
hypothesis. The corrective action is incomplete until production or fleet data
show that the same signature no longer escapes.}
