\chapter{WDM and wavelength-locked lasers}
\label{ch:wdm}

Early datacenter optics mostly ran one wavelength per fiber. That worked while
port counts were modest. At AI scale, fiber count itself becomes a first-order
cost and cable-plant problem, so the industry packed more channels onto each
strand. The price of that packing is control: once channel spacing tightens, or
once the modulator is a wavelength-selective ring, someone must keep laser and
filter locked together. Few phrases carry as much architectural information as
``wavelength-locked laser,'' because locking only appears under those
interconnect choices. Ring and MZM device physics stay in
\Cref{sec:siring,sec:simzm}; per-$\lambda$ laser ATP and aging stay in
\Cref{ch:lasers}. This chapter covers grids, lock loops, thermal crosstalk, MUX
budget, and CW-WDM architecture.

\section{Why multiplex wavelengths at all}
\label{sec:why-wdm}

At $100{,}000$+ accelerator scale, every extra fiber is another connector,
another patch, and another failure mode. \term{Wavelength-division multiplexing}
(WDM)\firstterm{WDM} puts many independent channels on a single fiber, each on its own
wavelength, so bandwidth per fiber rises without adding fiber. Each wavelength
can still be an ordinary IM/DD channel. WDM and IM/DD are orthogonal; you simply
run IM/DD \emph{per wavelength}.

Historically the industry climbed a ladder of spacing. Coarse CWDM4 used
$\approx$20~nm slots and uncooled lasers. LAN-WDM tightened that for 2~km-class
FR4. Dense grids and then CW-WDM O-band combs for CPO pushed spacing into the
100--800~GHz class and made active locking mandatory. Those spacings are
standardized grids, not vendor choices: the 20~nm CWDM slots follow the ITU-T
G.694.2 wavelength grid (18 channels, 1271--1611~nm), and the 50/100/200~GHz
datacom DWDM spacings follow the ITU-T G.694.1 frequency grid anchored at
193.1~THz~\firstcite{ituG694}. CWDM4 uses the four O-band lines of that CWDM
grid; the CW-WDM combs in \Cref{sec:cwwdm} define their own O-band grids for
dense integration. \Cref{tab:wdm-grids} is that ladder as you will meet it in
short-reach AI optics today.

\begin{table*}[ht]
\refstepcounter{table}\label{tab:wdm-grids}%
\centering
\setlength{\tabcolsep}{3.5pt}%
\footnotesize
\begin{tabular}{@{}p{0.16\linewidth}p{0.14\linewidth}p{0.12\linewidth}p{0.18\linewidth}p{0.32\linewidth}@{}}
\toprule
Grid family & Spacing (class) & Channels / fiber & Cooling / lock & Typical short-reach use \\
\midrule
CWDM4 & $\approx$20~nm & 4 & Uncooled; loose control & FR-class pluggables; faceplate WDM \\
LAN-WDM & $\approx$800~GHz ($\approx$4--5~nm @ 1310~nm) & 4 & Cooled or tight open-loop & 2~km-class FR4 (edge of book scope) \\
Datacom DWDM & 200/100/50~GHz & many & Locked to grid & Discrete DFB/EML DWDM modules \\
CW-WDM / CPO O-band & 100--800~GHz class (MSA spans) & 8 / 16 / 32 & Locked; often ring-tuned & CPO engines, optical I/O chiplets \\
\bottomrule
\end{tabular}
\par\medskip
{\raggedright\footnotesize\textbf{Table~\thetable.} WDM grids for short-reach AI interconnects.
CW-WDM MSA normative grids sit in O-band with 9/18/36~nm spans and 8/16/32-line
sets (\Cref{sec:cwwdm}); spacing is set by the chosen span and channel count, not
by Ethernet CWDM4.\par}
\end{table*}

\section{Why ``locked'' is the operative word}
\label{sec:why-locked}

WDM alone does not force active locking. CWDM4 packs four wavelengths with enough
spacing that uncooled lasers can wander and still stay in their slots. Locking
becomes the operative word only when either the channel spacing is tight or the
modulator itself is wavelength-selective. Those two situations drive nearly every
modern CPO and dense optical-I/O control loop.

\subsection{Tight DWDM grids}

To pack channels at 50--200~GHz class spacing\firstterm{DWDM}, each laser must sit on its grid
slot or adjacent channels collide. Emphasizing ``locking'' therefore hints at
spacing tighter than CWDM. You do not stress locking for coarse, uncooled CWDM4;
you do as soon as the grid looks like LAN-WDM, datacom DWDM, or a CW-WDM comb.

\subsection{Microring modulators and WDM locking}

Resonant ring and microdisk modulators are the dense-WDM workhorse
(\Cref{sec:siring}). Their resonance drifts by roughly $10$~GHz/\textdegree C in
silicon, so even a stable laser is not enough: the laser and the ring must be
wavelength-locked. Either lock the laser to the ring, or thermally tune the ring
onto the laser with a feedback loop. That laser--ring co-locking is the central
control problem in ring-based WDM links and in co-packaged optics, and it is why
neighbor heat and case-temperature ramps belong in validation, not only in the
thermal section of a datasheet.
\section{WDM filters, grids, and on-chip multiplexing}
\label{sec:wdm-hardware}

WDM is not only lasers and locking (\Cref{sec:cwwdm}): the PIC\firstterm{PIC} needs wavelength
selective routing. The MUX/demux stage is a first-class link-budget line item
(\Cref{sec:link-budget}), not a packaging footnote.

\paragraph{Hardware choices.}
\begin{description}
\item[AWG / echelle gratings] multiplex and demultiplex on silicon or glass.
      Insertion loss is often 2--5~dB per MUX stage; adjacent-channel crosstalk and
      passband ripple land in OMA and transmitter and dispersion eye closure
      quaternary (TDECQ).
\item[Ring filter banks] drop/port routing in microring banks sets how many
      $\lambda$ share a bus waveguide. Thermal tuning per ring is common
      (\Cref{sec:siring,sec:thermal-xtalk}).
\item[Hybrid] some engines use a coarse AWG plus fine ring filters; count every
      stage in the ledger.
\end{description}

\newthought{MZMs trade area for calm wavelength behavior.} When each lane carries
its own laser (DR/FR SiPh modules), silicon Mach--Zehnder modulators sidestep ring
locking (\Cref{sec:simzm}). Rings remain the default when many $\lambda$ share one
PIC and area dominates (\Cref{sec:siring,ch:networking}).

\paragraph{Where MUX defects land.}
\Cref{tab:mux-budget} maps common MUX faults to the measurement that catches them.

\begin{table*}[ht]
\refstepcounter{table}\label{tab:mux-budget}%
\centering
\setlength{\tabcolsep}{3.5pt}%
\footnotesize
\begin{tabular}{@{}p{0.22\linewidth}p{0.28\linewidth}p{0.22\linewidth}p{0.22\linewidth}@{}}
\toprule
Fault & Optical symptom & Hits & Catch with \\
\midrule
Stage insertion loss & Lower launch OMA on all $\lambda$ & Link budget OMA & Power meter / OMA \\
Passband ripple / tilt & Uneven OMA across bank & Weakest $\lambda$ & Per-$\lambda$ OMA map \\
Adjacent-channel crosstalk & Closed eyes, RLM/TDECQ up & Tx quality / BER & Isolation sweep + DCA \\
MUX imbalance & One $\lambda$ weak, neighbors OK & Single-lane BER & Per-lane power + BER \\
Grid misalignment & Filter edge clipping & TDECQ, unlock risk & OSA + lock status \\
\bottomrule
\end{tabular}
\par\medskip
{\raggedright\footnotesize\textbf{Table~\thetable.} MUX/demux defects and where they
appear in validation. Isolation and imbalance tests belong in ATP for any dense
WDM engine (\Cref{sec:lock-validation,sec:cwwdm-laser}).\par}
\end{table*}

Validation adds channel isolation sweeps, grid alignment across temperature, and
MUX imbalance (uneven OMA per $\lambda$). Treat the weakest channel as the
budget-limiting lane, not the average.

\section{Lock-loop mechanics}
\label{sec:locking-techniques}

Wavelength locking closes the loop between source and filter. Pick a technique
based on whether the laser, the ring, or both must be steered
(\Cref{sec:siring,ch:validation}).

\paragraph{Error-signal sources.}
\begin{description}
\item[Etalon-based wavelength locker]\firstterm{wavelengthLocker}
      A fixed reference etalon plus a pair of photodiodes produces an error signal
      proportional to wavelength offset; feedback trims laser temperature or current
      onto the grid. Common on discrete DFB/EML modules.
\item[Laser-to-ring thermal feedback]
      Monitor the ring's through/drop power (or a dither tone) and heat the ring, or
      trim laser current, to park the carrier on resonance. Default for dense
      microring WDM banks in CPO and optical I/O.
\item[Injection / external-cavity locking]
      Stabilize a laser's wavelength and linewidth against an external reference
      cavity; higher performance, more parts. Rare in short-reach volume products.
\item[Athermal design]
      Engineer the device so its resonance barely moves with temperature, reducing
      the control burden. Athermal does not remove MUX grid alignment; it shrinks
      the loop authority you need.
\item[Digital supervisory loop]
      CMIS-exposed monitors and firmware on modern modules; link training at
      224G/448G may iterate EQ and wavelength trim together
      (\Cref{sec:pluggables}).
\end{description}

\paragraph{What you trim.}
Three actuators show up repeatedly, and the bring-up order usually starts with the
slowest, highest-authority knob. Laser TEC / temperature moves the whole comb or
a single DFB on the frequency axis. Laser bias current is the fine wavelength trim
(and also changes power), so watch RIN and SMSR when you use it as a locker
(\Cref{sec:laser-drivers}). Ring heaters park each microring onto its assigned
$\lambda$ and are the per-channel control in dense banks
(\Cref{sec:thermal-xtalk}).

\paragraph{Capture versus hold.}
\emph{Capture} is acquiring lock from a cold or unlocked state: coarse scan of
heater or TEC until the monitor error signal crosses zero, then close the loop.
\emph{Hold} is rejecting thermal drift and neighbor crosstalk once locked. Loop
bandwidth must be fast enough to track case-temperature ramps and adjacent-heater
steps, but slow enough not to fight the data path or inject RIN through bias
modulation. Silicon rings at $\sim$10~GHz/\textdegree C set the disturbance scale:
a 1~\textdegree C neighbor step is tens of GHz of resonance walk, which is a large
fraction of a 100--200~GHz grid slot (\Cref{sec:siring}).

\paragraph{Failure signatures.}
Fleet failures often look like slow wavelength walk: drop monitor power falls,
TDECQ rises on one $\lambda$, or one lane in a WDM bank drops out while neighbors
stay up. Bisect laser versus ring by toggling TEC setpoint and ring heater
independently (\Cref{sec:instruments,sec:fleet-triage,sec:lock-validation}).

\section{Thermal crosstalk and heater budget}
\label{sec:thermal-xtalk}

Dense ring banks share a substrate. Heating one ring to stay on resonance shifts
neighbors. That is \firstterm{thermalCrosstalk}, and it is why a single-lane lock
test at room temperature is not a product test.

\paragraph{Where heat comes from.}
Self-heating from the ring's own heater (and absorbed optical power) shifts its
resonance, so the lock loop must settle with the lane at operating optical power,
not dark. Adjacent heaters on nearest-neighbor and next-nearest rings in a WDM
bank are the next disturbance; the worst case is all neighbors at max heater power
while you hold lock. Package and ASIC load add a common-mode walk: switch or XPU
case-temperature ramps and local hotspots move the whole bank, and a shared TEC or
cold plate sets how much of that walk the lock loop must reject
(\Cref{sec:prod-corners}).

\paragraph{Design and validation implications.}
Budget heater range with headroom for crosstalk, not just for the coldest and
hottest case alone. Layout (heater placement, thermal isolation trenches, ring
pitch) is a reliability and yield problem as much as a control problem. In
validation, simultaneous full-traffic on neighbors plus max case $T$ is a
\emph{lock} test: unlock, BER walk, or TDECQ rise on one $\lambda$ under neighbor
load points at thermal design, not at a bad laser die
(\Cref{sec:prod-corners,sec:fleet-triage}).

\section{External multi-wavelength sources (CW-WDM)}
\label{sec:cwwdm}

Dense WDM with ring modulators needs a source of many clean, stable wavelengths.
The industry answer is a \emph{disaggregated} external laser: a single
multi-wavelength continuous-wave (CW) module supplies a comb of wavelengths over
fiber to the photonic engine, where microrings imprint data onto each one. The
\term{CW-WDM MSA} standardizes those sources for AI, HPC, and high-density
optics~\firstcite{cwwdm}. Source-side measurement detail (per-channel power, SMSR,
RIN, lock under neighbor heat) is in \Cref{sec:cwwdm-laser}; this section is the
architecture contract.

\paragraph{What the MSA specifies (and what it does not).}
Rev~1.0 (June 2021) defines O-band wavelength grids, port configurations, and
measurement methods. It does \emph{not} standardize mechanical form factors,
management pins, or full link parameters; those stay application-specific or move
to form-factor MSAs such as ELSFP~\firstcite{elsfp,cwwdm}.

Core normative content:
\begin{itemize}
\item \textbf{Grid sets:} 8+1 and 16+1 lines in a 9~nm span; 8+1 / 16+1 / 32+1 in
      18~nm and 36~nm spans (shortest line optional in each set).
\item \textbf{Spacings (class):} for the 18~nm span, channel spacing is 400 / 200 /
      100~GHz for 8 / 16 / 32-line sets; the 9~nm and 36~nm spans scale spacing
      with span width (100--800~GHz class). Normative MSA grids are denser than
      5~nm; coarser CWDM-like spacings are informative only.
\item \textbf{Two physical configs:} \emph{modular} (each fiber carries one
      $\lambda$) and \emph{integrated} (each fiber carries the full comb).
\item \textbf{Power classes and AS parameters:} output power classes span low to
      high launch; SMSR, RIN, and linewidth floors are defined with measurement
      methods, with many limits marked application-specific (AS) in the normative
      tables.
\end{itemize}

Informative appendix examples (not universal product guarantees) often quote
$\approx$30~dB SMSR, $\approx-135$~dB/Hz RIN, $\approx$20~MHz linewidth,
$\pm$1~dB per-line power variation, and $-20$~dB ORL tolerance for 18~nm-span
examples.
Treat those as negotiation anchors; write the ATP to your link budget
(\Cref{sec:cwwdm-laser,tab:laser-prd}).

\paragraph{Why disaggregate the laser.}
External CW-WDM / ELSFP sources are field-replaceable. Lasers dominate wear-out
FIT (\Cref{ch:lasers,ch:reliability}); keeping them off the ASIC package turns a
COD or facet failure into a hot-swap, not a board pull. The photonic engine still
needs per-$\lambda$ lock to rings or filters.

\paragraph{Exemplar: SuperNova + TeraPHY.}
Ayar Labs' optical-I/O stack is aimed at \emph{scale-up} (XPU-to-XPU) rather than
switch fabric~\firstcite{ayar}:
\begin{description}
\item[TeraPHY] an optical-I/O chiplet co-packaged with the host XPU, carrying the
      microring modulators and receivers.
\item[SuperNova] the external CW light source, positioned as the first
      CW-WDM-MSA-compliant 16-wavelength source, delivering up to 16 wavelengths
      into each of 16 fibers. That is light for 256 data channels (vendor claim:
      about 16~Tb/s bidirectional), and roughly $64\times$ the wavelength count of
      CWDM4 pluggables.
\end{description}

Vendor performance claims versus pluggables plus electrical SerDes (5--10$\times$
bandwidth, 10$\times$ lower latency, 4--8$\times$ better power efficiency) are
marketing numbers; use them as orientation, not as ATP limits.\sidenote{Source:
\citehref{ayar}{Ayar Labs, SuperNova light source, ayarlabs.com}. Positioning:
Ayar targets scale-up optical I/O between accelerators; Broadcom and NVIDIA CPO
(\Cref{sec:cpo-status}) put the same building blocks on the \emph{switch}.}

\paragraph{System requirements the source must meet.}
For the PIC to close every lane, the comb must deliver, across temperature and
with all ports active, a small set of properties at once: per-line power flatness
(else MUX + modulator bank makes uneven OMA), grid placement and SMSR per line,
RIN under the specified ORL, and absolute wavelength stable enough that ring
heaters stay in range
(\Cref{sec:thermal-xtalk,sec:cwwdm-laser}). Miss any one and a single $\lambda$
will look like a modulator or lock-loop failure when the source is the real cause.

\subsection{Comb sources: one device, many lines}
\label{sec:comb-sources}

The SuperNova approach builds its comb from an array of discrete lasers, one
distributed-feedback (DFB) die per wavelength, combined onto the output fiber.
That is the shipping answer, and it scales cleanly to the 8 and 16 lines the MSA
grids call for. Past a few dozen lines the die count, the combining loss, and the
per-die wavelength trimming start to hurt, which is why a single device that emits
a whole comb is attractive. Three device classes compete.

\newthought{Quantum-dot mode-locked lasers} (\term{QD-MLL}s) are the front-runner
for a monolithic O-band comb. Mode locking in a single cavity produces evenly
spaced lines at the cavity round-trip rate; quantum-dot gain adds low RIN, a
near-zero linewidth-enhancement factor, and strong optical-feedback tolerance, the
same properties that make quantum-dot lasers attractive for isolator-free
co-packaging~\firstcite{qdcomb}\cite{qdrin}. Reported O-band demos carry
14$\times$100~Gb/s PAM4 over 10~km at $\sim$284~fJ/bit, and isolator-free variants
target interconnect capacity beyond 3.2~Tb/s. These are research results, not
qualified products, so treat the line counts and efficiencies as provisional.

\newthought{Kerr microcombs} take the opposite route: pump one high-$Q$
microresonator and let four-wave mixing fill in many evenly spaced lines on a
chip~\firstcite{microcomb}. The line count and the spacing uniformity are
excellent, and a 2025 demonstration added a monolithic demultiplexer that
autonomously locks to and tracks the comb lines. The catch is power.
Pump-to-comb conversion efficiency is modest, so each line leaves the chip weak and
usually needs a booster or per-line amplifier before it reaches the modulator bank
(\Cref{sec:soa-distribution}). Microcombs also need a clean pump laser and careful
thermal control to hold the soliton state.

\newthought{Gain-switched and quantum-dash combs} sit between the two: a directly
driven laser produces a flatter, lower-line-count comb with simple electronics.
They have reached multi-terabit aggregate rates in the lab but see less datacom
traction than QD-MLLs.

For any of them the contract from the MSA does not change: the source must hold
per-line power flatness, SMSR, RIN, and grid placement across temperature with
every port active (\Cref{sec:cwwdm-laser}). A comb that delivers 32 lines but
drops 6~dB across the band, or whose edge lines miss the grid, buys nothing over an
array of DFBs the PIC already knows how to drive.

\subsection{Gain and power distribution across the bank}
\label{sec:soa-distribution}

Whatever generates the comb, the light still has to survive the trip to the
modulators. One source feeds a multiplexer, a splitter tree, and per-line routing
before it reaches a ring, and each stage takes its cut. When the source is a
chip-scale comb with weak lines (\Cref{sec:comb-sources}), or when the fan-out is
large, a \term{semiconductor optical amplifier} (SOA) restores the budget.

\newthought{Where the gain sits.} A booster SOA placed right after the comb lifts
every line before the split, so one device pays for the whole fan-out. A per-line
SOA after the demultiplexer instead corrects line-to-line imbalance, at the cost of
one amplifier per wavelength. The receiver-side SOA preamplifier
(\Cref{tab:rxtech}) is a separate job: there the goal is sensitivity, here it is
launch power.

\newthought{The noise-figure cost.} An SOA adds amplified spontaneous emission
(\term{ASE}), and its noise figure sets how much. The quantum floor is 3~dB;
commercial O-band SOAs land near 6--7~dB with roughly 15~dB of gain and about
1.5~dB polarization-dependent gain~\firstcite{soa}. Every dB of noise figure eats
into the signal-to-noise ratio the receiver eventually sees, so an SOA that rescues
launch power can cost link margin if it runs deep into saturation or amplifies an
already-noisy comb. Quantum-dot SOAs grown on silicon are attractive for the same
reason QD lasers are: low noise, wide O-band gain, and CMOS-compatible integration.

\newthought{Holding the bank flat.} Gain is not uniform across the comb span, and
SOAs compress near saturation, so the line that starts strongest is not the line
that ends strongest. Per-line power flatness is a system spec (\Cref{sec:cwwdm}),
held with some mix of source-side pre-emphasis, gain-tilt control, and per-line
trimming. The alternative to any distribution-side gain is a higher-power source,
which is why array-of-DFB designs that already meet the launch budget often skip
the SOA entirely.

\subsection{Polarization on the CW distribution path}
\label{sec:polarization}

IM/DD is forgiving about polarization where it counts most. The photodiode is a
square-law detector, so the received state of polarization does not affect the
recovered bits (\Cref{sec:coherent-boundary}). A standard single-mode drop to the
receiver therefore needs no polarization control. The external-source and CW-WDM
architectures move the sensitive part upstream, onto the path between the laser and
the modulator.

\newthought{Where it matters.} Three elements on the CW feed care about the launched
state. A \term{TFLN} Mach--Zehnder drives the electro-optic effect through
TE-polarized light on one crystal axis (\Cref{sec:tfln-mzm}); light on the wrong
axis sees little modulation. Silicon grating couplers are polarization-selective by
construction, so coupling loss swings with the input state, a
\term{polarization-dependent loss} (PDL) that comes straight off the launch budget.
And a booster or per-line SOA adds polarization-dependent gain
(\Cref{sec:soa-distribution}), so a drifting state becomes a drifting per-line
power. None of these sit after the photodiode, so none show up in a receiver-side
budget: they act on light that has not yet been modulated.

\newthought{How it is held.} Keep the CW feed on \term{polarization-maintaining}
(PM) fiber and PM connectors from the external laser (\Cref{ch:lasers}) to the
modulator input, launched on the coupler's preferred axis. On-chip the light is
already single-polarization once it is in a TE waveguide, so the discipline is
really about the fiber run and the mate at the package. A co-packaged design with
the laser on the same substrate shortens that run to millimeters and sidesteps most
of the problem; an external-laser design pays for a PM path and its alignment
tolerance instead.

\section{Lock validation playbook}
\label{sec:lock-validation}

Instruments and BER methods live in \Cref{ch:validation}. What is special to WDM
is the order: you cannot trust a BER number on a ring bank until the comb is
identified, the resonances are parked, and the lock loops hold under neighbor
heat. The sequence below is the usual bring-up; skip a step and you will debug
the wrong domain.

\paragraph{Bring-up order.}
\begin{enumerate}
\item \textbf{Grid ID:} confirm each CW line (or DFB) is on the assigned channel
      with an OSA / wavemeter (\Cref{sec:cwwdm-laser}).
\item \textbf{Coarse align:} park rings near resonance with open-loop heater
      sweeps; check through/drop monitors.
\item \textbf{Close lock:} enable the feedback loop; verify capture on every
      $\lambda$ at operating optical power.
\item \textbf{Stress neighbors / temperature:} max case $T$, neighbor heaters and
      traffic on (\Cref{sec:thermal-xtalk,sec:prod-corners}). Confirm hold, not
      just capture.
\item \textbf{Close the link:} BER / TDECQ / sensitivity on the weakest lane
      first (\Cref{sec:tdecq,sec:link-budget}).
\end{enumerate}

\paragraph{Bisect laser versus ring.}
If one $\lambda$ unlocks or walks, change one actuator at a time:
\begin{itemize}
\item change laser TEC / current with ring heater fixed: if the error follows the
      laser, the source or its locker is wrong;
\item change ring heater with laser fixed: if the error follows the heater, the
      ring, monitor PD, or thermal crosstalk is wrong;
\item if both look fine but OMA is low, inspect MUX imbalance and connector/ORL
      (\Cref{tab:mux-budget,sec:optical-channel}).
\end{itemize}

\paragraph{Fleet telltales.}
Slow BER creep with rising bias on one line is often laser wear-out
(\Cref{sec:wearout-modes}). Sudden unlock under neighbor load with healthy LIV is
thermal crosstalk or lock firmware. One dark lane with neighbors up is COD, FAU,
or a single ring/heater fail; classify with \Cref{sec:fleet-triage} before you
open an 8D on the wrong supplier.

\section{What wavelength locking implies about an architecture}

Because locking is only worth its complexity under specific conditions, its
presence narrows the design space considerably.

\begin{table*}[ht]
\refstepcounter{table}\label{tab:wdm-read}%
\centering
\setlength{\tabcolsep}{5pt}%
\begin{tabular}{@{}p{0.30\linewidth}p{0.66\linewidth}@{}}
\toprule
Implication & Strength \\
\midrule
Some form of WDM is in use & Near-certain: locking only matters with WDM. \\
Dense (D)WDM rather than coarse CWDM & Likely: locking implies tight spacing. \\
Microring-based silicon photonics with external multi-wavelength (CW-WDM) sources, often trending toward co-packaged optics & Common in modern AI interconnects (e.g. Broadcom and NVIDIA Quantum-X programs), driven by fiber-count pressure at scale. \\
Discrete DFB/EML DWDM (no rings) & Also possible; locking alone does not prove rings. \\
\bottomrule
\end{tabular}
\par\medskip
{\raggedright\footnotesize\textbf{Table~\thetable.} What a wavelength-locking requirement implies.\par}
\end{table*}

The solid conclusion is that \textbf{WDM is present and wavelength control is
central}; whether the implementation is ring-based silicon photonics with external
multi-wavelength sources or discrete DFB/EML DWDM depends on the specific system.

\section{Engineering lens}
\label{sec:wdm-engineering-lens}

\subsection{How it works}

WDM buys fiber count back by stacking wavelengths, and the price is control: once
the grid is tight or the modulator is resonant, laser and filter must stay locked.
The chapter's lock loops, thermal budgets, and source specs all exist to hold that
lock across temperature and neighbor load.

\subsection{How it is measured}
\label{sec:wdm-how-measured}

Start with an OSA or wavemeter to identify every source line, then sweep each ring
or filter while recording through-port, drop-port, heater current, monitor-PD
signal, and lock error. Measure insertion loss and adjacent-channel leakage per
lane. Close the loop and repeat BER, OMA, and TDECQ under case-temperature ramps,
neighbor heater activity, source-power spread, and restart. Capture range and hold
range are different acceptance tests; \Cref{sec:lock-validation} gives the order,
and the source measurements follow the CW-WDM MSA methods~\cite{cwwdm}.

\subsection{How it fails}
\label{sec:wdm-how-fails}

One degraded lane points toward its laser line, ring, filter channel, heater,
monitor, or local fiber attach. All lanes moving together points toward the shared
CW source, thermal controller, supply, polarization path, or common MUX. Capture
failures occur during startup or a large temperature step. Hold failures occur
after lock, often under neighbor heat, power droop, or control-loop interaction.
Treat those as separate signatures.

\paragraph{One lane fails.}
Possible causes, roughly in order of how often each shows up in the field:

\begin{itemize}
\item \textbf{Wavelength locker failure:} the etalon-based error signal drifts or
      the locker photodiodes degrade, so the loop steers the laser to the wrong
      setpoint even though the laser itself is healthy (\Cref{sec:locking-techniques}).
\item \textbf{Ring drift:} the microring's own resonance walks off the assigned
      $\lambda$ under local heating or process aging, independent of the laser
      (\Cref{sec:siring}).
\item \textbf{AWG issues:} an arrayed-waveguide grating channel shifts passband
      center or picks up excess adjacent-channel crosstalk, clipping one lane at
      the filter edge while neighboring channels stay clean
      (\Cref{sec:wdm-hardware,tab:mux-budget}). Distinguish from a laser or ring
      fault by sweeping the source across the passband and watching for a
      grid-alignment signature rather than a lock-error signature.
\item \textbf{Thermal crosstalk:} an adjacent heater's disturbance exceeds this
      lane's hold-range budget while its own actuators read nominal
      (\Cref{sec:thermal-xtalk}).
\end{itemize}

\paragraph{Intermittent failure.}
Failures that appear and clear on their own point to control-loop dynamics rather
than a broken part:

\begin{itemize}
\item \textbf{Lock acquisition issues:} the loop fails to capture reliably from a
      cold or power-cycled state, so failures correlate with restarts or ELS
      hot-swaps rather than with steady-state operation (\Cref{sec:lock-validation}).
\item \textbf{Heater control instability:} the thermal feedback loop oscillates or
      overshoots, often triggered by a control-loop bandwidth that fights the data
      path or a gain that was tuned for a different neighbor-load condition. Shows
      up as a heater current that hunts rather than settles.
\item \textbf{Temperature excursions:} a transient case-temperature ramp (fan
      failure, load step, HVAC event) pushes the loop past its hold range
      momentarily; the lane recovers once the transient passes, which is the
      distinguishing signature versus a genuine hardware fault.
\end{itemize}

\failuremode{Wavelength drift}{BER rises with temperature, one WDM lane moves, and
lock error or heater demand grows.}{Laser wavelength drift, ring resonance drift,
thermal coupling, a saturated TEC or heater, or a weak monitor signal.}{OSA or
wavemeter, heater and TEC current, lock error, neighbor activity, and per-lane
BER.}{Restore thermal headroom, widen capture only after noise is checked, improve
calibration, and reduce shared thermal or supply coupling.}

\subsection{How it is debugged}
\label{sec:wdm-how-debugged}

First decide whether the fault affects one lane or all lanes. Freeze heater and
laser actuators long enough to measure the open-loop spectrum. Move the source
while holding the ring fixed, then move the ring while holding the source fixed.
If neither follows the failing BER, inspect MUX loss, polarization, connector ORL,
and the electrical lane. For intermittent unlock, align lock-error, heater,
temperature, supply, and neighbor-traffic traces on one time axis. A control loop
cannot be debugged from final register values alone.

\debugstory{One lane failed only after adjacent lanes began traffic.}{The source
line stayed on grid, while the suspect ring heater railed and lock error grew with
neighbor temperature.}{The optical source and receiver passed; the resonance was
being pushed out of hold range.}{Thermal coupling from the adjacent heater exceeded
the control-loop budget.}{The heater map and feed-forward terms were corrected,
and neighbor-load hold testing became an ATP corner.}

\section{Interview and design review questions}
\label{sec:wdm-design-review}

\paragraph{Concept.}
\begin{itemize}
\item Why is WDM preferable to adding fibers at scale?
\item Why does a microring modulator require wavelength locking while a
      Mach--Zehnder does not?
\item What happens if laser wavelength drifts by 2~nm in a 200~GHz-spaced WDM
      system?
\end{itemize}

\paragraph{Design.}
\begin{itemize}
\item What are capture and hold ranges across process, voltage, temperature, and
      source-power spread?
\item Which resource is shared across lanes, and what does its failure signature
      look like?
\item How much heater and TEC headroom remains at the worst thermal corner?
\item Can telemetry distinguish laser drift, ring drift, MUX loss, and monitor-PD
      error without opening the package?
\end{itemize}

\paragraph{Debug.}
\begin{itemize}
\item One WDM lane fails while neighbors are healthy. How do you determine whether
      the cause is the laser, ring, filter, heater, or connector?
\item All lanes unlock after a temperature step. What shared resource do you check
      first?
\item Which calibration or startup error can escape production and fail only after
      a field restart?
\end{itemize}

\paragraph{Manufacturing and operations.}
\begin{itemize}
\item What production test proves lock hold under worst-case neighbor heat?
\item How do you qualify a new CW-WDM source lot against the ATP?
\item What field telemetry distinguishes a source failure from a ring-bank failure?
\end{itemize}

\keyidea{Wavelength locking is a control-system problem wrapped around an optical
link. Measure the grid, capture, hold, shared thermal paths, and per-lane margin.
Then use one-lane versus all-lane behavior to split the source, ring bank, MUX,
thermal controller, and common supplies.}
