\chapter{Optical validation}
\label{ch:validation}

A datasheet that closes on a quiet bench is not a product. \emph{Validation} is
the work of proving a link meets spec and will keep meeting it across the
temperatures, hosts, connectors, and aging the fleet will actually see. That
discipline is the second theme of this book. This chapter walks the ladder from a
single device to a deployed fleet, the metrics measured at each stage, module and
system bring-up under production-like corners, and the data-driven debug mindset
the work demands.

\section{The validation ladder}

Optical programs fail in the same places again and again: a part that looks good
in characterization but cannot bring up on a production host, or a module that
passes ATP and then unlocks under neighbor heat. The practical way to avoid those
gaps is a ladder of increasing scale. Each rung answers a sharper question.

\begin{table*}[ht]
\refstepcounter{table}\label{tab:ladder}%
\centering
\setlength{\tabcolsep}{5pt}%
\begin{tabular}{@{}p{0.14\linewidth}p{0.26\linewidth}p{0.56\linewidth}@{}}
\toprule
Stage & Question & Typical activity \\
\midrule
Component & In spec? & laser LIV, wavelength/RIN/SMSR; PD responsivity and bandwidth; modulator response \\
Module bring-up & Does it link? & power-on, CMIS, CDR/BER (\Cref{sec:bringup}) \\
Characterization & How much margin, across corners? & sweep temperature, voltage, aging; TDECQ, sensitivity, link budget \\
Interop / system & Works with the switch/NIC/ASIC? & chassis thermal, host SerDes, prod-rep corners (\Cref{sec:bringup}) \\
Production readiness & Can the supplier build it at yield? & ATP, SPC, DVT/PVT gates (\Cref{sec:supplier-exec}) \\
Scaled deployment & Does the fleet stay up? & field telemetry, RMA, FIT tracking; triage tree (\Cref{sec:fleet-triage}) \\
\bottomrule
\end{tabular}
\par\medskip
{\raggedright\footnotesize\textbf{Table~\thetable.} From bench to fleet.\par}
\end{table*}

\section{The core IM/DD measurements}

Once the ladder is clear, the measurement list is organized around isolation:
transmitter, channel, and receiver. That split is older than PAM4. Long before
TDECQ, field engineers learned that a dark link can be a dead laser, a dirty
connector, or a dead TIA, and that guessing which one burns hours. Bisecting those
three domains is still how you keep debug from turning into simultaneous retunes
of everything.

\subsection{Transmitter}

Start with the light leaving the faceplate or the CPO fiber array. For PAM4, the
headline metric is \term{TDECQ}\firstterm{TDECQ} (transmitter and dispersion eye
closure quaternary): a reference equalizer is applied to the captured eye and the
residual penalty is reported in dB (\Cref{sec:tdecq}). Alongside it you read
\term{OMA}\firstterm{OMA} (outer), extinction ratio, and \term{RLM}\firstterm{RLM}
(level linearity), plus wavelength, spectral width, and RIN with a bias-driver
versus feedback bisect (\Cref{sec:laser-params,sec:laser-drivers,sec:rin-values}).

What else you add depends on the transmitter style. Laser-bearing modules need
LIV, threshold, slope, SMSR, and chirp checks for DMLs
(\Cref{sec:laser-params,sec:dfb-eml}). External MZMs (TFLN or silicon) need EO
$S_{21}$, $V_\pi$, quadrature bias versus temperature, and driver-path eye
symmetry at baud (\Cref{sec:simzm,sec:tfln-mzm,sec:tdecq}). Microring banks need
resonance alignment, thermal tuning, neighbor crosstalk, and peaking-network EO
$S_{21}$ (\Cref{sec:siring,ch:wdm}). The point of the list is not completeness for
its own sake: it is knowing which instrument answers which hypothesis when the
eye closes.

\subsection{Channel}
\label{sec:optical-channel}

If the transmitter looks clean into a golden receiver and the link still fails,
the channel is next. Insertion loss from fiber, connectors, MUX/de-MUX
(\Cref{sec:wdm-hardware}), and on-chip coupling (\Cref{sec:simzm,sec:siring}) is
the first ledger line; plan about 1--3~dB per fiber interface. Chromatic
dispersion (\Cref{sec:chirp-dispersion}) matters more on FR-class SMF sweeps than
on short DR links. Optical return loss (ORL) is the quiet killer: reflections
back into the laser raise RIN and seed burst errors, which is why many DR/FR
modules still carry isolators while some CPO engines rely on design margin and
monitor photodiodes instead (\Cref{sec:rin-values,ch:lasers}). Fiber attach
(MPO/MTP, FAU, grating couplers) shows up as both yield and reliability
(\Cref{sec:photonic-packaging}).

\subsection{Receiver}

Receiver work asks whether the front-end can still decide bits at the OMA that
survives the channel. Measure sensitivity (minimum OMA for the target BER) and
stressed-receiver sensitivity with a calibrated stressor for margin
(\Cref{sec:secq}), plus overload before the TIA saturates. Underneath those
system numbers sit the photodiode/TIA pair: responsivity, bandwidth, and
input-referred noise (\Cref{sec:pd-tia,ch:models}).

\subsection{Link level}

Only after Tx, channel, and Rx each look sane do you trust a full-link verdict:
pre-FEC BER against the KP4 threshold (\Cref{sec:kp4}), post-FEC BER, FEC
symbol-error histograms, and a signed link-budget ledger from transmitter OMA to
receiver sensitivity with penalties and remaining margin. That ledger is the
document you argue from in DVT; the BER alone is not.

\section{Transmitter and dispersion eye closure quaternary (TDECQ)}
\label{sec:tdecq}

\term{TDECQ} (transmitter and dispersion eye closure quaternary) deserves a closer
look because it is the metric that governs PAM4 transmitter acceptance. It answers
a specific question: \emph{how much worse is this transmitter than an ideal one,
after a realistic receiver has done what it can to clean up the signal?}

\subsection{How it is measured}

\begin{enumerate}
\item \textbf{Capture.} The optical waveform is acquired on a sampling
      oscilloscope (a DCA) through a standardized reference receiver (a
      fourth-order Bessel--Thomson filter at roughly half the baud rate) so every
      lab measures the same bandwidth.
\item \textbf{Equalize.} A defined \term{reference equalizer}, a
      \term{feed-forward equalizer}\firstterm{FFE} (FFE) with a small, bounded number of taps (commonly
      up to five), is applied. This models the modest equalization a real receiver
      would perform, so the transmitter is not penalized for \term{ISI}\firstterm{ISI} the system can
      remove anyway.
\item \textbf{Histogram.} Two narrow vertical histogram windows are placed inside
      the symbol (near 0.45 and 0.55 of the unit interval). The noise distribution
      is evaluated at the three PAM4 decision thresholds.
\item \textbf{Compute.} The algorithm finds the RMS Gaussian noise
      $\sigma$ that, added to the equalized signal, would just reach a target
      symbol error ratio of $4.8\times10^{-4}$ (the SER consistent with the KP4
      pre-FEC budget). TDECQ is the ratio, in dB, of the noise an \emph{ideal}
      transmitter could tolerate to the noise \emph{this} transmitter can
      tolerate:
      \[
        \mathrm{TDECQ} = 10\log_{10}\!\left(\frac{\sigma_{\text{ideal}}}
        {\sigma_{\text{measured}}}\right).
      \]
\end{enumerate}

A worse transmitter tolerates less added noise before failing, so
$\sigma_{\text{measured}}$ shrinks and TDECQ rises. Lower is better; typical
100--200G/lane specifications cap it in the low single-digit dB range.

\subsection{Related quantities and failure signatures}

\begin{description}
\item[SECQ] \firstterm{SECQ} the stressed-eye counterpart used on the receiver side, adding a
      calibrated stressor to test margin rather than transmitter quality alone.
      See \Cref{sec:secq}.
\item[RLM (relative level mismatch)] measures how evenly the four PAM4 levels are
      spaced; poor RLM (uneven levels) inflates TDECQ.
\end{description}

Because TDECQ folds several impairments into one number, the way it fails is
diagnostic: uneven levels point to modulator or driver linearity (RLM);
residual eye closure the equalizer cannot fix points to excess ISI or limited
bandwidth; a noise-limited result points to low OMA, RIN, or reflections. For
external MZMs (TFLN or silicon), also check EO $S_{21}$ bandwidth, $V_\pi$ and
bias quadrature drift with temperature, and RF return loss on the driver-to-modulator
path (\Cref{sec:simzm,sec:tfln-mzm,sec:siring}). This is
why \term{LPO}, which removes the module's own DSP, raises the stakes on
transmitter quality: there is less downstream equalization to hide behind, so
TDECQ-class metrics become even more central.

\section{SECQ and stressed-receiver testing}
\label{sec:secq}

\term{SECQ} (stressed eye closure quaternary) mirrors TDECQ on the \emph{receiver}:
instead of scoring transmitter quality with a reference equalizer, the test applies
a calibrated optical stressor (attenuation, ISI template, optional RIN) and asks
how much margin remains before the receiver hits the target pre-FEC BER.

Stressed-receiver sensitivity and overload tests (\Cref{sec:sensitivity}) use the
same philosophy: bracket the operating OMA range with impairments the link will see
in the field. For LPO, where the module DSP is gone, SECQ-style margin on the
host-side receiver (\Cref{sec:equalization,sec:conditioning}) is as important as
TDECQ on the transmitter.

\section{Instruments}
\label{sec:instruments}

A failing PAM4 link rarely announces which block is wrong. The bench is how you
force the answer: each instrument isolates one failure mode, and the loopback
topology tells you which side of the optical connector owns it.

\begin{description}
\item[DCA] \firstterm{DCA} (digital communication analyzer): sampling scope for PAM4 eyes, TDECQ,
      OMA, RLM (\Cref{sec:tdecq}). Needs a reference receiver filter matched to the
      PHY under test.
\item[BERT]\firstterm{BERT} bit-error ratio at pre- and post-FEC; FEC symbol histograms
      (\Cref{sec:kp4}).
\item[OSA]\firstterm{OSA} wavelength, SMSR, side modes, RIN estimates (\Cref{ch:lasers,sec:rin-values}).
\item[VOA / stressor assembly]\firstterm{VOA} calibrated attenuation and optional ISI for SECQ and
      sensitivity sweeps.
\item[Power meter] average power; pair with DCA for OMA.
\item[Thermal chamber + TEC controller] corner validation; essential for rings
      (\Cref{sec:siring,ch:wdm}) and laser grids.
\end{description}

Use electrical loopback (host SerDes), optical loopback (Tx$\to$Rx on module), and
golden-host/golden-module interop to bisect faults (\Cref{sec:optical-channel}).
If the fault follows the module under golden-host swap, stop blaming the SerDes;
if it stays with the host, stop opening laser FA.

\section{Building a link budget}
\label{sec:link-budget}

A link budget is a signed dB (or power) ledger from transmitter to receiver.
For IM/DD short reach, start from outer OMA at the Tx faceplate and subtract every
loss and penalty until you compare against receiver sensitivity (with target BER
and KP4 pre-FEC threshold, \Cref{sec:kp4,sec:sensitivity}).

\paragraph{Typical ledger (single-mode DR class).}
Start from Tx OMA on the DCA (or from average power and ER). Subtract
connector/coupling loss (1--3~dB per mated pair; fiber attach in CPO), fiber loss
($\sim$0.3--0.4~dB/km at 1310~nm; often negligible at 500~m), and MUX/de-MUX if
WDM (2--5~dB per stage, \Cref{sec:wdm-hardware}). Add penalties for TDECQ (already
in the OMA spec for many PMDs), dispersion (\Cref{sec:chirp-dispersion}), and
ORL/RIN reflection (\Cref{sec:optical-channel,sec:rin-values}). Compare the
remainder to stressed sensitivity at pre-FEC BER $2.4\times10^{-4}$, and keep
1--3~dB+ of production margin (more for fleet corners).
Electrical budgets parallel this for the host-to-module path: COM and pre-FEC BER
(\Cref{sec:com,sec:eye-budget,sec:equalization}). LPO requires \emph{both} ledgers
to close without module DSP help.

\section{Module management: CMIS}
\label{sec:cmis}

\subsection{What CMIS is, and why an optical engineer cares}

\term{CMIS}\firstterm{CMIS} (Common Management Interface Specification) is the
vendor-neutral management layer between a host (switch ASIC, NIC, or test fixture)
and a pluggable or on-board optical module. The host talks to the module over a
two-wire bus (TWI, I2C-like) through a paged register map: identity, power mode,
alarms, per-lane monitors, and (at 224G/448G) link-training and host signal-
integrity tuning extensions~\cite{cmis53}. CMIS covers QSFP-DD, OSFP, COBO, ELSFP,
and CPO engines that expose the same management contract.

You touch CMIS on every bring-up and every field triage. It is how the host learns
what module is seated, when lasers may turn on, what Tx/Rx power and temperature
look like, and whether a link failed at the management layer or the optical layer.
A module that passes BER on a bench with lasers forced on but cannot reach
ModuleReady on a production host will fail in the fleet (\Cref{sec:bringup}).

\subsection{The module state machine}

CMIS defines a module state machine the host drives. After presence detect and
power application, the module stays in low power until the host releases
\texttt{LPModeL} (or the CMIS 5.x \texttt{LowPwr} equivalent). The host reads
identifier pages, clears sticky interrupts, and steps the module toward
ModuleReady. Only then should Tx lanes or ELS lasers enable. ELSFP modules that
emit before ModuleReady are a reject: the host did not authorize light
(\Cref{sec:elsfp}).

Data paths have their own state machines in CMIS 5.x (data path states, and
network path states for media-side links). For bring-up, map the sequence in
\Cref{sec:bringup} onto these transitions: presence and Vcc, CMIS init and
ModuleReady, enable light, optical path check, electrical lock, traffic, snapshot.
Skipping step 2 and jumping to BER is how interop failures hide until production.

\subsection{The memory map: pages, monitors, control}

The lower memory map holds module identity, status, interrupt flags, and alarm
thresholds. Upper pages hold application descriptors, lane controls, tunable-laser
support, versatile diagnostics (VDM), and command-data-block (CDB) firmware
messaging~\cite{cmis53}. Hosts select an application (lane count, host interface,
media type) before bringing up traffic.

\term{DDM}\firstterm{DDM} (digital diagnostic monitoring) is the telemetry layer
you read at scale: per-lane Tx and Rx optical power, laser bias current when
exposed, module temperature, supply voltage, LOS/LOL flags, and alarm/warning
bits. On WDM parts you also get wavelength or channel ID. This is exactly what
\Cref{sec:fleet-triage} reads before anyone reaches for a DCA. On bring-up, dump
the register map you will use in the field and treat that dump as the golden
reference for later RMA comparisons.

\subsection{CMIS as a validation deliverable}

CMIS correctness is part of production readiness, not a firmware afterthought.
ATP should prove the state machine reaches ModuleReady across voltage and thermal
corners; DDM monitors track bench truth (CMIS Tx power versus DCA, module
temperature versus case $T$); alarms fire at the right thresholds; and firmware
revision is ECO-controlled like laser die revision (\Cref{sec:supplier-exec}).
Multi-source interop failures are often CMIS, media-type, or firmware mismatches,
not marginal TDECQ (\Cref{sec:bringup}). At fleet scale the register map is the
only eyes you have on a module in the rack. If CMIS is wrong, triage starts blind.

\section{Module and system bring-up}
\label{sec:bringup}

Characterization proves a sample can meet metrics on a quiet bench. Bring-up
proves a module (then a system) can be powered, managed, and linked the way
production and the fleet will actually run it. Lab-to-production programs fail in
the gap between those two if you only ever test golden hosts, clean fiber, and
room-temperature faceplates.

\paragraph{Module bring-up sequence.}
Run this order on every new module (pluggable, ELSFP, or CPO engine with CMIS).
Do not skip ahead to BER: a link that ``works'' with lasers forced on and CMIS
ignored will fail the first host that enforces the state machine
(\Cref{sec:elsfp}).

\begin{enumerate}
\item \textbf{Presence and power.} Detect module (\texttt{ModPrsL} or equivalent).
      Apply rails in the host power sequence. Confirm Vcc and module temperature
      in CMIS. Stay in low power (\texttt{LPModeL} asserted or ModuleLowPwr) until
      management is sane.
\item \textbf{CMIS init.} Read identifier, vendor, firmware rev, supported media.
      Clear sticky interrupts. Confirm the state machine can reach ModuleReady
      (or the pluggable equivalent) under host command. Dump the register map you
      will use in the field; that dump is your bring-up golden reference.
\item \textbf{Enable light.} Exit low power; enable Tx lanes / ELS lasers only after
      ModuleReady. Confirm Tx optical power and laser bias (if exposed) against the
      power class. Lasers that come up before the host asks are a reject for ELSFP
      (\Cref{sec:elsfp}).
\item \textbf{Optical path.} Mate fiber (clean first). Check Rx power and
      LOS\firstterm{LOS}. Optical loopback first if the host path is unproven.
\item \textbf{Electrical lock.} Bring host SerDes / module CDR. Confirm
      LOL\firstterm{LOL} clear, equalizer taps not pegged (\Cref{sec:equalization}).
      For LPO, this is the host eye and COM path (\Cref{sec:com,sec:drivers}).
\item \textbf{Traffic.} PRBS\firstterm{PRBS} or live FEC traffic. Pre-FEC BER vs.\ KP4 threshold
      (\Cref{sec:kp4}); glance at FEC symbol-error histogram shape.
\item \textbf{Quality snapshot.} On a Tx-capable path: OMA/RLM/TDECQ or module
      diagnostics that proxy them (\Cref{sec:tdecq}). Record CMIS + BER + case $T$
      together so later triage has a baseline (\Cref{sec:fleet-triage}).
\end{enumerate}

\Cref{tab:bringup-checklist} is the short form you can put on a lab wall.

\begin{table*}[ht]
\refstepcounter{table}\label{tab:bringup-checklist}%
\centering
\setlength{\tabcolsep}{3.5pt}%
\footnotesize
\begin{tabular}{@{}p{0.08\linewidth}p{0.22\linewidth}p{0.28\linewidth}p{0.36\linewidth}@{}}
\toprule
Step & Action & Pass signal & Fail $\to$ first look \\
\midrule
1 & Presence / Vcc / temp & CMIS alive, rails in range & cable, seat, PSU \\
2 & CMIS state machine & ModuleReady (or equiv.) & firmware, TWI, LPMode \\
3 & Enable Tx / ELS & Tx power in class; lasers on only when commanded & bias driver, enable pin, APC \\
4 & Fiber / Rx power & Rx power up; LOS clear & dirty MT, polarity, break \\
5 & CDR / SerDes lock & LOL clear; taps not saturated & host SI, LPO COM, retimer \\
6 & Pre-FEC BER & below KP4 target with margin & Tx quality, ORL, Rx sensitivity \\
7 & Snapshot & CMIS dump + BER + $T$ logged & (needed for RMA later) \\
\bottomrule
\end{tabular}
\par\medskip
{\raggedright\footnotesize\textbf{Table~\thetable.} Module bring-up checklist. LOS =
loss of signal; LOL = loss of lock. Limits come from the ATP and PMD, not from
this table.\par}
\end{table*}

\paragraph{Production-representative corners.}
\label{sec:prod-corners}

Bench corners ($T$, $V$) are necessary and not sufficient. Before you call DVT
or PVT done, run the corners that match how the fleet will abuse the link.
\Cref{tab:prod-corners} is the minimum set for IM/DD + laser programs.

\begin{table*}[ht]
\refstepcounter{table}\label{tab:prod-corners}%
\centering
\setlength{\tabcolsep}{3.5pt}%
\footnotesize
\begin{tabular}{@{}p{0.20\linewidth}p{0.28\linewidth}p{0.24\linewidth}p{0.22\linewidth}@{}}
\toprule
Corner & What to run & Why it catches & Points to \\
\midrule
Chassis thermal & Module in target rack/sled at airflow and power load; not only a quiet chamber on a bench fixture & Faceplate $T$ and TEC load differ from chamber setpoints & derate, TEC, ring unlock \\
Host rails live & Bias / CMIS powered from host supplies with SerDes traffic on & Switching noise into laser bias looks like RIN (\Cref{sec:laser-drivers}) & PSRR, ground, APC \\
Dirty fiber / ORL & Controlled contamination or ORL stress on MT/FAU; clean vs dirty BER & Field installs are not lab-clean; ORL raises RIN and bursts & connector, isolator, feedback \\
Cable plant & Production fiber length, MPO count, and bend radius & Extra loss and reflections eat margin the ledger assumed & link budget (\Cref{sec:link-budget}) \\
ELS hot-swap & Pull/replace ELSFP under traffic (or under controlled traffic stop per CMIS) & Service action the architecture promised (\Cref{sec:elsfp}) & state machine, mate cycles \\
Neighbor load & Adjacent modules/lanes at full traffic and max case $T$ & Crosstalk, shared supply droop, thermal crosstalk on rings & WDM lock, SI, PSU \\
LPO / linear path & Host COM and pre-FEC BER without module DSP crutch & LPO fails here first (\Cref{sec:224g-deploy,sec:com,sec:drivers}) & host FIR, module linearity \\
Voltage corners & Host Vcc min/max with traffic & Brown-out and CMIS glitches & power design, ATP \\
\end{tabular}
\par\medskip
{\raggedright\footnotesize\textbf{Table~\thetable.} Production-representative corners.
A quiet BERT at 25~$^\circ$C with pristine fiber is characterization, not
production readiness.\par}
\end{table*}

\paragraph{System bring-up.}
A module that passes on a golden host can still fail in a real chassis:

\begin{itemize}
\item \textbf{Host path:} run the same sequence on the target NIC/switch ASIC SerDes,
      not only the lab BERT. LPO and half-retimed modules expose host FIR/CTLE
      mistakes that a retimed module hid (\Cref{sec:conditioning,sec:pluggables}).
\item \textbf{Multi-lane / multi-module:} bring all lanes on a port, then neighbors
      in the same cage or tray. Watch thermal rise, supply droop, and CMIS temp
      alarms when the tray is loaded.
\item \textbf{Golden swap:} known-good module in the suspect host slot, then suspect
      module in a known-good slot. That single swap splits host vs.\ module before
      you open FA (\Cref{sec:fleet-triage}).
\item \textbf{Interop:} at least one other vendor host or module if the program
      claims multi-source. Interop failures are usually CMIS, media type, or
      electrical eye, not laser physics.
\item \textbf{ELS / CPO:} external laser modules add a second bring-up: ELSFP state
      machine and optical mate to the engine, then engine bring-up with light
      present (\Cref{sec:elsfp,sec:cpo-status}). A dark engine with a healthy ELS is
      an optical connector or FAU problem until proven otherwise.
\end{itemize}

\paragraph{Exit criteria before ``bring-up done.''}
Call module bring-up done only when: CMIS state machine and enable sequence are
correct; pre-FEC BER meets target on the \emph{target} host with margin; a CMIS+BER+$T$
snapshot is filed; and at least the chassis-thermal, host-rails, and ORL corners
in \Cref{tab:prod-corners} have been run on a representative unit. Call system
bring-up done when golden-swap has split host vs.\ module issues and multi-lane /
neighbor load has not opened a new failure mode. Everything after that is
characterization depth, supplier gates (\Cref{sec:supplier-exec}), or fleet triage
(\Cref{sec:fleet-triage}).

\keyidea{Bring-up is a sequence (presence $\to$ CMIS $\to$ light $\to$ lock $\to$
BER $\to$ snapshot), then a system proof on the real host, then
production-representative corners (chassis thermal, host-rail noise, ORL, ELS
hot-swap, neighbor load). A quiet bench pass is not DVT.}

\section{The debug mindset}
\label{sec:debug}

Debug at this level is data-driven, not opinion-driven. The method is disciplined
bisection: change one domain at a time, and let the measurement tell you whether
the transmitter, the channel, or the receiver moved.

\begin{enumerate}
\item Isolate transmitter versus channel versus receiver, using loopbacks.
\item Sweep temperature and voltage to expose corner-dependent failures.
\item Correlate failures to DSP equalizer tap values (\Cref{sec:equalization}) and FEC symbol-error
      statistics (\Cref{sec:kp4}); these tell you \emph{how} the link fails.
\end{enumerate}

The third step is where modern PAM4 links differ from older eye-mask work. Tap
saturation and FEC histograms often reveal the failure mode before a single
waveform screenshot does. Treat those as primary evidence, not as afterthoughts
logged once BER already fails.

\aside{Worked example. ``The link fails only at high temperature.'' Candidate
causes: laser wavelength drift off the grid or off a ring resonance; TEC
saturation; EAM bias shift; elevated receiver thermal noise; LIV rollover
(\Cref{sec:laser-aging,sec:laser-params}). You bisect with
measurements (spectrum for wavelength, LIV for the laser, receiver sensitivity
at temperature) rather than guessing.}

\section{Fleet and field triage}
\label{sec:fleet-triage}

Lab debug asks: \emph{what is broken on this unit?} Fleet triage asks:
\emph{which bucket does this failure belong in, and who owns the fix?} Optical
programs at fleet scale own that split across performance, reliability, and
manufacturability. Wrong bucket wastes weeks (sending a contaminated connector to
laser FA, or rewriting a SerDes FIR when the laser is rolling over).

\paragraph{Three buckets.}
Classify every field issue before deep root-cause work:

\begin{description}
\item[Performance] the design or operating point does not close the budget under
      the conditions seen in the fleet. Examples: TDECQ/RLM marginal at case
      temperature, host COM tight on LPO, ring unlock under thermal crosstalk,
      ORL-driven RIN that the architecture assumed away. Fix is usually retune,
      derate, firmware, or a design/spec change (\Cref{sec:tdecq,sec:com,sec:siring}).
\item[Reliability] the unit met spec at ship and later degraded. Examples: LIV
      threshold rise, SMSR collapse, EAM bias creep, COD, TEC wear, epoxy creep on
      fiber attach. Fix is Arrhenius-backed life projection, burn-in/screen, derating,
      or field-replaceable lasers
      (\Cref{sec:wearout-modes,sec:laser-aging,sec:gr468,sec:elsfp}).
\item[Manufacturability] a subpopulation fails early or never met the ATP; the
      issue tracks lot, date code, supplier site, or assembly step. Examples: FAU
      misalign yield cliff, solder void on a driver die attach, incoming DPPM spike,
      CMIS register map mismatch on one firmware rev. Fix is SPC, ATP tighten,
      first-article, DPA, and 8D/CAPA with the supplier
      (\Cref{sec:supplier-exec,sec:photonic-packaging}).
\end{description}

A single symptom can sit in more than one bucket until you bisect. The tree below
forces the split with telemetry first, then a short bench confirm, then an RMA
label.

\paragraph{Telemetry you actually read.}
At scale you rarely start with a DCA. Start with what the host and module already
report:

\begin{itemize}
\item \term{CMIS} monitors and alarms: module temperature, supply
      rails, Tx/Rx optical power, laser bias (when exposed), wavelength or channel
      ID on WDM parts, LOS/LOL flags, and interrupt history (\texttt{IntL} on ELSFP;
      \Cref{sec:elsfp}).
\item Host link state: CDR lock, pre-FEC BER, FEC symbol-error histogram shape
      (\Cref{sec:kp4}), equalizer tap saturation (\Cref{sec:equalization}).
\item Fleet context: rack position, case temperature, time since install, date
      code / lot, neighbor-link correlation (one bad fiber vs whole tray).
\end{itemize}

\paragraph{Decision tree (symptom $\to$ bucket).}
\Cref{tab:fleet-triage} is the working map. Read left to right: observe, check
telemetry, pick a provisional bucket, then run the named confirm measurement
before you open an RMA or change a design rule.

\begin{table*}[ht]
\refstepcounter{table}\label{tab:fleet-triage}%
\centering
\setlength{\tabcolsep}{3.5pt}%
\footnotesize
\begin{tabular}{@{}p{0.17\linewidth}p{0.22\linewidth}p{0.14\linewidth}p{0.22\linewidth}p{0.20\linewidth}@{}}
\toprule
Symptom & First telemetry check & Bucket & Confirm on bench / FA & Typical fix owner \\
\midrule
Link never comes up (fresh install) & CMIS presence, Vcc, Tx power flatline, LOS & Mfg or install & Visual fiber/connector; golden module swap; CMIS dump & Ops install; supplier ATP if lot-correlated \\
Intermittent LOS / burst errors & Rx power dropouts; FEC bursts; ORL events & Perf (ORL) or mfg (contam.) & Clean/inspect MT; ORL meter; RIN vs ORL (\Cref{sec:laser-drivers,sec:rin-values}) & Ops cleaning; packaging if repeat RMA \\
Pre-FEC BER high, power OK & Tap saturation; RLM/TDECQ if logged; case $T$ & Perf & DCA TDECQ/RLM; host COM; LPO vs retimed path (\Cref{sec:tdecq,sec:com}) & Host SI / module Tx design \\
BER rises only at high case $T$ & Module temp alarm; Tx power drop; $\lambda$ walk & Perf or reliability & LIV at $T$; OSA grid; TEC current; EAM bias (\Cref{sec:laser-aging}) & Derate / TEC / laser supplier \\
Slow BER creep over weeks/months & Bias current up for same Tx power; SMSR if monitored & Reliability & LIV/SMSR vs ship ATP; Arrhenius lot history & Laser wear-out; ELS replace \\
Sudden hard fail, was healthy & Last good CMIS snapshot; neighbor links OK & Reliability (COD) or mfg (ESD) & Dark LIV; DPA on facet/solder; date-code cluster? & FA + supplier 8D \\
One date code / site fails early & Lot Pareto; burn-in escape rate & Mfg & Incoming SPC vs ATP; FA on sample of lot & Supplier CAPA; hold shipment \\
WDM / ring unlock, power OK & Channel ID; thermal of neighbors; lock-loop status & Perf & Resonance tune; crosstalk; CW-WDM line power (\Cref{sec:lock-validation,sec:thermal-xtalk,sec:cwwdm-laser}) & Lock firmware / thermal design \\
ELSFP swap restores link & Old module CMIS vs new; connector cycles & Reliability or mfg (connector) & Inspect MT; mating-cycle count; laser LIV in returned module (\Cref{sec:elsfp}) & Laser vs connector split in FA \\
\bottomrule
\end{tabular}
\par\medskip
{\raggedright\footnotesize\textbf{Table~\thetable.} Fleet triage map: symptom to
provisional bucket to confirm measurement. Perf $=$ performance (design/operating
point); reliability $=$ time-dependent wear; mfg $=$ lot/process/install
excursion.\par}
\end{table*}

\paragraph{How to walk an incident (order of operations).}
\begin{enumerate}
\item \textbf{Stabilize and capture.} Freeze CMIS dump, host BER/FEC counters, rack
      $T$, and install age before anyone reseats the module. Reseating destroys
      connector evidence.
\item \textbf{Localize.} One link vs tray vs rack. Tray-wide points at power,
      cooling, or a shared ELS. Single-link points at that module, fiber, or host
      lane.
\item \textbf{Classify} with \Cref{tab:fleet-triage}. Write the bucket on the
      ticket before FA starts.
\item \textbf{Confirm} with the smallest measurement that can falsify the bucket
      (golden swap, clean/inspect, LIV, TDECQ, ORL). Do not skip to DPA.
\item \textbf{Act.}
      \begin{itemize}
      \item Performance: change operating policy (derate, FIR, lock loop) or open a
            design/spec defect.
      \item Reliability: replace (ELSFP hot-swap when available), update FIT burn-down,
            tighten burn-in or derate (\Cref{sec:laser-aging,sec:fit-example}).
      \item Manufacturability: quarantine lot, incoming hold, supplier 8D with DPA
            photos and ATP deltas (\Cref{sec:supplier-exec}).
      \end{itemize}
\item \textbf{Close the loop.} Feed the signature back into ATP and CMIS alarm
      thresholds so the next incident trips earlier.
\end{enumerate}

\paragraph{Worked paths (three common tickets).}

\emph{``High temp only.''}
CMIS shows module near thermal limit and Tx power sagging. Bucket starts as
performance (thermal design / derate) until LIV at temperature shows threshold
rise matching an aged lot, which flips it to reliability. Measure OSA wavelength
before blaming the laser: a ring unlock is still performance
(\Cref{sec:siring,ch:wdm}).

\emph{``Random burst errors, average power fine.''}
Check FEC histogram for clustered errors and CMIS for Rx power dropouts. Clean
and measure ORL. If RIN rises with ORL, it is performance/architecture
(feedback). If ORL is fine and bursts track a date code, it is mfg (intermittent
fiber attach). If bursts grow over months at fixed ORL, suspect laser or driver
aging (\Cref{sec:laser-drivers,sec:laser-aging}).

\emph{``ELSFP replace fixed it; returned module looks alive on the bench.''}
Alive LIV with high ORL sensitivity or dirty MT face means connector/ORL
(mfg/ops), not laser death. Dead or kinked LIV means reliability. Split those RMA
codes explicitly or your FIT math will blame the wrong wear-out mode
(\Cref{sec:elsfp,sec:photonic-packaging}).

\paragraph{RMA labels that keep FIT honest.}
RMA\firstterm{RMA} codes should be distinct, not a single ``optics fail'':
\begin{itemize}
\item laser wear-out (LIV/SMSR/EAM aging confirmed);
\item COD / sudden dark;
\item connector / contamination / ORL;
\item fiber attach / FAU;
\item driver / bias electronics;
\item host / SerDes / LPO eye (not module);
\item NFF\firstterm{NFF} (no-fault-found; track these; high NFF means bad triage).
\end{itemize}
NFF rate and lot Pareto are as important as FIT. A rising NFF with clean LIV
points at install practice or intermittent connectors, not Arrhenius.

\keyidea{Fleet triage is bucket-first: performance (budget/design), reliability
(time), or manufacturability (lot/process). Read CMIS and host FEC before you
touch a DCA; confirm with one falsifying measurement; label RMAs so FIT, ATP, and
supplier 8D point at the real owner (\Cref{tab:fleet-triage}).}
