\chapter{Reliability and manufacturing at scale}
\label{ch:reliability}

A link that closes in the lab can still fail the business case if lasers die in
the field or suppliers cannot hold yield. At gigawatt, multi-generation scale,
reliability and manufacturability stop being afterthoughts and become design
constraints: they decide whether you put the laser on the ASIC package or in a
replaceable module, how hard you derate, and what ATP language you freeze with
partners. This chapter covers the vocabulary of failure at scale, the
qualification flows that project field life, and the supplier-execution work these
systems demand.

\section{The language of scale: FIT and DPPM}
\label{sec:fit-dppm}

Fleet arguments need two different numbers. One is about life in the field; the
other is about quality at the factory door.

\begin{description}
\item[FIT (failures in time)]\firstterm{FIT} failures per $10^{9}$ device-hours. Multiply a
      per-laser FIT by the number of lasers in a fleet and by hours to estimate
      failures per day.
\item[DPPM (defective parts per million)] the manufacturing-quality counterpart,
      measured at incoming or outgoing inspection.
\end{description}

\aside{An order-of-magnitude exercise: a
$100{,}000$-GPU cluster with several optical links per GPU has on the order of
$10^{5}$--$10^{6}$ lasers. Even a small per-laser FIT then implies a nontrivial
number of link failures per day, which is exactly why field-replaceable external
laser sources and redundancy are attractive.}

\section{Qualification flows}

Optoelectronics inherited a common qualification language from telecom:
\term{Telcordia GR-468-CORE}\firstcite{gr468}. The core stress tests still show up
on every laser and module program:

\begin{itemize}
\item HTOL (high-temperature operating life) and burn-in.
\item Temperature cycling and damp heat.
\item Electrostatic-discharge and mechanical stress.
\end{itemize}

\term{Arrhenius} acceleration underpins life projection: raising temperature
accelerates wear-out by a factor set by the activation energy, so a bounded
high-temperature test projects years of field life. Screening (burn-in) removes
infant-mortality parts before deployment.

\paragraph{GR-468 in practice.}
\label{sec:gr468}

Telcordia GR-468-CORE is the common qualification language for optoelectronic
modules and discrete lasers. For optical engineers the actionable pieces are
test-plan alignment (map your ATP to GR-468 stress sequences such as HTOL,
temperature cycle, damp heat, and ESD so supplier and customer agree on
pass/fail), activation energy (FIT projections use Arrhenius acceleration;
document $E_a$ and confidence bounds when converting 1000-hour HTOL to field
years), sample-size humility (qualification lots are small; production SPC catches
drift that qual missed, \Cref{tab:npi}), and boundary clarity: qualify the laser
die, the hermetic package, and the module assembly separately when failures split
across those boundaries
(\Cref{sec:photonic-packaging,sec:laser-aging,sec:elsfp}).

\paragraph{GR-1221: the passive-component companion.}
GR-468 covers active optoelectronics. Its companion, \term{Telcordia GR-1221-CORE}
(Generic Reliability Assurance Requirements for Passive Optical Components), covers
the parts GR-468 does not: connectors, fiber couplers, WDM filters and MUX/DEMUX,
splitters, and isolators~\firstcite{gr1221}. It uses the same style of stress
sequence (damp heat, temperature cycling, mechanical, and aging tests) but scores
pass/fail on insertion loss and return loss rather than on LIV. A short-reach link
that leans on an on-package or blind-mate MUX and on external multi-wavelength
sources carries a passive reliability budget that lives in GR-1221, not GR-468
(\Cref{sec:connector-reliability,ch:wdm}). Split the qual the same way you split
the FIT: active laser die under GR-468, silicon under JESD47, passive optics under
GR-1221.

\paragraph{ATP sketch: EML module or ELSFP.}
A short acceptance sketch lives with the qual hooks in \Cref{sec:elsfp}; the full
ATP-as-contract, SPC, and 8D workflow is in \Cref{sec:supplier-exec}. Failures that
pass qual but fail field usually sit in derating policy or connector contamination
(\Cref{sec:laser-aging,sec:fleet-triage}).

\section{Electronics reliability: driver, TIA, and DSP silicon}
\label{sec:ic-reliability}

GR-468 covers the optoelectronic parts of the link: the laser die, the
photodiode, and the hermetic or non-hermetic package around them. The modulator
driver, TIA, retimer, and DSP (\Cref{sec:drivers,sec:pd-tia}) are ordinary CMOS or
SiGe BiCMOS ICs, and they wear out and fail by a different, better-documented set
of mechanisms. Treat them with the semiconductor industry's own qualification
language, not with Arrhenius laser-aging math borrowed from \Cref{sec:laser-aging}.

\paragraph{JESD47: the silicon-side GR-468.}
\firstterm{JESD47} JEDEC JESD47 is the baseline stress-test-driven qualification
flow for a new IC, a device family, or a process change: temperature cycling,
HTOL, HTSL (high-temperature storage life), autoclave or HAST (highly accelerated
stress test) for moisture, and mechanical shock and vibration~\firstcite{jesd47}.
It plays the same role for driver and TIA silicon that GR-468 plays for the
laser: a common list of stresses that a supplier runs once and a customer accepts
instead of renegotiating a qual plan on every design win.

\paragraph{ESD and latch-up: failure modes GR-468 does not test.}
Two mechanisms are specific to ICs and absent from the laser-side wear-out map in
\Cref{tab:wearout-map}:

\begin{description}
\item[ESD] a discharge event during handling or assembly damages a gate oxide or
      junction. Component-level classification uses the human-body model (HBM) and
      charged-device model (CDM) test standards, \term{ANSI/ESDA/JEDEC JS-001} and
      \term{JS-002}~\firstcite{jsesd}. A driver or TIA datasheet HBM/CDM rating is
      the number that protects the part on the factory floor, at fiber-attach and
      wire-bond stations where a laser die is also exposed.
\item[Latch-up] a parasitic thyristor structure in CMOS turns on under an
      overvoltage or current-injection event and holds a low-impedance path until
      power is cycled. \term{JESD78}\firstterm{JESD78} defines the overvoltage and
      $\pm100$~mA current-injection test that classifies susceptibility by supply
      and signal pin~\firstcite{jesd78}. A latched driver IC can look like a dead
      laser on the bench (no light, no LIV signature) until you check the supply
      current instead of the optical path.
\end{description}

Both mechanisms are 100\%-screen or design-margin items, not something you
project with an activation energy. If a driver fails ESD or latch-up in the
field, that is a manufacturability or design-margin bucket item
(\Cref{sec:fleet-triage}), not a wear-out FIT argument.

\paragraph{AEC-Q100: a borrowed grade, not a requirement.}
\term{AEC-Q100}\firstterm{AECQ100} is the automotive industry's qualification
standard for ICs, built on the same JEDEC JESD47/JESD22 stress methods with
tighter ESD targets and named temperature grades from Grade~3
($-40$ to $85$\textdegree C) up to Grade~0 ($-40$ to $150$\textdegree C)~\firstcite{aecq100}.
Datacenter optics does not require Q100; the fleet lives in a controlled data hall,
not an engine bay. It is still a useful signal: a driver, TIA, or retimer die that
also ships in an automotive part number typically carries a published Q100 grade,
and that grade is a fast proxy for the ESD/latch-up margin and temperature-cycle
depth behind the datasheet, without re-running the qual plan yourself.

\paragraph{Where this lands in the ATP.}
Fold IC-level qual into the same acceptance and SPC structure used for the laser
(\Cref{tab:atp-laser,sec:supplier-exec}): require the supplier's JESD47 qual
report and HBM/CDM/latch-up ratings for driver and TIA die at DVT, add an ESD
handling audit to the incoming-QC checklist alongside laser LIV/SMSR sampling,
and treat a driver/TIA silicon revision the same way you treat a laser die
revision or a CMIS firmware rev: an ECO that needs first-article requalification,
not a silent BOM swap.

\section{Wear-out modes to know}
\label{sec:wearout-modes}

Arrhenius math, derating, and the worked FIT example live in
\Cref{sec:laser-aging}. This section is the mechanism catalog: how each failure
shows up in ATP and telemetry, and which triage bucket owns it
(\Cref{sec:fleet-triage}). Do not run process CAPA on a wear-out part, and do not
burn FIT math on a dirty connector.

\paragraph{Infant mortality versus wear-out versus packaging.}
Field failures come in three clocks, and mixing them up wastes CAPA. Infant
mortality is early fails from latent defects; burn-in and HTOL screens remove them
before ship (\Cref{sec:gr468}). Wear-out is gradual or sudden end-of-life in the
laser or EAM under temperature, current, and optical power, projected with
Arrhenius and derating (\Cref{sec:laser-aging}). Packaging and assembly faults
(FAU align, epoxy creep, solder voids, connector wear) often dominate field
returns once lasers are screened (\Cref{sec:photonic-packaging}).
Destructive physical analysis (facet cross-section, EDX, FAU section) is required
when the signature is ambiguous or when you need evidence for supplier 8D
(\Cref{sec:supplier-exec}).

\paragraph{Mechanism map.}
\Cref{tab:wearout-map} is the working list for laser-bearing modules and CPO/ELS
paths. Customize limits in the ATP; keep the classification discipline.

\begin{table*}[ht]
\refstepcounter{table}\label{tab:wearout-map}%
\centering
\setlength{\tabcolsep}{3pt}%
\footnotesize
\begin{tabular}{@{}p{0.16\linewidth}p{0.22\linewidth}p{0.28\linewidth}p{0.26\linewidth}@{}}
\toprule
Mechanism & Observable & ATP / telemetry & Triage bucket \\
\midrule
COD (facet) & Sudden dark or hard fail; was healthy & Dark LIV; DPA facet; date-code cluster? & Reliability (COD) or mfg (ESD) \\
Gradual facet / active region & $I_\mathrm{th}$ up, slope down over life & LIV trend vs ship ATP; HTOL lot history & Reliability (wear-out) \\
SMSR collapse & Side modes rise; modal noise / BER & OSA SMSR vs floor at $T$ & Reliability; watch aging \\
EAM aging (EML) & TDECQ/RLM creep at fixed bias & EAM bias sweep + DCA; bias creep log & Reliability (EAM) \\
RIN rise & BER floor up; feedback sensitive & RIN @ ORL; isolator / connector check & Perf if ORL; reliability if isolator \\
TEC / thermal control & Unlock or $\lambda$ walk; LIV may look fine & TEC current, case $T$, lock status & Perf (lock) or reliability (TEC) \\
Coupling / FAU / solder & Loss step, intermittent LOS, shock-related & ORL, mate cycles, DPA FAU/solder & Manufacturability / packaging \\
Driver/TIA latch-up (ESD) & Sudden hard fail; no light, no LIV signature; supply current spikes & Supply current vs bias; JESD78 rating; date-code cluster? & Mfg (ESD) or design margin \\
Connector wear / contamination & ORL creep after repeated mate cycles; RIN floor rise & Mate-cycle count vs IEC~61300-2-2 rating; endface grade (IEC~61300-3-35) & Manufacturability / packaging \\
\bottomrule
\end{tabular}
\par\medskip
{\raggedright\footnotesize\textbf{Table~\thetable.} Wear-out and packaging mechanisms
versus observables. Arrhenius projection and derating for the laser rows:
\Cref{sec:laser-aging}. Electronics stress qualification:
\Cref{sec:ic-reliability}. Connector reliability: \Cref{sec:connector-reliability}.
Field classification workflow: \Cref{sec:fleet-triage}.\par}
\end{table*}

\term{Catastrophic optical damage} (COD) is the sudden facet failure under optical
or electrical overstress. Gradual facet and active-region degradation move
threshold and slope on a slower clock. EAM aging shifts the absorption curve and
shows up as transmitter and dispersion eye closure quaternary (TDECQ) or RLM
drift before the DFB LIV looks dead. TEC and lock failures
mimic optical wear-out until you bisect heater versus laser
(\Cref{sec:lock-validation}). Coupling and FAU faults belong with packaging, not
with Arrhenius $E_a$ arguments.

\section{The reliability bathtub: three failure regimes}
\label{sec:bathtub}

Field failures follow three regimes with different clocks and different fixes.
Mixing them wastes corrective action on the wrong mechanism.

\begin{description}
\item[Infant mortality (early life)] Latent defects from manufacturing: weak
      solder joints, marginal laser die, contamination, firmware bugs that trip on
      first thermal cycle. Burns down rapidly with time. Fixed by burn-in,
      screening, tighter ATP, and first-article control (\Cref{sec:gr468}).
\item[Useful life (constant rate)] Random failures at a roughly steady FIT:
      cosmic-ray-induced single-event upsets, handling damage during unrelated
      service actions, and isolated material defects that passed all screens. Fixed
      by redundancy, field-replaceable modules, and design margin. Note: connector
      wear and contamination accumulate with mate cycles and are usage-driven
      degradation (\Cref{sec:connector-reliability}), not constant-rate random
      failures; budget them separately from the flat-hazard FIT.
\item[Wear-out (end of life)] Physics-driven degradation: laser facet, active
      region, EAM absorption curve shift, TEC aging, epoxy creep. Onset depends
      on temperature, current, optical power, and time. Fixed by derating,
      Arrhenius-based life projection, and planned replacement intervals
      (\Cref{sec:laser-aging}).
\end{description}

A rising failure rate after years of service is wear-out and calls for replacement
planning, not supplier 8D. A cluster of early failures on a new lot is infant
mortality and calls for tighter screens. A steady trickle with no date-code
correlation is useful-life random and calls for redundancy and fast repair. The
triage tree in \Cref{sec:fleet-triage} forces this classification before
corrective action starts.

\section{Yield analysis}
\label{sec:yield-analysis}

Yield is not one number. It splits by stage and by failure mode, and each split
points at a different owner.

\begin{description}
\item[Wafer / die yield] Process-limited: waveguide loss, ring resonance spread,
      heater shorts, photodiode dark current. Caught at wafer probe. Owner: foundry
      SPC.
\item[Assembly yield] Packaging-limited: fiber-array attach alignment, solder voids,
      wirebond pull, epoxy placement. Caught at module ATP. Owner: assembly supplier.
\item[Test yield (first-pass)] ATP-limited: units that fail one or more acceptance
      criteria on first pass. May include measurement-system false rejects (gauge
      R\&R). Owner: test engineering.
\item[Escaped DPPM (post-screen field failures)] Units that passed all screens but
      fail in the fleet. Each escape is either a preventable coverage gap (the
      screen exists but missed the defect) or a residual latent failure that no
      cost-effective screen separates from good units. Owner: quality and
      reliability engineering.
\end{description}

Track yield by ATP row, lot, supplier site, tester, and date code. A yield drop
that correlates with one tester is likely a measurement problem. A yield drop that
correlates with one supplier lot is likely a process problem. A yield drop with no
observed correlation requires further investigation: verify gauge repeatability,
expand stratification (shift, fixture, material lot, firmware), and test guardband
or specification mismatch as hypotheses before concluding. Do not open supplier
corrective action until the measurement system is cleared (\Cref{sec:hvm-test}).

\section{Escaped defect analysis}
\label{sec:escaped-defects}

An escaped defect is a unit that passed every production screen and failed in the
field. Post-screen field failures split into two categories with different
corrective actions:

\paragraph{Preventable coverage escapes.}
A defect that a cost-effective screen could have caught but did not, because the
screen was absent, miscalibrated, or insufficiently stressed. Each one requires an
ATP or process-control change. Common mechanisms in optical modules:

\begin{itemize}
\item \textbf{Contamination:} particles trapped during assembly that only move under
      thermal cycling or vibration, shifting coupling or raising ORL intermittently.
\item \textbf{Marginal alignment:} fiber-attach or FAU position within spec at room
      temperature but outside at the thermal corner, creating a
      temperature-dependent loss that the ATP chamber did not run.
\item \textbf{Weak solder or wirebond:} passes pull test but fatigues under
      repeated thermal cycling, creating intermittent opens or increased resistance.
\item \textbf{Firmware corner:} a state-machine path or calibration table entry that
      production never exercises but the fleet hits on a specific boot sequence,
      temperature bin, or host interaction.
\item \textbf{Packaging stress:} residual mechanical stress from underfill cure or
      lid attach that relaxes over time, shifting alignment or birefringence.
\end{itemize}

For each preventable escape, trace the failure signature back to the earliest point
in the production flow where it could have been caught, and add or tighten the
screen there.

\paragraph{Residual latent failures.}
A defect that no cost-effective screen can separate from good units at the time of
test. Examples include cosmic-ray-induced single-event upsets, rare material
inclusions below detection limits, and process-marginal units that sit inside all
guardbands but fail under a specific field combination of temperature, ORL, and
neighbor load. These go into the residual FIT model, not into ATP. Document why no
reasonable screen exists so the decision is auditable and revisitable as test
technology improves.

\section{Photonic packaging and module-level failures}
\label{sec:photonic-packaging}

Fleet FIT is not only laser wear-out. Once lasers are screened and derated,
module and packaging failures often dominate field returns: the part that shipped
with a clean LIV still loses light after shock, humidity, or a thousand ELSFP mate
cycles. Fiber attach and FAU alignment fail from shock, humidity ingress, and
epoxy creep; CPO fiber-array units add assembly steps that wafer test cannot catch
(\Cref{sec:cpo-status}). Hybrid stacks (TFLN-on-Si, InP laser on Si, flip-chip
drivers) introduce solder voids, underfill cracks, and RF return-loss drift
(\Cref{sec:tfln-mzm,sec:drivers}). Thermal paths matter too: uncooled datacom
versus liquid-cooled XPO/CPO, and TEC failure that looks like wavelength drift off
grid or off ring (\Cref{sec:siring,ch:wdm}).

\paragraph{Connector reliability: MPO, mating cycles, and endface quality.}
\label{sec:connector-reliability}

Multi-fiber connectors are the highest-touch mechanical interface in the fleet:
every ELSFP swap, every fiber-attach unit (FAU)\firstterm{FAU} rework, and every cable-plant
install mates and unmates an MPO\firstterm{MPO}. The MPO/MT ferrule family (rectangular, 6.4~mm
$\times$ 2.5~mm, guide-pin aligned, 8/12/16/24 fibers per row) is standardized in
\term{IEC 61754-7}, split into one-fibre-row and two-fibre-row
parts~\firstcite{iec617547}. That standard fixes geometry, not lifetime; lifetime
comes from two companion test methods. \term{IEC 61300-2-2} specifies the
mate/unmate cycling test connector datasheets are rated against, and \term{IEC
61300-3-35} grades endface scratches, pits, and debris into pass/fail zones on the
fiber core and cladding~\firstcite{iec61300}. TIA-568.3 sets 500 cycles as the
structured-cabling
mating-durability floor; MPO/MTP-class connectors in practice are commonly rated
well above 1000 cycles, but that headroom erodes fast with the wrong cleaning
discipline (\Cref{sec:optical-channel}).

Three practical consequences follow for an ELSFP or CPO fiber-attach program.
First, ORL creep is a mating-cycle and cleaning problem before it is a laser
problem: a rising RIN floor after repeated ELS swaps (\Cref{tab:wearout-map}) is
diagnosed with an IEC~61300-3-35-style endface inspection, not a laser FA request.
Second, mate-cycle count belongs in the same telemetry you already read for CMIS
and DDM (\Cref{sec:cmis}); track it per connector, not per module, since a
connector can outlive several module swaps or vice versa. Third, write the
mating-cycle and endface-grade limits into the ATP explicitly
(\Cref{tab:atp-laser}) rather than inheriting a generic MPO datasheet number: an
ELS bank that hot-swaps weekly reaches a 500-cycle floor in under ten years, and a
CPO fiber array that is field-serviced more aggressively reaches it faster still.

ELSFP cycling adds connector wear and contamination that raise ORL
(\Cref{sec:optical-channel,sec:elsfp}); the mating-cycle and endface-grade limits
above are exactly the numbers that turn ``the connector feels loose'' into an ATP
line item instead of a guess.

Destructive physical analysis (cross-section, EDX) and structured 8D/CAPA with
suppliers close the loop from RMA to design rule
(\Cref{sec:supplier-exec,sec:fleet-triage}). Without that loop, packaging FIT gets
mis-attributed to laser Arrhenius models and the wrong part gets redesigned.

\section{Production test at volume}
\label{sec:hvm-test}

\subsection{Test time is a cost, coverage is a risk}

Every second in the acceptance test plan (ATP) times millions of units is line
capacity and real money. Every skipped measurement is escaped DPPM in the field
(\Cref{sec:fit-dppm}). The core tension in high-volume manufacturing is not
``test or don't test'' but how much coverage you buy per second. The expensive
optical steps are thermal soak and corner runs, TDECQ on a sampling scope, BER
dwell long enough to trust a low pre-FEC target, laser burn-in, and mate-cycle
stress on ELSFP connectors. Some screens are statistical (sample burn-in from a
lot, audit TDECQ on a subset). Others must be 100\%: CMIS state machine sanity,
basic LIV/SMSR pass, and any test that catches a safety or enable-sequence fault
(\Cref{sec:cmis,sec:laser-safety}).

\subsection{Where the test happens: wafer, die, module, system}

Push defect detection as far upstream as correlation allows. Wafer-level or PIC
probe catches process shifts (waveguide loss, ring resonance drift, bad heaters)
before fiber attach and packaging spend. Killing a bad die at probe is orders of
magnitude cheaper than an RMA (\Cref{sec:photonic-packaging}). Module ATP is the
full functional test: optical power class, TDECQ or proxy, sensitivity spot-check,
CMIS bring-up, and connector/ORL on ELS parts. System or golden-host bring-up
catches interop: media type, firmware rev, equalizer defaults, and the corners in
\Cref{sec:prod-corners}. Wafer test cannot catch fiber attach, FAU alignment,
epoxy creep, or connector wear. Those failures must survive to module ATP and,
for some signatures, to fleet telemetry (\Cref{sec:fleet-triage}).

\subsection{ATE-to-bench correlation}

Production testers are built for speed and cost, not lab fidelity. The number that
matters is correlation: does the ATE TDECQ, OMA, or sensitivity track the DCA/BERT
reference within a known offset and spread? Set ATP limits with guardbands derived
from that spread. A drifting tester or a stale golden unit shows up as a yield
cliff or a DPPM escape. Keep a golden module (and golden laser subassembly for
ELS), run gauge R\&R across testers and shifts, and correlate CMIS monitors to
bench instruments the same way you correlate TDECQ (\Cref{sec:cmis}). If the ATE
and the DCA disagree, fix the correlation before you argue with the supplier about
spec.

\subsection{Screens, guardbanding, and SPC}

Burn-in and HTOL screens trade infant-mortality escape rate against test time and
cost (\Cref{sec:wearout-modes,sec:gr468}). Test limits are usually guardbanded
tighter than the customer spec so field DPPM stays inside target under drift.
SPC control charts on LIV, SMSR, RIN, TDECQ, and mate-cycle yield by lot, site,
and date code catch a process shift before it becomes an 8D (\Cref{sec:supplier-exec}).
Production test is a yield, DPPM, and cost trade under a fixed reliability target.
It is not a pass/fail checkbox after the optics already work on a golden bench.

\section{Supplier execution playbook}
\label{sec:supplier-exec}

The supplier path is milestones, performance targets, quality, and
manufacturability triage. That is not a soft skill. It is a concrete contract:
requirements, gates, acceptance tests, process control, and corrective action when
a lot goes wrong.

\paragraph{NPI gates and exit criteria.}
\Cref{tab:npi} is the usual stage map. For lasers and IM/DD modules, write exit
criteria that a supplier can fail clearly, not slogans.

\begin{table*}[ht]
\refstepcounter{table}\label{tab:npi}%
\centering
\setlength{\tabcolsep}{4pt}%
\footnotesize
\begin{tabular}{@{}p{0.10\linewidth}p{0.22\linewidth}p{0.34\linewidth}p{0.28\linewidth}@{}}
\toprule
Gate & Question & Laser / optics exit criteria (examples) & Who signs \\
\midrule
EVT & Does it work at all? & First light; CMIS bring-up sequence (\Cref{sec:bringup}); LIV/SMSR/RIN on engineering samples; one link closes BER & Optical eng + supplier FAE \\
DVT & Spec across corners? & Full ATP at $T$/$V$ corners; prod-rep corners (\Cref{sec:prod-corners}); TDECQ/OMA/sensitivity; GR-468 stress plan frozen; FIT model agreed & Optical eng + reliability \\
PVT & Buildable at yield? & Multi-lot yield vs ATP; SPC charts live; burn-in escape rate; FAIR on production tooling; bring-up on production host & Optical eng + SQE / mfg \\
MP & Sustained quality? & Steady DPPM; RMA Pareto owned; ECO control on CMIS/firmware and process & Program + supplier QM \\
\bottomrule
\end{tabular}
\par\medskip
{\raggedright\footnotesize\textbf{Table~\thetable.} NPI gates with laser-relevant exit
criteria. EVT/DVT/PVT/MP are stage names, not a schedule; dates and sample sizes
belong in the program plan with the supplier.\par}
\end{table*}

Hold a gate if the exit data are missing. Shipping PVT material without frozen
ATP limits is how field NFF and FIT arguments start (\Cref{sec:fleet-triage}).

\paragraph{Requirements and ATP are the contract.}
ATP\firstterm{ATP} and the requirements doc are the contract. Write both and keep
them versioned together:

\begin{enumerate}
\item \textbf{Requirements / PRD slice for the laser path:} fill
      \Cref{tab:laser-prd,sec:laser-reqs} (power class, grid, RIN@ORL, SMSR,
      derating, CMIS, FIT). Version it with the ATP.
\item \textbf{Acceptance test plan (ATP):} the measurable tests that prove those
      requirements on every ship lot (or on a defined sample). Map each ATP line to
      a GR-468 or design-validation stress where life is claimed (\Cref{sec:gr468}).
\end{enumerate}

\Cref{tab:atp-laser} is a working ATP checklist for an EML pluggable or an ELSFP
CW module. Customize limits from the datasheet and the link budget; do not invent
numbers in the ATP itself.

\begin{table*}[ht]
\refstepcounter{table}\label{tab:atp-laser}%
\centering
\setlength{\tabcolsep}{3.5pt}%
\footnotesize
\begin{tabular}{@{}p{0.20\linewidth}p{0.22\linewidth}p{0.28\linewidth}p{0.24\linewidth}@{}}
\toprule
ATP item & Instrument / method & Pass intent & Ties to \\
\midrule
LIV ($I_\mathrm{th}$, slope, kink) & SMU + power meter & kink-free bias window at rated $T$ & wear-out, derate (\Cref{sec:laser-params}) \\
SMSR & OSA & single-mode vs.\ spec floor & modal noise, aging \\
RIN (intrinsic + stressed ORL) & PD + ESA & $\mathrm{RIN}_x\mathrm{OMA}$ / quiet floor & BER floor (\Cref{sec:rin-values}) \\
Wavelength / grid & OSA / wavemeter & channel ID; $d\lambda/dT$ & WDM lock (\Cref{ch:wdm}) \\
Optical power class & power meter & ELSFP / MSA class met & link budget \\
EAM bias / chirp (EML) & bias sweep + DCA & ER, RLM, TDECQ at baud & Tx quality (\Cref{sec:tdecq}) \\
CMIS / TWI bring-up & host or CMIS tool & registers, alarms, state machine & field telemetry \\
Connector / ORL & mate cycles + ORL meter & cycles + endface grade vs IEC~61300 limits & packaging (\Cref{sec:connector-reliability,sec:elsfp}) \\
Burn-in screen & HTOL sample or 100\% screen & infant mortality culled & GR-468 (\Cref{sec:gr468}) \\
Driver/TIA ESD, latch-up & supplier JESD47/HBM/CDM report; sample audit & rating on file; no latch-up under current injection & IC reliability (\Cref{sec:ic-reliability}) \\
Thermal class & chamber at case $T$ & LIV/RIN/CMIS still pass & derate policy \\
\bottomrule
\end{tabular}
\par\medskip
{\raggedright\footnotesize\textbf{Table~\thetable.} Acceptance checklist for laser-bearing
modules (EML or ELSFP). Limits are program-specific; the structure is what you
negotiate with the supplier.\par}
\end{table*}

\paragraph{Incoming QC and SPC.}
Qual lots are small. Production catches drift that qual missed.

\begin{itemize}
\item \textbf{Incoming:} sample or 100\% screen against a subset of the ATP (at least
      power, CMIS, and a laser LIV/SMSR sample). Track DPPM by date code and site.
\item \textbf{SPC}\firstterm{SPC}: control charts on $I_\mathrm{th}$, slope, SMSR, Tx
      power, and burn-in fallout. A process shift is a hold, not a hope.
\item \textbf{First-article / FAIR:} when tooling, epi, or assembly site changes,
      rerun a defined FAIR package before open PO volume. Treat CMIS firmware revs
      the same way as process changes.
\end{itemize}

\paragraph{Excursions: 8D / CAPA.}
When a lot fails ATP, incoming, or field triage lands in the manufacturability
bucket (\Cref{sec:fleet-triage}), run structured corrective action:

\begin{enumerate}
\item \textbf{Contain:} quarantine WIP and ship holds; identify suspect date codes
      in the fleet.
\item \textbf{Evidence pack:} failing ATP rows, CMIS dumps, LIV/SMSR/RIN plots, DPA
      photos (facet, solder, FAU cross-section) compared to a golden unit.
\item \textbf{8D / CAPA}\firstterm{eightD}: root cause with the supplier (process step,
      material lot, firmware), corrective action, and preventive control (ATP tighten,
      SPC limit, poke-yoke).
\item \textbf{Verify:} re-run FAIR on the corrected process; watch field RMA codes
      for that date-code family for a defined burn-in window.
\end{enumerate}

Do not close 8D on ``operator error'' without a control that would have caught it
at ATP. If FA shows laser wear-out on a young unit, it may be a reliability screen
gap, not a supplier process bug; reclassify with \Cref{sec:fleet-triage} before you
argue FIT.

\paragraph{Milestone hygiene with partners.}
Align the partner calendar to gates, not slideware:

\begin{itemize}
\item freeze requirements before DVT samples are built;
\item freeze ATP limits before PVT yield is claimed;
\item freeze FIT/$E_a$ assumptions before reliability marketing numbers ship;
\item require ECO notice on laser die revision, TEC vendor, FAU epoxy, driver/TIA
      silicon revision (\Cref{sec:ic-reliability}), and CMIS firmware.
\end{itemize}

Your job in those meetings is to name the measurement that would kill the gate.
If nobody can point to an ATP row or a corner, the milestone is not real.

\keyidea{Reliability at scale is mechanism discipline plus supplier gates.
Classify failures with the wear-out map (\Cref{tab:wearout-map}) before you argue
FIT or open 8D. Laser wear-out gets Arrhenius and GR-468; driver and TIA silicon
gets JESD47, HBM/CDM ESD, and latch-up ratings (\Cref{sec:ic-reliability}); MPO
connectors get mating-cycle and endface-grade limits
(\Cref{sec:connector-reliability}). Supplier execution is a gated contract:
requirements and ATP at DVT, multi-lot SPC and FAIR at PVT, then 8D with evidence
when a lot or field Pareto says manufacturability. Do not run process CAPA on a
wear-out failure, and do not burn FIT math on a dirty or worn connector.}

\section{From component FIT to fabric availability}
\label{sec:fabric-availability}

The FIT arithmetic in \Cref{sec:fit-example} gives a rate: about $0.6$ laser
failures per day for a fleet of $5\times10^5$ lasers at 50~FIT. That number sizes
the RMA pipeline and the ELS spares bin (\Cref{sec:elsfp}), but it does not say
what a failure costs or how a running job survives one. Two facts turn a
per-component rate into a fabric problem.

First, a training or large inference job is synchronous. A collective
(\Cref{sec:collectives}) waits for its slowest member, so a single dead or slow
link stalls the whole group, not just one endpoint (\Cref{sec:inference-bottlenecks}).
A link that flaps for a second is a stall for every accelerator in that collective.
The optical FIT the earlier chapters budget therefore matters out of proportion to
its share of the parts count.

Second, at cluster scale failures are continuous, not rare. Meta's published
Llama~3 run is the clearest public data point: 16{,}384 H100 GPUs over 54 days
logged 466 interruptions (419 unexpected), roughly one every three hours, while
holding about 90\% effective training time~\firstcite{metallama3}. GPU and HBM3
faults dominated at close to half; network switch and cable faults were 35 events,
8.4\% of the total. The optical link is a minority of hard job stops, but 8.4\% of
a failure every three hours is still tens of network events per run, and the ELS,
module, and connector FIT this chapter budgets
(\Cref{sec:fit-dppm,sec:connector-reliability}) lands in exactly that bucket.

So the design question shifts. It is no longer ``how reliable is one link'' but
``how does a fabric of $10^5$ links keep a job running through a failure every few
hours.'' The answers are architectural, and the optical engineer feeds each one.

\begin{description}
\item[Redundancy and rails.] Rail-optimized topologies (\Cref{sec:topologies})
      already give parallel planes; dual-plane and dual-ToR designs let a lost link
      degrade bandwidth instead of dropping an endpoint. Redundancy multiplies the
      link and laser count, which feeds straight back into the FIT budget: more
      resilience is more parts that can fail.
\item[Detection and reroute.] Transient faults stay below the job. KP4 FEC
      (\Cref{sec:kp4}) absorbs the error bursts a marginal link throws; link-level
      retry and sub-second link-flap detection plus adaptive routing steer traffic
      off a degraded link before the scheduler notices. Vendor fabrics (NVIDIA
      Spectrum-X and Quantum, Broadcom Tomahawk) advertise adaptive or ``cognitive''
      routing and link-level retry for this. Treat the specifics as vendor
      orientation, but the mechanism is why transient optical faults rarely reach
      the hard-stop bucket above.
\item[Topology reconfiguration.] When a link or rack dies for good, an optical
      circuit switch re-wires the topology around it in milliseconds, so the
      scheduler routes around the dead node instead of stalling the pod
      (\Cref{sec:ocs})~\citep{tpuv4ocs}. Component FIT still applies; the fabric
      survives each failure by re-wiring optically.
\item[Sparing and field service.] Hot spare nodes and lanes cover the interval
      between failure and repair. Field-replaceable external lasers (\Cref{sec:elsfp})
      make a dead laser a faceplate swap rather than a fabric outage, which is the
      architectural reason ELS decouples laser FIT from switch FIT. The connector
      mating-cycle and endface budget (\Cref{sec:connector-reliability}) sets how
      many of those swaps the plant survives.
\end{description}

The cost of a failure closes the loop. A hard interruption is lost compute plus
the time to detect, reroute or reschedule, and restart from the last checkpoint.
Fast detection and reroute shrink that lost time, which is the fabric-level reason
the module work in this chapter pays off: derating (\Cref{sec:laser-aging}),
burn-in and screens (\Cref{sec:gr468}), and a tight ATP (\Cref{sec:supplier-exec})
lower the failure rate, and a resilient fabric lowers the cost of each failure that
slips through. The two multiply.

\section{Engineering lens}
\label{sec:reliability-engineering-lens}

\subsection{How it works}

At fleet scale, reliability and manufacturability are design constraints: a modest
per-part FIT times millions of parts is a steady stream of failures. The chapter's
qualification, yield, and supplier discipline all aim at keeping that stream small
and classifiable.

\subsection{How it is measured}
\label{sec:reliability-how-measured}

Reliability is measured as a distribution over stress and time. Qualification
records failures by mechanism, stress, lot, and sample history. Production records
yield, fallout by ATP row, measurement distributions, gauge repeatability, and
escaped defects per million. Fleet data add install age, temperature, firmware,
supplier lot, return code, and no-fault-found rate. Keep the chain from wafer or
die measurement through module ATP to field return so a drift can be traced to its
first observable point (\Cref{sec:hvm-test,sec:supplier-exec,sec:fleet-triage}).

\subsection{How it fails}
\label{sec:reliability-how-fails}

Programs fail at scale through wear, variation, and escapes. Wear shifts a unit
after ship. Variation produces a weak tail across wafer, lot, site, or assembly
line. An escape is a defect the current screen cannot see or a control that was not
run. Yield can drop because the product changed, the process moved, incoming
material moved, calibration changed, or the tester moved. Do not open supplier
corrective action until the measurement system is cleared.

\failuremode{Yield drop}{First-pass yield falls from its stable baseline, often on
one ATP row, tester, lot, or shift.}{Process drift, supplier-lot variation,
assembly change, calibration or fixture drift, software revision, or a changed
guardband.}{Pareto by test and lot, golden-unit history, gauge repeatability,
station correlation, incoming data, and destructive analysis on selected
failures.}{Contain suspect material, clear the tester, identify the first changed
input, correct it, and verify with a controlled lot before release.}

\subsection{How it is debugged}
\label{sec:reliability-how-debugged}

For a yield fall, freeze software, limits, and suspect material. Split by tester,
shift, lot, supplier site, and ATP row. Run a golden unit across stations and a
failing unit on a reference bench. If the failure follows the station, repair the
measurement system. If it follows the unit, inspect upstream process data and
choose failure analysis that can distinguish the top mechanisms. Contain first,
find root cause second, change the process third, and verify the fix on fresh data.

\debugstory{Module yield fell sharply after a supplier lot change.}{The failure
Pareto pointed to one-lane TDECQ. Golden units passed all stations, and failed
units kept the bad lane on the reference bench. Cross-sections showed a shifted
fiber-array attach.}{The electrical path and testers were stable.}{An assembly
fixture change moved one fiber row outside its coupling window.}{The lot was held,
the fixture was restored, first-article coupling checks were tightened, and the
supplier control plan was revised.}

\section{Interview and design review questions}
\label{sec:reliability-design-review}

\paragraph{Concept.}
\begin{itemize}
\item What distinguishes infant mortality, useful-life random, and wear-out
      failures?
\item Why does qualification on a small sample not guarantee field reliability?
\item What does FIT mean, and how does fleet size amplify it?
\end{itemize}

\paragraph{Design.}
\begin{itemize}
\item Which failure mechanisms are accelerated by each qualification stress, and
      which mechanisms are not covered?
\item What is tested on every unit, what is sampled, and why is that escape risk
      acceptable?
\item Can yield and field returns be sliced by wafer, lot, supplier site, assembly
      line, tester, firmware, and date code?
\item What spare, repair, and reroute plan limits the fleet cost of a failure that
      still escapes?
\end{itemize}

\paragraph{Debug.}
\begin{itemize}
\item Yield dropped from 95\% to 70\% on one ATP row. What are your first three
      actions?
\item A field return looks alive on the bench. What does a high NFF rate tell you?
\item What measurement-system check runs before a supplier or process excursion is
      declared?
\end{itemize}

\paragraph{Manufacturing and operations.}
\begin{itemize}
\item Which corrective action changes a control so the same defect cannot escape
      again?
\item What data must a supplier provide at each NPI gate to pass?
\item How do you set guardband limits that balance yield loss against DPPM escape?
\item What is the cost of a wrong RMA classification (laser wear-out versus
      connector contamination)?
\end{itemize}

\keyidea{Production readiness joins mechanism-based qualification, a stable
measurement system, controlled supplier changes, and field data that preserve lot
history. When yield or fleet health moves, contain the population, clear the test
system, find the first changed input, and verify the corrective control with new
data.}
