Artificial intelligence has become an infrastructure problem. Training and
serving frontier models at scale is no longer limited only by the accelerator at
the center of the rack, but by how many accelerators can be wired together
efficiently, reliably, and within a fixed power envelope. That wiring is
increasingly optical, and the optics (especially the lasers inside them) have
become a first-order lever on the cost, power, and reliability of the whole
system.

This book is a concise technical overview of that layer: the short-reach optical
interconnects that stitch together AI datacenters, from in-package optical I/O
out to intra-rack links, deliberately setting aside the 2--10~km campus links that
belong to coherent optics (\Cref{sec:reach}). It concentrates on the subjects that
decide whether these links work at scale: IM/DD physics and vocabulary; lasers
and external light sources; WDM and wavelength locking; quantitative noise and
sensitivity models; validation from bench to fleet; and reliability and
manufacturing at volume. Two chapters bracket those fundamentals: why
inference-scale computing puts the interconnect on the critical path
(\Cref{ch:role}), and how AI datacenter networks are built where optics dominate
cost and power (\Cref{ch:networking}).

\newthought{How to use this handbook.} Start from the requirement, not the
component. Move downward from system requirements to architecture, subsystem,
component, and only then to the physics needed to make or test the decision. A
choice at one layer constrains the next. A VCSEL points toward 850~nm, multimode
fiber, and direct modulation. A DFB at 1310~nm points toward single-mode fiber
and leaves a choice among direct modulation, an EML, or an external modulator.
Neither path is best in isolation. Each closes a different reach, power, cost,
manufacturing, and service model.

Every major chapter asks four questions: How does it work? How is it measured,
and what uncertainty does the measurement remove? How does it fail? How is it
debugged?
The aim is operational judgment. Validation builds confidence by reducing
uncertainty across characterization, margin, environment, qualification, and
production. Debugging narrows hypotheses with measurements rather than matching
a symptom to a memorized list. Throughout the book, watch how power, noise,
timing, and spectral margins erode together before a link crosses its BER limit.

\newthought{On sources.} The industry moves quickly. Where the text cites specific
products or figures (co-packaged-optics programs, per-lane roadmaps, energy-per-bit
trends) it draws on public disclosures current as of early 2026, cited in the
references. Where a claim is an inference rather than an established fact, the text
says so. History and trend notes are included where they help explain why today's
defaults exist, not as a full chronology of the field.

\keyidea{The through-line: at gigawatt, multi-generation scale, the optical
interconnect and its lasers are a first-order lever on power, cost, and
reliability. Everything here serves that argument.}
