% Selected references and further reading. The industry moves quickly; refresh
% the standards drafts and vendor announcements, which change often.
\begin{thebibliography}{99}

\bibitem[Broadcom et al.(2026)]{jalapeno2026}
Broadcom and hyperscaler partner.
\newblock \emph{\href{https://investors.broadcom.com/news-releases/news-release-details/openai-and-broadcom-unveil-llm-optimized-intelligence-processor}{LLM-optimized inference processor announcement (Broadcom, June 2026)}}.

\bibitem[Miller(2009)]{miller2009}
D.~A.~B.~Miller.
\newblock \citehref{miller2009}{Device requirements for optical interconnects to silicon chips}.
\newblock \emph{Proceedings of the IEEE} 97(7):1166--1185, 2009.

\bibitem[Miller(2017)]{miller2017}
D.~A.~B.~Miller.
\newblock \citehref{miller2017}{Attojoule optoelectronics for low-energy information processing and
communications}.
\newblock \emph{Journal of Lightwave Technology} 35(3):346--396, 2017.

\bibitem[Daudlin et al.(2025)]{daudlin2025}
S.~Daudlin, A.~Rizzo, S.~Lee, \emph{et al.}
\newblock \citehref{daudlin2025}{Three-dimensional photonic integration for ultra-low-energy,
high-bandwidth interchip data links}.
\newblock \emph{Nature Photonics} 19:502--509, 2025.

\bibitem[Chang et al.(2023)]{chang2023}
P.-H.~Chang, A.~Samanta, P.~Yan, \emph{et al.}
\newblock \citehref{chang2023}{A 3D integrated energy-efficient transceiver realized by direct bond
interconnect of co-designed 12~nm FinFET and silicon photonic integrated
circuits}.
\newblock \emph{Journal of Lightwave Technology} 41(21):6741--6755, 2023.

\bibitem[Pirmoradi et al.(2025)]{pirmoradi2025}
A.~Pirmoradi, H.~Hao, K.~Omirzakhov, \emph{et al.}
\newblock \href{https://arxiv.org/abs/2507.12452}{A single-chip 1.024~Tb/s silicon photonics PAM4 receiver}.
\newblock arXiv:2507.12452, 2025.

\bibitem[Avicena(2025)]{avicena2025}
Avicena.
\newblock \emph{Ultra-low-power microLED link at 200~fJ/bit (LightBundle)}.
\newblock ECOC 2025 demonstration.

\bibitem[IEEE 802.3dj(2026)]{ieee8023dj}
IEEE P802.3dj Task Force.
\newblock \href{https://www.ieee802.org/3/dj/}{200~Gb/s, 400~Gb/s, 800~Gb/s, and 1.6~Tb/s Ethernet}.
\newblock 200~Gb/s per lane (112~GBd PAM4), KP4 FEC RS(544,514); draft entered SA
ballot February 2026; approval expected $\sim$late 2026.

\bibitem[IEEE 802.3bj(2014)]{ieee8023bj}
IEEE Std 802.3bj-2014.
\newblock \href{https://standards.ieee.org/standard/802_3bj-2014.html}{100~Gb/s Operation Over Backplanes and Copper Cables}.
\newblock Introduced \texttt{100GBASE-KP4} (backplane, PAM4, four lanes) and mandatory
RS(544,514) FEC (Clause~91); the ``KP4'' label for that code.

\bibitem[IEEE E4AI(2025--26)]{e4ai448}
IEEE 802.3 ``Ethernet for AI'' ad hoc.
\newblock \href{https://www.ieee802.org/3/ad_hoc/E4AI/}{E4AI assessment and 400G/lane CFI materials}.
\newblock 3.2~TbE port demand; fast-track 400/800/1600G on 400G/lane; follow-on
3.2T PHYs.

\bibitem[IEEE 400G/lane SG(2026)]{ieee400gpl}
IEEE 802.3 400~Gb/s/lane Signaling Study Group.
\newblock \href{https://www.ieee802.org/3/400GPL/}{400~Gb/s per lane signaling for electrical and SMF
reaches up to 500~m}; chartered 13 March 2026.
\newblock PAR/CSD for scale-up copper and IM/DD optics; aligns with CEI-448G /
3.2T module roadmap.

\bibitem[IEEE 802.1(2018--20)]{ieee8021ae}
IEEE.
\newblock \href{https://standards.ieee.org/ieee/802.1AE/7154}{IEEE Std 802.1AE-2018, MAC Security (MACsec)}
and \href{https://standards.ieee.org/ieee/802.1X/12247}{IEEE Std 802.1X-2020, Port-Based Network Access Control}.
\newblock MACsec: connectionless user-data confidentiality, frame integrity, and
data origin authenticity at the MAC layer, transparent to the MAC client. 802.1X:
port-based access control and key agreement that establishes MACsec associations.
Line-rate encryption for 800G/1.6T links lives in the switch ASIC or NIC, not the
transceiver.

\bibitem[OIF(2025)]{oif224}
Optical Internetworking Forum.
\newblock \emph{\href{https://www.oiforum.com/technical-work/hot-topics/common-electrical-i-o-cei-224g/}{CEI-224G and CEI-448G}} Common Electrical I/O projects
(OIF-FD-CEI-448G-01.0 framework, 2025).

\bibitem[OIF CEI-448G(2025)]{cei448}
Optical Internetworking Forum.
\newblock \href{https://www.oiforum.com/wp-content/uploads/OIF-FD-CEI-448G-01.0.pdf}{Next Generation CEI-448G Framework}
(OIF-FD-CEI-448G-01.0, 2025).
\newblock 448~Gb/s/lane target; PAM4 at 224~GBd ($f_N\approx112$~GHz), PAM6 at
173~GBd ($f_N\approx87$~GHz), PAM8 at 149~GBd; connector-limited channels
$\approx$90~GHz; package BW $\gtrsim$115~GHz at 0.5~mm pitch; CPO die-to-OE as
PAM4 bypass; SDO liaison map (\S2.3) and scale-up/out metrics (Table~1).

\bibitem[UALink(2025)]{ualink200}
Ultra Accelerator Link Consortium.
\newblock \href{https://ualinkconsortium.org/specifications/}{UALink 200G 1.0 Specification}.
\newblock April 2025.
\newblock Open scale-up interconnect; 200G/lane; up to 1\,024 accelerators per pod.

\bibitem[UEC(2025)]{uec10}
Ultra Ethernet Consortium.
\newblock \href{https://ultraethernet.org/specification/}{UEC Specification 1.0}.
\newblock June 2025.
\newblock Scale-out Ethernet stack for AI/HPC; Ultra Ethernet Transport (UET).

\bibitem[IBTA(2023)]{ibta}
InfiniBand Trade Association.
\newblock \href{https://www.infinibandta.org/ibta-specification/}{InfiniBand Architecture Specification} (Volumes 1 and 2).
\newblock XDR release October 2023 (Vol.~1 R1.7, Vol.~2 R1.5): 200~Gb/s/lane
SerDes, 800~Gb/s per four-lane port, 1.6~Tb/s switch-to-switch; NDR at 400~Gb/s
per port. Scale-out fabric alongside Ethernet; reuses QSFP/OSFP form factors and
MPO fiber. Dominant vendor NVIDIA (Mellanox). Provisional: XDR rates are from the
2023 spec announcement; confirm against the published specification.

\bibitem[SNIA SFF(2025)]{snia448}
SNIA SFF Technology Affiliate.
\newblock \href{https://www.oiforum.com/wp-content/uploads/OIF_448G_AI_Workshop-SNIA-2025-03-31_v2_AConstantine.pdf}{SFF 448G project overview} (OIF 448G AI Workshop).
\newblock March 2025.
\newblock Backplane/storage/compute focus; SFF-TA-1043; complements OIF front-panel CEI.

\bibitem[OCP(2025)]{ocpai}
Open Compute Project.
\newblock \href{https://www.opencompute.org/projects/optical-circuit-switching}{Optical Circuit Switching subproject}; Open Systems for AI.
\newblock 2025.
\newblock Open rack/cluster designs; OCS for AI fabric reconfiguration.

\bibitem[Google Jupiter(2022)]{jupiterocs}
Poutievski, L., \emph{et al.} (Google).
\newblock \href{https://doi.org/10.1145/3544216.3544265}{Jupiter evolving: transforming Google's datacenter network via optical circuit switches and software-defined networking}.
\newblock Proc.\ ACM SIGCOMM, 2022.
\newblock MEMS OCS under SDN control; Palomar $136\times136$ OCS, $\sim$2~dB insertion
loss, millisecond switching, circulators for bidirectional links; reported $\sim$30\%
capex, $\sim$41\% power, and 3x reconfiguration gains on a direct-connect topology.

\bibitem[Jouppi et al.(2023)]{tpuv4ocs}
Jouppi, N.~P., \emph{et al.} (Google).
\newblock \href{https://arxiv.org/abs/2304.01433}{TPU v4: An optically reconfigurable supercomputer for machine learning with hardware support for embeddings}.
\newblock Proc.\ ISCA, 2023.
\newblock 4096 accelerators wired through 48 OCS into a reconfigurable 3D torus;
per-job topology and route-around-failure for availability and utilization.

\bibitem[Cignal AI(2025)]{ocsmarket}
Cignal AI.
\newblock \href{https://cignal.ai/2025/12/the-optical-circuit-switching-market-4q25/}{The optical circuit switching market}.
\newblock 2025.
\newblock OCS technology survey (MEMS, liquid crystal, piezoelectric, robotic, silicon
photonic) and transceiver impact: FR single-fiber over DR, circulators, higher-power
optics. Analyst report; figures provisional.

\bibitem[Meta Llama 3(2024)]{metallama3}
Grattafiori, A., \emph{et al.} (Meta AI).
\newblock \href{https://arxiv.org/abs/2407.21783}{The Llama 3 herd of models}.
\newblock arXiv:2407.21783, 2024.
\newblock Reliability snapshot (\S3.3.4): 16{,}384 H100 GPUs, 54-day pre-training run,
466 interruptions (419 unexpected, $\sim$one every 3~hours), $\sim$90\% effective
training time; $\sim$78\% hardware, GPU/HBM3 $\sim$47\%, network switch/cable 35
events (8.4\%).

\bibitem[UCIe(2022--25)]{ucie}
UCIe Consortium.
\newblock \href{https://www.uciexpress.org/}{Universal Chiplet Interconnect Express (UCIe) Specification}.
\newblock Open die-to-die chiplet interconnect (physical layer, protocol stack,
software model). 1.0 (March 2022); 2.0 (August 2024) adds 3D packaging and in-field
manageability; 3.0 (2025) roughly doubles bandwidth. Parallel package/interposer
counterpart to CEI XSR; relevant to co-packaged optical I/O and optical-UCIe work.

\bibitem[PCI-SIG(2025)]{pcie7}
PCI-SIG.
\newblock \href{https://pcisig.com/}{PCI Express 7.0 Specification} and Optical Workgroup interconnect revision.
\newblock Released 11~June~2025: 128~GT/s per lane, PAM4, flit encoding
($\sim$512~GB/s bidirectional on a x16 link). The PCI-SIG Optical Workgroup (2023)
targets a technology-agnostic optical PCIe interface for reach beyond copper.
Provisional: 2025 release; rates and optical-interface scope pending cross-check
against the final published specification.

\bibitem[CXL(2025)]{cxl4}
CXL Consortium.
\newblock \href{https://computeexpresslink.org/}{Compute Express Link (CXL) 4.0 Specification}.
\newblock Released 18~November~2025 on a PCIe 7.0 base (128~GT/s, PAM4); coherent
CPU/accelerator/memory fabric with memory pooling and bundled ports. Relevant to
in-rack memory disaggregation and optical CXL. Provisional: 2025 release; verify
details against the final published specification.

\bibitem[Synopsys 448G(2025)]{synopsys448}
Synopsys.
\newblock \href{https://www.synopsys.com/content/dam/synopsys/white-papers/designing-for-448g-modulation-dsp-wp.pdf}{Designing for 448G: Modulation, DSP, and Channel Feasibility}.
\newblock White paper, 2025.
\newblock PAM4 (224~GBd, 112~GHz Nyquist) vs.\ PAM6 (179~GBd, 90~GHz); dual-mode
448G SerDes channel simulations on 224G-class links.

\bibitem[OIF 448G Workshop(2025)]{huawei448}
OIF 448G AI Workshop (Lumentum/Huawei et al.).
\newblock \href{https://www.oiforum.com/wp-content/uploads/OIF-448G-AI-Workshop-Huawei-Kuschnerov_Matuz.pdf}{448G/lane native modulation and gearbox transition}.
\newblock 2025.
\newblock First 448G optical modules may gear-box 224G SerDes; electrical/optical
format alignment preferred for LPO and retimed architectures.

\bibitem[Cignal AI(2025)]{cignal448}
Cignal AI.
\newblock \href{https://cignal.ai/2025/04/oifs-448g-workshop-the-need-for-speed/}{OIF's 448G Workshop: The Need for Speed}.
\newblock April 2025.
\newblock Industry coalescing on PAM6 for electrical chip-to-module; PAM4 at
224~GBd for optics; OFC 2025 400G PAM4 optical demos.

\bibitem[St-Arnault et al.(2025)]{tfmln32t}
C.~St-Arnault \emph{et al.}
\newblock \href{https://arxiv.org/html/2503.24147}{Net 3.2~Tbps 225~Gbaud PAM4 O-band IM/DD 2~km transmission with CMOS 3~nm SerDes and TFLN modulators}.
\newblock arXiv:2503.24147, 2025 (preprint).
\newblock Eight $\times$225~GBd PAM4 lanes; FR8/DR8; HD-FEC threshold.

\bibitem[Liu et al.(2025)]{tfln224}
Y.~Liu \emph{et al.}
\newblock \href{https://arxiv.org/abs/2311.05119}{High-speed thin-film lithium niobate modulator with transparent conductive oxide electrodes}.
\newblock arXiv:2311.05119; OFC 2025, paper M3K.2.
\newblock 108~GHz EO BW; 224~Gb/s PAM4 O-band; $V_\pi L=1.02$~V$\cdot$cm.

\bibitem[He et al.(2025)]{tfln390}
M.~He \emph{et al.}
\newblock \href{https://arxiv.org/abs/2411.15037}{Dual-band thin-film lithium niobate modulators for 390~Gb/s PAM8 transmission}.
\newblock \emph{Laser \& Photonics Reviews}, 2025; arXiv:2411.15037.
\newblock Extrapolated 220~GHz EO BW; $V_\pi L\approx1.08$--1.33~V$\cdot$cm; sub-fJ/bit lab demo.

\bibitem[OFC 2026 W1A.3]{tfln200mm}
OFC 2026, paper W1A.3.
\newblock \href{https://doi.org/10.1364/OFC.2026.W1A.3}{200-mm silicon-compatible TFLN platform}.
\newblock 110~GHz EO modulators; $V_\pi L=1.9$~V$\cdot$cm; foundry-scale integration path.

\bibitem[He et al.(2019)]{he2019}
M.~He \emph{et al.}
\newblock \href{https://doi.org/10.1038/s41566-019-0378-6}{High-performance hybrid silicon and lithium niobate Mach--Zehnder modulators}.
\newblock \emph{Nature Photonics} \textbf{13}, 359--364 (2019).
\newblock Hybrid Si--LN integration template for co-packaged assemblies.

\bibitem[HyperLight(2025)]{hyperlight110}
HyperLight Corp.
\newblock \href{https://www.hyperlightcorp.com/news/hyperlight-announces-110-ghz-modulators}{110-GHz packaged TFLN IQ modulators for 240-GBaud-class signaling}.
\newblock September 2025 (vendor announcement; provisional).

\bibitem[MACOM(2026)]{macom448}
MACOM Technology Solutions.
\newblock \href{https://www.macom.com/updates/news/2026/macom-announces-two-new-448g-per-lane-drivers-for-3-2t-data-cent}{448G per-lane PAM4 modulator drivers (MAOM-025408 MZM, MAOM-022404 EML)}.
\newblock March 2026 (OFC).
\newblock $>$120~GHz RF bandwidth; SiPh, EML, and TFLN platforms; wire-bond or bumped die.

\bibitem[Semtech(2026)]{semtech224}
Semtech Corporation.
\newblock \href{https://www.semtech.com/company/press/semtech-launches-224-gbps-ic-family-for-linear-optics-era}{224~Gb/s/lane MZM drivers (GN1877/GN1887) and TIA family for LPO/LRO/CPO}.
\newblock March 2026 (OFC); CEI-224G-Linear / LPO-MSA positioning (vendor announcement).

\bibitem[Da Silva et al.(2026)]{rfic460drv}
L.~Da~Silva \emph{et al.}
\newblock \href{https://rfic-ieee.org/paper_abstract/1148}{A 460~Gb/s PAM-4 linear distributed driver with 105~GHz BW for TFLN modulators in 130~nm SiGe BiCMOS}.
\newblock RFIC 2026: 105.7~GHz BW, 2.25~$V_{\mathrm{pp}}$, 232~GBd PAM4 with offline DSP (research).

\bibitem[OFC W4J.4(2026)]{ofc420tosa}
OFC 2026, paper W4J.4.
\newblock \href{https://doi.org/10.1364/OFC.2026.W4J.4}{A 420~Gb/s/lane O-band PAM-4 TOSA based on thin-film lithium niobate}.
\newblock Co-designed TFLN modulator + driver; 210~GBd / 420~Gb/s PAM4.

\bibitem[OFC W3E.6(2026)]{ofc180drv}
OFC 2026, paper W3E.6.
\newblock \href{https://doi.org/10.1364/OFC.2026.W3E.6}{180~GBaud PAM4 driver--modulator engine for IM/DD in the O-band}.
\newblock Co-packaged 76~GHz InP MZM with 224~GBd-class linear differential EML driver.

\bibitem[Aimone et al.(2026)]{tflndiffdrv}
A.~Aimone \emph{et al.}
\newblock \href{https://doi.org/10.1364/OFC.2026.M4D.4}{Truly-differential drive of TFLN TWE-MZM by linear SiGe driver in a codesigned hybrid integrated assembly}.
\newblock OFC 2026, M4D.4: $V_\pi L=1.1$~V$\cdot$cm; 140~GBd PAM8; 1.4~pJ/bit.

\bibitem[Lin et al.(2026)]{ring224}
M.-W.~Lin \emph{et al.}
\newblock \href{https://doi.org/10.1364/OFC.2026.Th2A.14}{Record-high 90-GHz silicon microring modulator and 224-Gb/s PAM4 operation}.
\newblock OFC 2026, paper Th2A.14.
\newblock Inductive and wavelength tuning; compact RLC model; 90-GHz EO BW; CPO target.

\bibitem[Peng et al.(2026)]{ring400}
C.-W.~Peng \emph{et al.}
\newblock \href{https://doi.org/10.1364/OFC.2026.M2A.2}{Ultra-high bandwidth silicon microring modulator with T-coil peaking for $>$400~Gbps transmission}.
\newblock OFC 2026, paper M2A.2.
\newblock $>$110~GHz EO BW; 416~Gb/s PAM4, TDECQ 2.88~dB at 1~Vpp.

\bibitem[OFC M2A.1(2026)]{ring256}
OFC 2026, paper M2A.1.
\newblock \href{https://doi.org/10.1364/OFC.2026.M2A.1}{256~Gb/s silicon Euler microring modulator with 3~THz FSR and $>$67~GHz bandwidth}.
\newblock O-band Euler ring; inductive peaking; 256~Gb/s PAM4.

\bibitem[Dong et al.(2026)]{simzm400}
P.~Dong \emph{et al.}
\newblock \href{https://doi.org/10.1364/OFC.2026.Th4A.4}{400G/lane PAM4 modulation using silicon Mach--Zehnder modulators}.
\newblock OFC 2026 postdeadline, paper Th4A.4.
\newblock Si MZM with commercial SiGe driver (2.5~V swing); 400~Gb/s/lane PAM4 IM/DD.

\bibitem[OFC M2A.6(2026)]{simzmcompact}
OFC 2026, paper M2A.6.
\newblock \href{https://doi.org/10.1364/OFC.2026.M2A.6}{Compact Mach--Zehnder modulator in 300-mm silicon photonic platform toward 400Gbps/lane}.
\newblock 500~$\mu$m MZM; 94.7~GHz median EO BW; 2.4~dB median IL at 1310~nm.

\bibitem[OFC W3E.5(2026)]{simzmdiff}
OFC 2026, paper W3E.5.
\newblock \href{https://doi.org/10.1364/OFC.2026.W3E.5}{Toward 400~G/lane silicon differential-drive Mach--Zehnder modulator with $>$80~GHz bandwidth}.
\newblock Differential-drive Si MZM; 81.8~GHz 3-dB EO BW; 100~GBd PAM8 eyes (lab setup).

\bibitem[CW-WDM MSA(2021)]{cwwdm}
CW-WDM MSA.
\newblock \emph{\href{https://cw-wdm.org/}{Continuous-Wave WDM Multi-Source Agreement}}
Technical Specifications Rev~1.0 (4~June~2021).
\newblock O-band grids: 8/16/32-line sets in 9/18/36~nm spans; modular (one
$\lambda$/fiber) and integrated (full comb/fiber) configs; measurement methods for
power, RIN, SMSR, linewidth. Form factor and management left to suppliers /
ELSFP. Informative 18~nm examples often cite $\approx$30~dB SMSR,
$\approx-135$~dB/Hz RIN, $\approx$20~MHz linewidth.

\bibitem[OIF ELSFP(2023)]{elsfp}
Optical Internetworking Forum.
\newblock \href{https://www.oiforum.com/wp-content/uploads/OIF-ELSFP-02.0.pdf}{External Laser Small Form-Factor Pluggable (ELSFP) Implementation Agreement}.
\newblock OIF-ELSFP-02.0, 2023.
\newblock 24-pin card-edge electrical interface; CMIS management; blind-mate MT optical connector; optical/thermal/electrical power classes.

\bibitem[Ayar Labs(2026)]{ayar}
Ayar Labs.
\newblock \emph{\href{https://ayarlabs.com/supernova/}{SuperNova light source and TeraPHY optical I/O chiplet}}.
\newblock ayarlabs.com.

\bibitem[Arista/XPO MSA(2026)]{xpo}
Arista Networks (with Coherent, Marvell, Lightmatter, et al.).
\newblock \emph{XPO: eXtra-dense Pluggable Optics multi-source agreement}
(12.8~Tb/s liquid-cooled pluggable).
\newblock Announced OFC 2026; arista.com white paper, March 2026.

\bibitem[Säckinger(2018)]{sackinger2018}
E.~Säckinger.
\newblock \citehref{sackinger2018}{\emph{Analysis and Design of Transimpedance Amplifiers for Optical
Receivers}}.
\newblock Wiley, 2018 (Chapter 4: Receiver Fundamentals).

\bibitem[100G Lambda MSA / IEEE 802.3]{rinspec}
100G Lambda MSA and IEEE 802.3.
\newblock \href{https://www.ieee802.org/3/bs/public/}{\emph{400G-FR4 and 100GBASE-BR technical specifications}} ($\mathrm{RIN}_{17.1}\mathrm{OMA}\le-136$~dB/Hz; RIN method per IEEE Std 802.3 clause 52.9.6).

\bibitem[100G Lambda MSA(2017--20)]{lambdamsa}
100G Lambda MSA Group.
\newblock \emph{\href{https://100glambda.com/}{100G Lambda MSA specifications}} (100G-FR/LR, 400G-DR4/FR4/LR4).
\newblock Formed 2017 (22 promoter members). Specifications built on 100~Gb/s per
wavelength PAM4 ($\approx$53~GBd) over duplex single-mode fiber, 2--40~km reaches;
defined the single-$\lambda$ 100G approach IEEE 802.3 adopted into its DR/FR/LR PMD
clauses and that the LPO MSA 100G-DR-LPO profile inherits, including the
$\mathrm{RIN}_x\mathrm{OMA}$ transmitter method.

\bibitem[Koheron DRV200]{koherondrv200}
Koheron.
\newblock \href{https://www.koheron.com/}{\emph{DRV200 low-noise laser diode driver datasheet}}
(current noise density at 1~kHz: 55 / 270 / 480~pA$/\sqrt{\mathrm{Hz}}$ for 40 / 200 / 400~mA variants).

\bibitem[QD-on-Si RIN(2020--2025)]{qdrin}
Intensity-noise studies of epitaxial/heterogeneous quantum-dot lasers on silicon.
\newblock \emph{Opt. Lett.} 45(17):4887 (2020, p-doped QD RIN $\approx-150$~dB/Hz);
\emph{J. Photonics} (2024, quiet-pump/injection-locked RIN to $\sim-168$~dB/Hz);
\emph{JLT} 43(4):1855 (2025, feedback-tolerant QD on 300~mm Si).

\bibitem[RIO PLANEX]{eclplanex}
RIO Lasers.
\newblock \href{https://www.lasercomponents.com/fileadmin/user_upload/home/Datasheets/rio-laser/orion-planex-external-cavity-laser.pdf}{\emph{PLANEX 1550~nm external-cavity laser datasheet}}.
\newblock Planar gain-chip and Bragg-grating external cavity with narrow linewidth,
low phase noise, and wavelength stability; sensing and metrology product rather
than a short-reach IM/DD default. Vendor orientation, provisional.

\bibitem[Ge/Si PD(2025)]{gesipd}
Waveguide germanium-on-silicon photodiodes for 200~GBd and beyond.
\newblock Shahin et al., \emph{OECC/PSC} (2025), recessed Ge/Si PIN: 106~GHz, $0.93$~A/W, $<\!10$~nA dark;
\newblock \emph{ACS Photonics} (2025), SiN-coupled lateral Ge PD $>\!110$~GHz at 1~mA (O-band).

\bibitem[Krishnamoorthy et al.(2022)]{tia112}
A 112-Gb/s $-8.2$-dBm-sensitivity 4-PAM linear TIA in 16-nm CMOS with co-packaged photodiodes.
\newblock \emph{IEEE J. Solid-State Circuits} 58(4), 2023 (32~GHz BW, $16.9$~pA/$\sqrt{\text{Hz}}$).

\bibitem[224G SiGe TIA(2026)]{tia224}
A $4\times112$-GBaud linear PAM-4 TIA with 65-GHz bandwidth and $13.2$-pA/$\sqrt{\text{Hz}}$ input noise for 1.6-T links.
\newblock \emph{IEEE Trans. Microw. Theory Tech.}, 2026 (55-nm SiGe BiCMOS, 224~Gb/s, 1.22~pJ/bit).

\bibitem[UMC-APD(2025)]{apd}
\href{https://doi.org/10.1038/s41467-025-66047-6}{Ultrafast avalanche photodiode exceeding 100~GHz bandwidth}.
\newblock \emph{Nat.\ Commun.} (2025): Ge/Si UMC-APD, 105~GHz at gain~$\approx$7; GBP~$\sim$4800~GHz;
224/260~Gb/s PAM4 at $-10.9$/$-10.1$~dBm ($\sim$9~dB over PIN).

\bibitem[Shahin et al.(2026)]{gesiapd100}
A.~Shahin \emph{et al.}
\newblock \href{https://doi.org/10.1109/JLT.2026.3670103}{Towards 100~GHz waveguide Ge/Si avalanche photodiodes for high-speed optical interconnects}.
\newblock \emph{JLT} 2026: $>$100~GHz peak BW; 70~GHz at 2~A/W; $\sim$7~V; 300~mm SOI.

\bibitem[OFC Th3F.2(2026)]{ofc360apd}
OFC 2026, paper Th3F.2.
\newblock \href{https://doi.org/10.1364/OFC.2026.Th3F.2}{360~Gbps Ge-on-Si avalanche photodiodes operating in the O- and C-band}.
\newblock Up to 180~GBd PAM4 below HD-FEC; 70/100~GHz BW; 2/1.5~A/W.

\bibitem[Semtech OFC demos(2026)]{semtech448demo}
Semtech Corporation.
\newblock \href{https://www.semiconductor-today.com/news_items/2026/mar/semtech-160326.shtml}{OFC 2026: 224G optical demos and 448G/lane PMD ICs (TN622 driver, TN14740 TIA)}.
\newblock Vendor demonstration; provisional until datasheet limits are published.

\bibitem[UTC-PD(2023)]{utc}
\href{https://arxiv.org/abs/2501.02812}{Wideband uni-traveling-carrier photodiodes for near-300~Gbps communications}.
\newblock Type-II GaInAsSb/InP UTC-PD $>\!110$~GHz; InP MUTC-PD 206~GHz (arXiv:2501.02812, 2025).

\bibitem[OIF CEI-224G(2024--25)]{cei224}
Optical Internetworking Forum, \href{https://www.oiforum.com/technical-work/hot-topics/common-electrical-i-o-cei-224g/}{Common Electrical I/O (CEI-224G) framework and
implementation agreements}; CEI interoperability demos, OFC/ECOC 2025.
\newblock Reach classes: XSR (die-to-die), VSR (200~mm host + module), MR
($\sim$500~mm, $\sim$32--34~dB), LR ($\sim$1~m, up to $\sim$40~dB die-to-die).
CEI-224G-Linear extends the CEI-112G-Linear methodology to 224G full-linear
modules (LPO/CPO/NPO): TP1/TP1a and TP4/TP4a electrical specs without module
DSP/CDR.

\bibitem[OIF CEI Demo OFC(2025)]{cei224demo}
Optical Internetworking Forum,
\emph{\href{https://www.oiforum.com/wp-content/uploads/OIF_CEI_Demo_OFC2025.pdf}{448G, 224G, 112G CEI Interoperability Demo}},
OFC 2025.
\newblock Project map: CEI-224G-XSR ($\lesssim$50~mm), VSR (200+20~mm), MR
($\sim$500~mm), LR ($\sim$1000~mm), and CEI-224G-Linear (no module DSP);
BER $10^{-15}$ with FEC allowed. Live Linear/RTLR/VSR/LR demos.

\bibitem[IEEE 802.3df 224G PCB(2022)]{traceloss}
M.~Li et al., \emph{\href{https://www.ieee802.org/3/df/public/22_07/li_3df_01a_2207.pdf}{224G Package and PCB Investigations and COM Reference Model}},
IEEE 802.3df, 2022.
\newblock M7N stripline loss at 56~GHz: $\approx$2.8~dB/inch (regular),
$\approx$1.9~dB/inch (skip-layer); next-gen target 1~dB/inch at Nyquist.

\bibitem[COBO / OBO--NPO--CPO]{cobo}
Consortium for On-Board Optics, \emph{On-Board Optical Module Specification}; and
comparative analyses of pluggable, on-board (OBO), near-packaged (NPO), and
co-packaged (CPO) optics architectures.

\bibitem[Synopsys 224G DAC(2024)]{dac224}
Reliable 2~m+ passive DAC connectivity for AI-scale networks with 224G PHY IP
(BER $\sim$2\,$\times$10$^{-9}$ at 2~m); passive DAC $\sim$0.5--1~m nominal, AEC to
$\sim$2.5~m at 224G.

\bibitem[Roofline(2009)]{roofline}
S.~Williams, A.~Waterman, and D.~Patterson.
\newblock \emph{\href{https://doi.org/10.1145/1498765.1498785}{Roofline: An Insightful Visual Performance Model for Multicore
Architectures}}. Communications of the ACM, 2009.
\newblock Arithmetic intensity (FLOP/byte) vs.\ the compute/bandwidth balance
point; the framework behind the decode memory-bound argument.

\bibitem[LPO/LRO(2025)]{lpo}
\href{https://eps.ieee.org/wp-content/uploads/2026/03/Linear-Pluggable-Optics_V2-UPDATED.pdf}{Linear Pluggable Optics (LPO) and Linear-Receive Optics (LRO/TRO)}: DSP-less and
half-retimed module architectures.
\newblock IEEE EPS overview (2026); OIF/MSA interop demos, OFC 2025. Typical 800G:
DSP module $\sim$14--18~W, $\sim$8--10~ns/hop; LPO $\sim$7--9~W, $<$3~ns, $<$2~km;
LRO $\sim$9~W. Redriver = CTLE$+$VGA (no CDR); retimer = EQ$+$CDR.
Names component proponents (Macom, Semtech, Maxlinear) and OFC~2024--25 demos
from Eoptolink, Macom, Marvell, Alphawave, and Innolight.

\bibitem[LBNL(2024)]{lbnl2024}
A.~Shehabi et al.
\newblock \emph{\href{https://eta-publications.lbl.gov/sites/default/files/2024-12/lbnl-2024-united-states-data-center-energy-usage-report.pdf}{2024
United States Data Center Energy Usage Report}}. Lawrence Berkeley National
Laboratory, December 2024.
\newblock US data centers: 58~TWh (2014) $\to$ 176~TWh (2023, $\sim$4.4\% of US
electricity) $\to$ projected 325--580~TWh (6.7--12\%) by 2028.

\bibitem[Cignal AI(2023)]{cignal}
Cignal AI, \emph{\href{https://cignal.ai/2023/08/linear-drive-market-opportunity/}{The
Linear Drive Market Opportunity}} (2023) and OFC~2025 show report.
\newblock Market view: LPO stays a small share, especially at 800G, because the
installed fabric is DSP-based; 100G/lane 800GbE LPO late to market; at 1.6T,
LRO favored over LPO on power/thermal. Provisional (analyst reports).

\bibitem[Eoptolink(2024)]{eoptolinklpo}
Eoptolink, \emph{\href{https://www.eoptolink.com/news/13-new-products/348-eoptolink-demonstrates-industry-1st-200g-lane-lpos-with-100g-lane-800g-lpos-entering-mass-production}{Industry-first
200G-per-lane LPO; 100G/lane 800G LPO entering mass production}}, March 2024.
\newblock 200G/$\lambda$ four-channel LPO, no DSP/CDR; Gen2 100G/lane 800G and
400G SMF (OSFP, QSFP-DD, QSFP112), full TP2 compliance claimed. Vendor,
provisional.

\bibitem[Macom(2024)]{macomlpo}
Macom, \emph{\href{https://ir.macom.com/news-releases/news-release-details/macom-showcase-200g-lane-products-optical-fiber-communication}{Macom
to showcase 200G-per-lane products at OFC}}, March 2024.
\newblock PURE~DRIVE 200~Gb/s LPO, extensible to 212~Gb/s/lane for 1.6T; linear
TIA and driver as the key blocks. Vendor, provisional.

\bibitem[Marvell(2024)]{marvelllpo}
Marvell, \emph{\href{https://www.marvell.com/company/newsroom/marvell-introduces-1-6-tbps-lpo-chipset.html}{Marvell
introduces 1.6~Tbps LPO chipset}}, December 2024.
\newblock 200G/lane TIA and laser-driver chipset for 800G and 1.6T LPO, targeting
scale-up XPU compute fabrics. Vendor, provisional.

\bibitem[Alphawave/InnoLight(2024)]{alphawavelpo}
Alphawave Semi and InnoLight, \emph{\href{https://awavesemi.com/press-release/alphawave-semi-and-innolight-collaborate-to-demonstrate-low-latency-linear-pluggable-optics-with-pcie-6-0-subsystem-solution-for-high-performance-ai-infrastructure-at-ofc-2024/}{Low-latency
LPO with a PCIe~6.0 subsystem for AI infrastructure}}, March 2024.
\newblock 64~Gb/s/lane PCIe~6.0 controller+PHY over InnoLight LPO OSFP optics
(OFC~2024); later a 128~Gb/s/lane PCIe~7.0-ready platform. Vendor, provisional.

\bibitem[LPO MSA(2025)]{lpomsa}
LPO MSA, \emph{\href{https://www.lpo-msa.org/files/live/sites/lpomsa/files/specs/LPO_MSA_Specification_v1p01.pdf}{100G-DR-LPO Specification}},
rev.~1.0, 19~March 2025.
\newblock 100~Gb/s/lane PAM4 SMF to 500~m; builds on IEEE 802.3 and OIF
CEI-112G-LINEAR. TDECQ/TECQ $\le$3.4~dB; OMA coupled to max(TECQ,TDECQ); host
RS(544,514) FEC. Template for linear-module optical limits; 224G linear modules
follow CEI-224G-Linear electrically while optical PMDs track 802.3dj.

\bibitem[Telcordia(2004)]{gr468}
Telcordia.
\newblock \emph{GR-468-CORE: Generic Reliability Assurance Requirements for
Optoelectronic Devices}.

\bibitem[Telcordia(1999)]{gr1221}
Telcordia.
\newblock \emph{GR-1221-CORE: Generic Reliability Assurance Requirements for
Passive Optical Components}, Issue~2 (January 1999).
\newblock Passive-component companion to GR-468: connectors, couplers, WDM
filters, splitters, and isolators. Same stress families (damp heat, temperature
cycle, mechanical, aging) scored on insertion and return loss rather than LIV.

\bibitem[Broadcom(2025)]{th6davisson}
Broadcom Inc.
\newblock \emph{Tomahawk 6 -- Davisson: 102.4~Tb/s co-packaged-optics Ethernet
switch} (press release and briefs).
\newblock October 2025.

\bibitem[NVIDIA(2025)]{nvphotonics}
NVIDIA.
\newblock \emph{Spectrum-X and Quantum-X silicon photonics co-packaged-optics
switches} (GTC 2025; Hot Chips 2025, Shainer).

\bibitem[TSMC(2026)]{coupe}
TSMC.
\newblock \emph{Compact Universal Photonic Engine (COUPE): SoIC-X electronic/
photonic integration; mass-production milestone}.
\newblock 2026.

\bibitem[OIF(2020)]{oif400zr}
Optical Internetworking Forum.
\newblock \href{https://www.oiforum.com/technical-work/400zr/}{OIF Implementation Agreement for 400ZR}.
\newblock 400~Gb/s coherent DWDM pluggable; 15~W module power budget; DCI reach to
$\sim$120~km; interoperable QSFP-DD/OSFP coherent optics.

\bibitem[OIF CMIS(2024)]{cmis53}
Optical Internetworking Forum.
\newblock \href{https://www.oiforum.com/wp-content/uploads/OIF-CMIS-05.3.pdf}{CMIS 5.3 Implementation Agreement} (OIF-CMIS-05.3).
\newblock Host-to-module management over two-wire bus; module and data-path state
machines; DDM, VDM, CDB; link training; ELSFP and CPO extensions.

\bibitem[Redfish/OpenConfig/SONiC(2015--25)]{netmgmt}
DMTF, OpenConfig, and the SONiC community.
\newblock \href{https://www.dmtf.org/standards/redfish}{DMTF Redfish (DSP0266)},
\href{https://www.openconfig.net/}{OpenConfig models and gNMI}, and
\href{https://sonicfoundation.dev/}{SONiC}.
\newblock Box- and fleet-level management and telemetry above OIF CMIS. Redfish:
RESTful server/switch management API (v1.0, August 2015). OpenConfig:
vendor-neutral YANG models and streaming telemetry. SONiC: open network OS with
dataplane telemetry. Where per-module optical monitors surface at cluster scale.

\bibitem[IEC(2021)]{iec60825}
IEC.
\newblock \href{https://webstore.iec.ch/en/publication/74712}{IEC 60825-1:2014/AMD1:2021} and
\href{https://webstore.iec.ch/en/publication/66912}{IEC 60825-2:2021}.
\newblock Laser product classes (1, 1M, 3R, 3B, 4) and OFCS hazard levels at
accessible fiber ports.

\bibitem[ITU-T(2012)]{g664}
ITU-T.
\newblock \href{https://www.itu.int/rec/T-REC-G.664-201202-I/en}{Recommendation ITU-T G.664 (02/2012)}.
\newblock Optical safety procedures; APR with automatic restart; reduce to Hazard
Level 1M within 3~s of continuity loss.

\bibitem[ITU-T(2003--20)]{ituG694}
ITU-T.
\newblock \href{https://www.itu.int/rec/T-REC-G.694.1/en}{Recommendation ITU-T G.694.1 (10/2020), Spectral grids for WDM applications: DWDM frequency grid}
and \href{https://www.itu.int/rec/T-REC-G.694.2/en}{ITU-T G.694.2 (12/2003), CWDM wavelength grid}.
\newblock G.694.1 anchors the DWDM grid at 193.1~THz with 12.5/25/50/100~GHz
spacings and integer multiples; G.694.2 defines the 20~nm CWDM grid, 18 channels
from 1271 to 1611~nm. Datacom DWDM locks to G.694.1; CWDM4 uses four O-band lines
of G.694.2.

\bibitem[ITU-T(2016--24)]{itufiber}
ITU-T.
\newblock \href{https://www.itu.int/rec/T-REC-G.652/en}{Recommendation ITU-T G.652 (11/2016), Characteristics of a single-mode optical fibre and cable}
and \href{https://www.itu.int/rec/T-REC-G.657/en}{ITU-T G.657 (08/2024), Characteristics of a bending-loss insensitive single-mode optical fibre and cable}.
\newblock G.652.D sets the standard single-mode fiber used in short-reach links:
attenuation $\approx$0.2~dB/km at 1550~nm, $\approx$0.32~dB/km at 1310~nm, and
zero-dispersion wavelength near 1310~nm. G.657 defines bend-insensitive fiber for
tight-radius routing in patch panels and shelves; category A stays G.652-compatible,
category B trades some standard-fiber interoperability for tighter bend radii.

\bibitem[ISO/IEC and TIA(2017--24)]{fibercabling}
ISO/IEC and TIA.
\newblock \href{https://www.iso.org/standard/66763.html}{ISO/IEC 11801:2017, Information technology: Generic cabling for customer premises}
and the \href{https://www.tiaonline.org/}{TIA-492 series, Detail Specifications for Optical Fibers}.
\newblock ISO/IEC 11801 defines the cabled fiber performance classes used in
datacenter patch plant: OM1--OM5 for multimode and OS1/OS2 for single-mode. TIA-492
gives the matching detailed fiber specifications (for example TIA-492AAAD
corresponds to OM4); TIA-568 adopts the same OM/OS class names for structured
cabling.

\bibitem[JEDEC(2025)]{jesd47}
JEDEC Solid State Technology Association.
\newblock \href{https://www.jedec.org/standards-documents/docs/jesd-47g}{JESD47M, Stress-Test-Driven Qualification of Integrated Circuits}.
\newblock Published August 2025. Baseline acceptance-test flow (temperature cycle,
HTOL/HTSL, autoclave/HAST, mechanical shock and vibration) for qualifying an IC as
a new product, a product family, or after a process change. The silicon-side
counterpart to Telcordia GR-468-CORE for optoelectronics.

\bibitem[JEDEC/ESDA(2024--25)]{jsesd}
JEDEC Solid State Technology Association and ESD Association.
\newblock \href{https://www.jedec.org/standards-documents/docs/js-001-2023}{ANSI/ESDA/JEDEC JS-001-2024, Human Body Model (HBM) ESD Sensitivity Testing}
and \href{https://www.jedec.org/standards-documents/docs/js002-2014}{JS-002-2025, Charged Device Model (CDM) ESD Sensitivity Testing}.
\newblock Component-level ESD classification for driver, TIA, and DSP silicon.
Supersede the older JEDEC-only JESD22-A114 (HBM) and JESD22-C101 (CDM) test methods.

\bibitem[JEDEC(2023)]{jesd78}
JEDEC Solid State Technology Association.
\newblock \href{https://www.jedec.org/standards-documents/docs/jesd-78b}{JESD78F.02, IC Latch-Up Test}.
\newblock Published November 2023. Overvoltage and $\pm100$~mA current-injection
test method for parasitic-thyristor latch-up susceptibility in CMOS, bipolar, and
BiCMOS ICs; the failure mode COD is to a laser facet.

\bibitem[AEC(2023--25)]{aecq100}
Automotive Electronics Council.
\newblock \href{http://aecouncil.com/}{AEC-Q100 Rev-J, Failure Mechanism Based Stress Test Qualification for Integrated Circuits}.
\newblock Automotive-grade IC qualification built on the JEDEC JESD47/JESD22
methods with tightened ESD targets and temperature grades (Grade~3:
$-40$ to $85$\textdegree C through Grade~0: $-40$ to $150$\textdegree C).
Not required for datacenter optics, but a useful borrow when a driver or TIA die
also ships in an automotive part number and the datasheet cites a Q100 grade.

\bibitem[IEC(2014--17)]{iec617547}
IEC.
\newblock \href{https://webstore.ansi.org/standards/iec/iec61754eden2014}{IEC 61754-7-1:2014, Fibre optic connector interfaces -- Type MPO connector family: one fibre row}
and \href{https://webstore.iec.ch/en/publication/30239}{IEC 61754-7-2:2017, Type MPO connector family: two fibre rows}.
\newblock Mechanical interface standard for the MPO/MT connector family: 6.4~mm
$\times$ 2.5~mm rectangular ferrule, guide-pin alignment, 8/12/16/24-fiber counts.
Superseded the single-part IEC 61754-7:2008 to add active-device receptacles and
up-angled plugs.

\bibitem[IEC(2009--22)]{iec61300}
IEC.
\newblock \href{https://webstore.ansi.org/standards/iec/iec61300ed2009-1296572}{IEC 61300-2-2:2009, Basic test and measurement procedures -- Mating durability}
and \href{https://webstore.iec.ch/en/publication/64254}{IEC 61300-3-35:2022, Visual inspection of fibre optic connectors and fibre-stub transceivers}.
\newblock IEC 61300-2-2 defines the mate/unmate cycling test method that connector
datasheets rate against; TIA-568.3 sets 500 cycles as the structured-cabling
durability floor, and MPO/MTP-class connectors are commonly rated well above
1000. IEC 61300-3-35 grades endface scratches, pits, and debris against zones on
the fiber core and cladding for pass/fail inspection.

\bibitem[RS-FEC latency(2024)]{feclatency}
Concatenated RS-BCH coded modulation latency study (incl.\ KP4).
\newblock \href{https://arxiv.org/abs/2402.09364}{Performance-complexity-latency trade-offs of concatenated codes}, arXiv:2402.09364, 2024; and vendor RS(544,514) implementation notes.
\newblock KP4 RS(544,514) decode latency in practice $\approx$20--100~ns depending
on implementation and clock; encode is comparable. Academic/vendor figures,
provisional.

\bibitem[AI switch latency(2025--26)]{switchlat}
Vendor switch datasheets (Ethernet and InfiniBand).
\newblock \href{https://cloudswit.ch/product/800gbe-switch-512tbps-enterprise-sonic/}{800GbE 51.2~Tb/s SONiC switch, $\sim$560~ns port-to-port}; NVIDIA Quantum-X800 InfiniBand advertises sub-100~ns port-to-port with in-network compute.
\newblock Per-hop O-E-O switch latency: a few hundred ns for cut-through Ethernet,
under $\sim$100~ns for InfiniBand. Vendor orientation, provisional.

\bibitem[Chen et al.(2025)]{hcf2025}
Y.~Chen, M.~N.~Petrovich, E.~Numkam~Fokoua, \emph{et al.} (Southampton/Microsoft).
\newblock \href{https://doi.org/10.1038/s41566-025-01633-0}{Record-low-loss hollow-core fibre} (double nested antiresonant nodeless fibre, DNANF).
\newblock \emph{Nature Photonics}, September 2025: 0.091~dB/km at 1550~nm (below the
$\sim$0.14~dB/km silica floor), $\approx$0.2~dB/km over a 66~THz window; air guiding
gives group index near 1.0, $\sim$45\% faster propagation (about a third lower
latency) than silica. Microsoft reports Azure network deployment. Record result,
provisional.

\bibitem[QD comb(2022--2025)]{qdcomb}
Quantum-dot mode-locked laser (QD-MLL) frequency combs for datacom.
\newblock \href{https://www.mdpi.com/2076-3417/12/4/1836}{High-efficiency QD lasers as comb sources for DWDM}, \emph{Appl.\ Sci.} 12(4):1836 (2022): among DFB arrays, nonlinear microcombs, and semiconductor MLLs, QD-MLLs currently offer the best efficiency, simplicity, and size for a single-device comb.
\newblock O-band IM/DD demos: a 1.3~\textmu m InAs/InGaAs QD-MLL comb carrying 14$\times$100~Gb/s PAM4 over 10~km SMF at $\sim$284~fJ/bit; and an isolator-free QD-MLL comb reported for interconnects beyond 3.2~Tb/s (\href{https://arxiv.org/abs/2506.02402}{arXiv:2506.02402}, 2025). Research demos, provisional.

\bibitem[Kerr microcomb source(2025)]{microcomb}
Integrated Kerr/soliton microcomb multi-wavelength sources.
\newblock \href{https://doi.org/10.1038/s41467-025-61288-x}{Integrated multi-port multi-wavelength coherent optical source}, \emph{Nat.\ Commun.} 16 (2025): a Kerr microcomb with a monolithically integrated demultiplexer that autonomously locks to and tracks the comb lines, aimed at beyond-Tb/s links.
\newblock A microcomb turns one pump into hundreds of evenly spaced lines on a chip, but per-line power is low (pump-conversion efficiency is modest), so a booster or per-line SOA is usually required (\Cref{sec:soa-distribution}). Research demos, provisional.

\bibitem[O-band SOA(2019--2023)]{soa}
Semiconductor optical amplifiers for datacom booster/distribution use.
\newblock Commercial O-band SOA modules quote $\sim$15~dB gain, noise figure $\le$7~dB, and $\le$1.5~dB polarization-dependent gain (\href{https://www.anritsu.com/en-US/sensing-devices/products/soa}{Anritsu SOA datasheet}); the quantum NF floor is 3~dB. \href{https://pubs.acs.org/doi/10.1021/acsphotonics.9b00903}{O-band quantum-dot SOAs grown on a CMOS-compatible silicon substrate} (\emph{ACS Photonics}, 2019) give wide O-band gain with low noise for co-packaged integration.
\newblock An SOA restores launch power after comb generation and fan-out, at the cost of added ASE. Vendor/academic figures.

\bibitem[CPC/NPC(2025--2026)]{cpc}
Co-packaged and near-package copper, copper's answer to the same reach wall that drives CPO.
\newblock \href{https://www.molex.com/content/molex/molex-dot-com/language-masters/en/industries-applications/data-center-connector-solutions/near-asic/co-packaged-copper.html}{Molex co-packaged copper (CPC)}: on-substrate routing validated at 224~Gb/s PAM4 with a stated roadmap to 448G, aimed at short scale-up runs where power, latency, and cost favor copper over optics. The Impress compression substrate connector and mating cable support 224~Gb/s PAM4 and beyond.
\newblock \href{https://www.marvell.com/blogs/224g-long-range-serdes-scale-up-scale-inside.html}{Marvell 224G long-reach SerDes} reports $\sim$4~pJ/bit and targets CPC cables, near-packaged optics, and CPO. Near-package copper (NPC) mates the connector near the ASIC rather than onto the package substrate; shared NPC/NPO sockets have been proposed. Vendor orientation, provisional.

\bibitem[FiberMall(2026)]{switchthermal}
FiberMall.
\newblock \href{https://www.fibermall.com/blog/osfp-thermal-management-guide.htm}{OSFP thermal management: data center cooling guide}.
\newblock 2026.
\newblock A 32-port 800G switch dissipates $>$1~kW, with optical modules $\sim$50\% of the thermal load; observed power runs 15--25\% above rated. Vendor orientation, provisional.

\bibitem[Cao et al.(2025)]{cpocooling}
Cao, \emph{et al.}
\newblock \href{https://doi.org/10.1007/s12200-025-00156-4}{Simulation and experimental investigation of liquid-cooling thermal management for high-bandwidth co-packaged optics}.
\newblock Front.\ Optoelectron., 2025.
\newblock Power density and on-package thermal crosstalk limit CPO reliability; liquid cooling investigated to hold optical-engine and laser temperature.

\bibitem[Semtech(2025)]{lpocost}
Semtech.
\newblock \href{https://blog.semtech.com/webinar-recap-low-power-1.6t-datacom-transceivers-and-the-path-to-3.2t}{Low-power 1.6T datacom transceivers and the path to 3.2T}.
\newblock Webinar recap, 2025.
\newblock Linear modules cited at $\sim$50--60\% less power than fully retimed; a
500{,}000-GPU cluster estimated to save $>$100~MW and $\sim$\$100~million/year in
electricity. Vendor orientation, provisional.

\bibitem[Epoch AI(2025)]{dctco}
Epoch AI.
\newblock \href{https://epoch.ai/data-insights/ai-datacenter-cost-breakdown}{Total cost of ownership of a one-gigawatt AI data center}.
\newblock Data insight, 2025.
\newblock A 1~GW AI site estimated near \$38~billion up-front capex and $\sim$\$0.9~billion/year opex. Analyst breakdown, provisional.

\end{thebibliography}
